### 2013-02-20 docs/parco13/ios-book-article.tex
Start directory for ParCo 2013 paper

\documentclass{IOS-Book-Article}

\usepackage{mathptmx}

%\usepackage{times}
%\normalfont
%\usepackage[T1]{fontenc}
%\usepackage[mtplusscr,mtbold]{mathtime}
%
\begin{document}
\begin{frontmatter}              % The preamble begins here.

%\pretitle{Pretitle}
\title{Instructions for the Preparation of an\\
Electronic Camera-Ready Manuscript in\\ \LaTeX}
\runningtitle{IOS Press Style Sample}
%\subtitle{Subtitle}

\author[A]{\fnms{Book Production} \snm{Manager}%
\thanks{Corresponding Author: Book Production Manager, IOS Press, Nieuwe Hemweg 6B,
1013 BG Amsterdam, The Netherlands; E-mail:
bookproduction@iospress.nl.}},
\author[B]{\fnms{Second} \snm{Author}}
and
\author[B]{\fnms{Third} \snm{Author}}

\runningauthor{B.P. Manager et al.}
\address[A]{Book Production Department, IOS Press, The Netherlands}
\address[B]{Short Affiliation of Second Author and Third Author}

\begin{abstract}
These instructions are designed for the Preparation of an Electronic
Camera-Ready Manuscript in \LaTeX{} and should be read carefully. If you
have any questions regarding the instructions, please contact the Book
Production Department, by e-mail: \textit{bookproduction@iospress.nl}
or fax: +31-20-6870039.
\end{abstract}

\begin{keyword}
electronic camera-ready manuscript\sep IOS Press\sep
\LaTeX\sep book\sep layout
\end{keyword}
\end{frontmatter}

\thispagestyle{empty}
\pagestyle{empty}

\section*{Introduction}
This document provides instructions for style and layout and how to submit the final
version. Although it was written for individual authors contributing to IOS Press books,
it can also be used by the author/editor preparing a monograph or an edited volume.

Authors should realize that the manuscript submitted by the volume editor to IOS
Press will be almost identical to the final, printed version that appears in the book,
except for the pagination and the insertion of running headlines. Proofreading as
regards technical content and English usage is the responsibility of the author.

A template file for \LaTeX2e is available from the Author's Corner on
www.iospress.nl. \LaTeX{} styles required for the \LaTeX{} template are also
available.\footnote{For authors using MS Word separate Instructions as well
as a Word template are available from the Author's Corner on
www.iospress.nl.}

\section{Typographical Style and Layout}

\subsection{Type Area}
The text output area is automatically set within an area 12.4 cm
horizontally and 20 cm vertically. Please do not use any
\LaTeX{} or \TeX{} commands that affect the layout or formatting of
your document (i.e. commands like \verb|\textheight|,
\verb|\textwidth|, etc.).

\subsection{Font}

The font type for running text (body text) is 10~point Times New Roman.
There is no need to code normal type (roman text). For literal text, please use
\texttt{type\-writer} (\verb|\texttt{}|)
or \textsf{sans serif} (\verb|\textsf{}|). \emph{Italic} (\verb|\emph{}|)
or \textbf{boldface} (\verb|\textbf{}|) should be used for emphasis.

\subsection{General Layout}
Use single (1.0) line spacing throughout the document. For the main
body of the paper use the commands of the standard \LaTeX{}
``article'' class. You can add packages or declare new \LaTeX{}
functions if and only if there is no conflict between your packages
and the \texttt{IOS-Book-Article}.

Always give a \verb|\label| where possible and use \verb|\ref| for cross-referencing.

\subsection{(Sub-)Section Headings}
Use the standard \LaTeX{} commands for headings: {\small \verb|\section|, \verb|\subsection|, \verb|\subsubsection|, \verb|\paragraph|}.
Headings will be automatically numbered.

Use initial capitals in the headings, except for articles (a, an, the), coordinate
conjunctions (and, or, nor), and prepositions, unless they appear at the beginning
of the heading.

\subsection{Footnotes and Endnotes}
Please keep footnotes to a minimum. If they take up more space than roughly 10\% of
the type area, list them as endnotes, before the References. Footnotes and endnotes
should both be numbered in arabic numerals and, in the case of endnotes, preceded by
the heading ``Endnotes''.

\subsection{References}

References to the literature should be mentioned in the main text by arabic numerals in
square brackets. Use the Citation-Sequence System, which means they are ``listed and
numbered in the sequence in which they are 1st cited. ($\ldots$) Subsequent citations of the
same document use the same numbers as that of its initial citation'' \cite{r1}.

As regards the content, form and punctuation of the References, if the volume
editor has not expressed a preference in this matter, authors should select the format
most appropriate to their article, and use it \textit{consistently}.

\section{Illustrations}

\subsection{General Remarks on Illustrations}
The text should include references to all illustrations. Refer to illustrations in the
text as Table~1, Table~2, Figure~1, Figure~2, etc., not with the section or chapter number
included, e.g. Table 3.2, Figure 4.3, etc. Do not use the words ``below'' or ``above''
referring to the tables, figures, etc.

Do not collect illustrations at the back of your article, but incorporate them in the
text. Position tables and figures at the top or bottom of a page, with at least 2 lines
extra space between tables or figures and the running text.

Illustrations should be centered on the page, except for small figures that can fit
side by side inside the type area. Tables and figures should not have text wrapped
alongside.

Place figure captions \textit{below} the figure, table captions \textit{above} the table.
Use bold for table/figure labels and numbers, e.g.: \textbf{Table 1.}, \textbf{Figure 2.},
and roman for the text of the caption. Keep table and figure captions justified. Center
short figure captions only.

The minimum \textit{font size} for characters in tables is 8 points, and for lettering in other
illustrations, 6 points.

On maps and other figures where a \textit{scale} is needed, use bar scales rather than
numerical ones of the type 1:10,000.

\subsection{Quality of Illustrations}
Use only Type I fonts for lettering in illustrations.

Include graphics in Encapsulated postscript (EPS) format. Do \textit{not} use illustrations
taken from the Internet. The resolution of images intended for viewing on a screen is
not sufficient for the printed version of the book.

If you are incorporating screen captures, keep in mind that the text may not be
legible after reproduction (using a screen capture tool, instead of the Print Screen
option of PC's, might help to improve the quality).

\subsection{Color Illustrations}
Please note, that illustrations will only be printed in color if the volume editor agrees to
pay the production costs for color printing. However, you should \textit{not} use color in
illustrations that must be printed in black and white, because this will greatly reduce the
print quality. (Note that in software the default often is color, so you may have to
change the settings for these illustrations.)

Illustrations that must be printed in colour should be enclosed as CMYK-encoded
EPS files.

\section{Equations}
Number equations consecutively, not section-wise. Place the numbers in parentheses at
the right-hand margin, level with the last line of the equation. Refer to equations in the
text as Eq. (1), Eqs. (3) and (5).

\section{Fine Tuning}

\subsection{Type Area}
\textbf{Check once more that all the text and illustrations are inside the type area and
that the type area is used to the maximum.} You may of course end a page with one
or more blank lines to avoid `widow' headings, or at the end of a chapter.

\subsection{Capitalization}
Use initial capitals in the title and headings, except for articles (a, an, the), coordinate
conjunctions (and, or, nor), and prepositions, unless they appear at the beginning of the
title or heading.

\subsection{Page Numbers and Running Headlines}
You do not need to include page numbers or running headlines. These elements will be
added by the publisher.

\section{Submitting the Manuscript}
Submit the following to the volume editor:

\begin{enumerate}
\item The main source file, and any other required files. Do not submit more than
one version of any item.

\item Identical high resolution PDF file with all fonts embedded. First produce a
Postscript file from \LaTeX{} with DVIPS version 5.56 or higher. The option
``-M'' (don't make fonts) should be indicated. Use Adobe Acrobat Distiller to
produce the PDF and choose the job option \textit{Press-Optimized}.
\end{enumerate}

\begin{thebibliography}{99}

\bibitem{r1}
\textit{Scientific Style and Format: The CBE manual for authors,
editors and publishers}. Style Manual Committee, Council of Biology Editors.
Sixth ed. Cambridge University Press, 1994.

\bibitem{r2}
L.U. Ante, Cem surgere: Surgite postquam sederitis, qui manducatis panem doloris,
\textit{Omnes} \textbf{13} (1916), 114--119.

\bibitem{r3}
T.X. Confortavit, \textit{Seras}, Portarum, New York, 1995.

\bibitem{r4}
P.A. Deus, Ater hoc et filius et mater praestet nobis,
\textit{Paterhoc} \textbf{66} (1993), 856--890.

\end{thebibliography}
\end{document}

### 2013-02-20 docs/parco13/abstract.tex
Start parco 2013 document

% PLEASE USE THIS FILE AS A TEMPLATE FOR THE PUBLICATION 
% Check file IOS-Book-Article.tex
%

% preferred topic area: algorithms
% 
% at most 5 keywords
% 
% relevance
% 
% originality

\documentclass{IOS-Book-Article}     %[seceqn,secfloat,secthm]

\usepackage{mathptmx}
%\usepackage[T1]{fontenc}
%\usepackage{times}%
%
%%%%%%%%%%% Put your definitions here

%%%%%%%%%%% End of definitions
\begin{document}
\begin{frontmatter}          % The preamble begins here.
%
%\pretitle{}
\title{Title}
\runningtitle{Running title}
%\subtitle{}

% For one author:
\author[A]{\fnms{Carsten} \snm{Burstedde}%
\thanks{Corresponding author.  E-mail: \texttt{burstedde@ins.uni-bonn.de}}},
\author[B]{\fnms{Donna} \snm{Calhoun}},
\author[C]{\fnms{Kyle} \snm{Mandli}} and
\author[C]{\fnms{Andy R.} \snm{Terrel}}
\runningauthor{C.\ Burstedde et.al.}
\address[A]{Institut f\"ur Numerische Simulation, Universit\"at Bonn, Germany}
\address[B]{Boise State University, Idaho, USA}
\address[C]{Institute for Computational Engineering and Sciences,\\
The University of Texas at Austin, USA}
%
% Two or more authors:
%\author[A]{\fnms{} \snm{}},
%\author[B]{\fnms{} \snm{}}
%\runningauthor{}
%\address[A]{}
%\address[B]{}
%
%%\begin{abstract}

%%\end{abstract}

\begin{keyword}
adaptive mesh refinement,
clawpack,
HPC,
manycore
\end{keyword}

\end{frontmatter}

\cite{BursteddeWilcoxGhattas11}

%%%%%%%%%%% The article body starts:

\section*{Introduction}

%%%%%%%%%%% The bibliography starts:
\bibliographystyle{siam}
\bibliography{ccgo}
\end{document}

### 2013-02-26 docs/parco13/abstract.tex
Edits
Write abstract for extended abstract

@@
 % Check file IOS-Book-Article.tex
-%
 
@@
 
+\usepackage{amsmath,amssymb}
+
+\newcommand{\sR}{\mathbb{R}}
+
 %%%%%%%%%%% End of definitions
@@
 
-% For one author:
+% Two or more authors:
+%\author[A]{\fnms{} \snm{}},
+%\author[B]{\fnms{} \snm{}}
+%\runningauthor{}
+%\address[A]{}
+%\address[B]{}
+%
 \author[A]{\fnms{Carsten} \snm{Burstedde}%
@@
 The University of Texas at Austin, USA}
+
+\begin{abstract}
 %
-% Two or more authors:
-%\author[A]{\fnms{} \snm{}},
-%\author[B]{\fnms{} \snm{}}
-%\runningauthor{}
-%\address[A]{}
-%\address[B]{}
+We present a new hybrid paradigm for parallel adaptive mesh refinement (AMR)
+that combines the scalability and lightweight architecture of tree-based AMR
+with the computational efficiency of grid-based solvers for hyperbolic
+conservation laws.  The key idea is to interpret each leaf of the AMR hierarchy
+as one uniform compute patch in $\sR^d$ of dimension $m^d$, where $m$ is
+customarily between 8 and 32.  Thus, computation on each grid can be optimized
+for speed, while we inherit the flexibility of adaptive meshes.  In our work we
+choose the p4est AMR library as mesh backend since it allows to compose the
+mesh from multiple mapped blocks, thus enabling us to work with the cubed
+sphere and other nontrivial multiblock geometries.
 %
-%%\begin{abstract}
-
-%%\end{abstract}
+\end{abstract}
 
@@
 
+\section*{}
+
+\section*{}
+
 %%%%%%%%%%% The bibliography starts:

### 2013-02-26 docs/parco13/abstract.tex
Collect some key points

--- 2013-02-20 docs/parco13/abstract.tex
@@
 
+\usepackage{amsmath,amssymb}
+\usepackage{cite}
+
+\newcommand{\sR}{\mathbb{R}}
+
 %%%%%%%%%%% End of definitions
@@
 %\pretitle{}
-\title{Title}
-\runningtitle{Running title}
+\title{ForestClaw:
+        Hybrid forest-of-octrees AMR for hyperbolic conservation laws}
+\runningtitle{ForestClaw}
 %\subtitle{}
 
-% For one author:
+% Two or more authors:
+%\author[A]{\fnms{} \snm{}},
+%\author[B]{\fnms{} \snm{}}
+%\runningauthor{}
+%\address[A]{}
+%\address[B]{}
+%
 \author[A]{\fnms{Carsten} \snm{Burstedde}%
@@
 The University of Texas at Austin, USA}
+
+\begin{abstract}
 %
-% Two or more authors:
-%\author[A]{\fnms{} \snm{}},
-%\author[B]{\fnms{} \snm{}}
-%\runningauthor{}
-%\address[A]{}
-%\address[B]{}
+We present a new hybrid paradigm for parallel adaptive mesh refinement (AMR)
+that combines the scalability and lightweight architecture of tree-based AMR
+with the computational efficiency of patch-based solvers for hyperbolic
+conservation laws.  The key idea is to interpret each leaf of the AMR hierarchy
+as one uniform compute patch in $\sR^d$ of dimensions $m^d$, where $m$ is
+customarily between 8 and 32.  Thus, computation on each patch can be optimized
+for speed, while we inherit the flexibility of adaptive meshes.  In our work we
+choose the p4est AMR library as mesh backend since it allows to compose the
+mesh from multiple mapped octrees, thus enabling us to work with the cubed
+sphere and other nontrivial multiblock geometries.
 %
-%%\begin{abstract}
-
-%%\end{abstract}
+\end{abstract}
 
@@
 
-\cite{BursteddeWilcoxGhattas11}
-
 %%%%%%%%%%% The article body starts:
@@
 
+Experience from integration with large-scale adaptive-mesh PDE solvers.
+
+Design goals: Lightweight, modular, reuse.
+
+Separation of mesh on the one hand from discretization and solvers on the other.
+
+Mesh information is discrete (tree nodes have integer coordinates in
+$(0, 2^L($ where $L$ is the maximum allowed refinement level of the tree.
+
+sorted ordering of tree nodes, lookup tables, .
+
+p4est \cite{BursteddeWilcoxGhattas11}
+
+rank as is common with parallelization using the MPI framework
+\cite{Forum94, SnirOttoHuss-LedermanEtAl96}.
+
+Challenge: patch neighborhood information is known to the mesh backend
+but needs to be used to handle ghost information.
+
+\section*{Hybrid forest-of-octrees AMR}
+
+Iterators over all leafs, optionally restricted to a given level.
+
+same block, same rank: local copy.
+
+neighbor block: permutation of patch points due to non-aligned
+coordinate systems
+
+neighbor processor: data needs to be fetched over the network first.
+
+Trees have different coordinate systems.
+
+$L$ is 30 for 2D and 19 for 3D, so it allows for deep hierarchies.
+
+\section*{Patch-based numerics at the leaf level}
+
+\section*{Numerical results}
+
 %%%%%%%%%%% The bibliography starts:
@@
 \bibliography{ccgo}
+
 \end{document}

### 2013-02-26 docs/parco13/abstract.tex
Write intro

@@
 
+With the advent of high-throughput coprocessors, such as GPGPUs or the MIC
+chip, comes the opportunity to sustain unprecedented rates of floating point
+operations at comparably high integration density and low cost.  These
+architectures, however, require careful structuring of the data layout and
+memory access patterns to fully utilize their multithreading and vectorization
+capabilities.
+
+Consequently, it is not clear a priori how to accelerate PDE solvers that use
+adaptive mesh refinement, especially when working with unstructured meshes.  It
+has been realized early though that it helps to aggregate degrees of freedom
+(DOF) at the element level, as it is done with high-order spectral element
+\cite{TufoFischer99}, finite volume, or discontinuous Galerkin
+\cite{BursteddeGhattasGurnisEtAl10} methods.
+%
+% For example, SEs optimized for computational speed have originally been
+% implemented on unstructured conforming meshes \cite{TufoFischer99} and later
+% been extended to non-conforming adaptive meshes \cite{FischerKruseLoth02,
+% BursteddeGhattasGurnisEtAl10}.
+
+To enable hardware acceleration for dynamic AMR, we like to build upon the
+forest-of-octrees paradigm because of its low overhead and parallel scalability
+\cite{BursteddeWilcoxGhattas11}.  We go beyond the traditional high-order
+element by introducing a complete compute patch with $m^d$ DOFs as our
+macro-element.  In fact, this approach resembles block-structured AMR
+\cite{ColellaGravesKeenEtAl07}
+% todo cite John B.\ Bell
+except that the patches are not overlapping, which enables us to capitalize on
+our previous experience with AMR PDE solvers
+\cite{BursteddeStadlerAlisicEtAl13}.
+
+\section*{Design principles}
+
+Now combine.
+
 Experience from integration with large-scale adaptive-mesh PDE solvers.
-
 Design goals: Lightweight, modular, reuse.
@@
 
-p4est \cite{BursteddeWilcoxGhattas11}
-
 rank as is common with parallelization using the MPI framework
-\cite{Forum94, SnirOttoHuss-LedermanEtAl96}.
+%\cite{Forum94, SnirOttoHuss-LedermanEtAl96}.
 
@@
 
+$z$-order
+
 Iterators over all leafs, optionally restricted to a given level.
+Random access is possible.  Looping over the patches in the order prescribed
+by the forest leads to a high percentage of cache reuse, since the ghost
+patches are likely closeby with respect to the $z$-order.
 
@@
 
-neighbor processor: data needs to be fetched over the network first.
+neighbor rank: data needs to be fetched over the network first.
 
@@
 
+Patch: contains the local coordinates and copies of neighboring ghost degrees
+of freedom.  The layer of ghost copies is usually 2 points deep.
+
 \section*{Numerical results}
@@
 %%%%%%%%%%% The bibliography starts:
-\bibliographystyle{siam}
+\bibliographystyle{unsrt}
 \bibliography{ccgo}

### 2013-02-26 docs/parco13/abstract.tex
Tweak for space

@@
 for speed, while we inherit the flexibility of adaptive meshes.  In our work we
-choose the p4est AMR library as mesh backend since it allows to compose the
-mesh from multiple mapped octrees, thus enabling us to work with the cubed
-sphere and other nontrivial multiblock geometries.
+choose to work with the p4est AMR library since it allows to compose the mesh
+from multiple mapped octrees, thus enabling the cubed sphere and other
+nontrivial multiblock geometries.
 %
@@
 
-\begin{keyword}
-adaptive mesh refinement,
-clawpack,
-HPC,
-manycore
-\end{keyword}
+% \begin{keyword}
+% adaptive mesh refinement,
+% clawpack,
+% HPC,
+% manycore
+% \end{keyword}
 
@@
 except that the patches are not overlapping, which enables us to capitalize on
-our previous experience with AMR PDE solvers
-\cite{BursteddeStadlerAlisicEtAl13}.
+our previous experience with AMR PDE solvers.
+% \cite{BursteddeStadlerAlisicEtAl13}.
 

### 2013-02-26 docs/parco13/abstract.tex
Extend intro

@@
 \usepackage{cite}
+\usepackage{color}
+\usepackage{xspace}
 
+\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}\xspace}
 \newcommand{\sR}{\mathbb{R}}
@@
 architectures, however, require careful structuring of the data layout and
-memory access patterns to fully utilize their multithreading and vectorization
+memory access patterns to exhaust their multithreading and vectorization
 capabilities.
@@
 Consequently, it is not clear a priori how to accelerate PDE solvers that use
-adaptive mesh refinement, especially when working with unstructured meshes.  It
-has been realized early though that it helps to aggregate degrees of freedom
-(DOF) at the element level, as it is done with high-order spectral element
-\cite{TufoFischer99}, finite volume, or discontinuous Galerkin
+adaptive mesh refinement, especially when working with unstructured meshes.  Of
+course, it has been realized early that it helps to aggregate degrees of
+freedom (DOF) at the element level, as it has been done with high-order
+spectral element \cite{TufoFischer99}, finite volume
+\todo{add favorite citation}, or discontinuous Galerkin
 \cite{BursteddeGhattasGurnisEtAl10} methods.
@@
 
-To enable hardware acceleration for dynamic AMR, we like to build upon the
-forest-of-octrees paradigm because of its low overhead and parallel scalability
-\cite{BursteddeWilcoxGhattas11}.  We go beyond the traditional high-order
-element by introducing a complete compute patch with $m^d$ DOFs as our
-macro-element.  In fact, this approach resembles block-structured AMR
-\cite{ColellaGravesKeenEtAl07}
+To enable hardware acceleration for parallel dynamic AMR, we like to build upon
+the forest-of-octrees paradigm because of its low overhead and proven
+scalability \cite{BursteddeWilcoxGhattas11}.  As an important extension, we go
+beyond the traditional high-order element and define each mesh element to be a
+dense compute patch with $m^d$ DOFs.  In fact, this approach resembles
+block-structured AMR \cite{ColellaGravesKeenEtAl07}
 % todo cite John B.\ Bell
 except that the patches are not overlapping, which enables us to capitalize on
-our previous experience with AMR PDE solvers.
-% \cite{BursteddeStadlerAlisicEtAl13}.
+our previous experience with scalable FE solvers for PDEs
+\cite{BursteddeStadlerAlisicEtAl13}.  A particular implementation of such a
+patch is available with the Clawpack software \cite{LeVeque97} that has been
+designed to solve hyperbolic conservation laws on a uniform compute patch and
+successfully used in the context of block-structured AMR \todo{AMRClaw
+citation}.
+
+In this paper we describe how we design the coupling of forest-of-octree AMR
+with Clawpack at the leaf level.  We comment on challenges that arise in
+enabling efficient parallelism and multiblock geometries and close with a range
+of numerical examples that demonstrate the conceptual improvements in relation
+to other approaches.
 

### 2013-02-26 docs/parco13/abstract.tex
Add a todo

@@
 block-structured AMR \cite{ColellaGravesKeenEtAl07}
-% todo cite John B.\ Bell
+\todo{favorite BoxLib citation}
 except that the patches are not overlapping, which enables us to capitalize on

### 2013-02-26 docs/parco13/abstract.tex
Continue working

@@
 
-Now combine.
+The starting point of our work is defined by the p4est algorithms for
+forest-of-octrees AMR on the one hand, and the Clawpack algorithms for the
+numerical solution of hyperbolic conservation laws
+% on uniformly gridded domains
+on the other.  Both are specialized codes with the following characteristics:
+\begin{center}
+\begin{tabular}{l|l|l}
+& \multicolumn{1}{c|}{\texttt{p4est}} & \multicolumn{1}{c}{Clawpack} \\
+\hline
+subject & hexahedral nonconforming mesh &  hyperbolic PDE on $[0, 1]^d$ \\
+structure & forest of octrees & patch of $m^d$ DOFs \\
+atomic unit & octree leaf & one DOF \\
+parallelization & MPI & threads \\
+memory access & distributed & shared \\
+data type & integers & floating point values \\
+language & C & Fortran 77 \\
+dependencies & none & none \\
+\hline
+\end{tabular}
+\end{center}
+The proposed 1:1 correspondence between a leaf and a patch thus combines two
+hithertwo disjoint models in a modular way that permits the reuse of existing,
+verified and performant codes if the integration is done well.  The resulting
+parallel programming model is a hybrid and often referred to as MPI+X.
+
+Multiblock handling
+
+Identification: an octree is a block, and a leaf is a patch.
+
+Each block is understood as a reference unit cube with an optional per-block
+geometric mapping.  The coordinate system of
+
+Information provided by the mesh:
+Sequence of blocks, list of patches for each.  Each patch
+has a list of neighbor patches including information on
 
@@
 
+\section*{Acknowledgements}
+
+The authors acknowledge valuable discussion with Randy LeVeque, Marsha Berger,
+and Hans-Petter Langtangen.  The leaf/grid paradigm was independently presented
+by B.\ as part of a talk at the SCI Institute, Utah, in July 2011.
+
 %%%%%%%%%%% The bibliography starts:

### 2013-02-26 docs/parco13/abstract.tex
Throw in some notes

@@
 language & C & Fortran 77 \\
-dependencies & none & none \\
+dependencies & none & Python \\
 \hline
@@
 Multiblock handling
+generic transformations
+yields periodicity as a special case
 
@@
 
+AMR Metadata $\ll$ numerical data
+
 Experience from integration with large-scale adaptive-mesh PDE solvers.
@@
 sorted ordering of tree nodes, lookup tables, .
+
+Dimension independence
 
@@
 
+fast (O(log N/P)) neighbor and parent/child lookups
+horizontal/vertical tree traversal
+
+subcycling: level-wise load balancing as done in structured AMR or
+geometric adaptive multigrid \cite{SundarBirosBursteddeEtAl12}
+
+\todo{What is BearClaw (tree-based / S.\ Mitran)?}
+
 \section*{Patch-based numerics at the leaf level}

### 2013-03-11 docs/parco13/abstract.tex
Go through first two pages of parco paper

--- 2013-02-26 docs/parco13/abstract.tex
@@
 \newcommand{\sR}{\mathbb{R}}
+\newcommand{\forestclaw}{Forestclaw\xspace}
+\newcommand{\pforest}{\texttt{p4est}\xspace}
 
@@
 %\pretitle{}
-\title{ForestClaw:
+\title{\forestclaw:
         Hybrid forest-of-octrees AMR for hyperbolic conservation laws}
-\runningtitle{ForestClaw}
+\runningtitle{\forestclaw}
 %\subtitle{}
@@
 for speed, while we inherit the flexibility of adaptive meshes.  In our work we
-choose to work with the p4est AMR library since it allows to compose the mesh
+choose to work with the \pforest AMR library since it allows to compose the mesh
 from multiple mapped octrees, thus enabling the cubed sphere and other
@@
 
-\section*{Introduction}
+\section{Introduction}
 
@@
 the forest-of-octrees paradigm because of its low overhead and proven
-scalability \cite{BursteddeWilcoxGhattas11}.  As an important extension, we go
-beyond the traditional high-order element and define each mesh element to be a
+scalability \cite{BursteddeWilcoxGhattas11}.  This approach identifies each
+octree leaf with a mesh element.  In this work, we go
+beyond the traditional high-order element and define each element to be a
 dense compute patch with $m^d$ DOFs.  In fact, this approach resembles
@@
 our previous experience with scalable FE solvers for PDEs
-\cite{BursteddeStadlerAlisicEtAl13}.  A particular implementation of such a
-patch is available with the Clawpack software \cite{LeVeque97} that has been
-designed to solve hyperbolic conservation laws on a uniform compute patch and
-successfully used in the context of block-structured AMR \todo{AMRClaw
+\cite{BursteddeStadlerAlisicEtAl13}.  The Clawpack software \cite{LeVeque97}
+provides a popular implementation of such a patch.  It has
+been designed to solve hyperbolic conservation laws
+% on a uniform compute patch
+and successfully used in the context of block-structured AMR \todo{AMRClaw
 citation}.
@@
 
-\section*{Design principles}
+\section{Design principles}
 
-Now combine.
+The starting point of our work is defined by the \pforest algorithms for
+forest-of-octrees AMR on the one hand, and the Clawpack algorithms for the
+numerical solution of hyperbolic conservation laws
+% on uniformly gridded domains
+on the other.  Both are specialized codes with the following characteristics:
+\begin{center}
+\begin{tabular}{l|l|l}
+& \multicolumn{1}{c|}{\pforest} & \multicolumn{1}{c}{Clawpack} \\
+\hline
+subject & hexahedral nonconforming mesh &  hyperbolic PDE on $[0, 1]^d$ \\
+structure & forest of octrees & patch of $m^d$ DOFs \\
+atomic unit & octree leaf & one DOF \\
+parallelization & MPI & threads (Manyclaw variant) \\
+memory access & distributed & shared \\
+data type & integers & floating point values \\
+language & C & Fortran 77 \\
+dependencies & none & Python \\
+%\hline
+\end{tabular}
+\end{center}
+The proposed 1:1 correspondence between a leaf and a patch thus combines two
+hithertwo disjoint models in a modular way:
+\begin{enumerate}
+\item We permit the reuse of existing, verified and performant codes.
+% if the integration is done well
+\item We preserve the separation of the mesh on one hand from the
+discretization and solvers on the other.
+\item The AMR metadata (24 bytes per leaf) is insignificant compared to
+the numerical data ($m^d$ floating point values per patch).
+\item The resulting parallel programming model is a hybrid and often referred
+to as MPI+X.
+\end{enumerate}
 
-Experience from integration with large-scale adaptive-mesh PDE solvers.
-Design goals: Lightweight, modular, reuse.
+% yields periodicity as a special case
+A particular feature of \forestclaw is that the generic handling of multiblock
+geometries is inherited from \pforest, identifying each octree as a block, and
+each leaf as a patch.  Each block is understood as a reference unit cube with
+its own geometric mapping.  The connectivity of the blocks is static and can be
+created by external hexahedral mesh generators, elimintating the need to encode
+it by hand.
 
-Separation of mesh on the one hand from discretization and solvers on the other.
+A main challenge is presented by the fact that the patch neighborhood is only
+known to the mesh backend, from which it needs to be propagated to the
+numerical code in \forestclaw that implements the interaction with ghost
+patches.  To this end, we define an interface that allows distributed read-only
+access to the sequence of blocks, the list of patches for each, and a lookup of
+neighbor patches (and their relative orientation which is nontrivial between
+blocks).  The exchange of ghost data and parallel partitioning is provided
+transparently by the backend.  Mesh modification directives, such as local
+refinement and coarsening, are called from \forestclaw and relayed to the mesh
+backend.
 
-Mesh information is discrete (tree nodes have integer coordinates in
-$(0, 2^L($ where $L$ is the maximum allowed refinement level of the tree.
+% AMR Metadata $\ll$ numerical data
 
-sorted ordering of tree nodes, lookup tables, .
+% Experience from integration with large-scale adaptive-mesh PDE solvers.
+% Design goals: Lightweight, modular, reuse.
 
-rank as is common with parallelization using the MPI framework
-%\cite{Forum94, SnirOttoHuss-LedermanEtAl96}.
+% Mesh information is discrete (tree nodes have integer coordinates in
+% $(0, 2^L($ where $L$ is the maximum allowed refinement level of the tree.
 
-Challenge: patch neighborhood information is known to the mesh backend
-but needs to be used to handle ghost information.
+% Dimension independence
 
-\section*{Hybrid forest-of-octrees AMR}
+% rank as is common with parallelization using the MPI framework
+% \cite{Forum94, SnirOttoHuss-LedermanEtAl96}.
+
+\section{Parallelization}
 
@@
 
-\section*{Patch-based numerics at the leaf level}
+fast (O(log N/P)) neighbor and parent/child lookups
+horizontal/vertical tree traversal
+
+subcycling: level-wise load balancing as done in structured AMR or
+geometric adaptive multigrid \cite{SundarBirosBursteddeEtAl12}
+
+\todo{What is BearClaw (tree-based / S.\ Mitran)?}
+
+\section{Patch-based numerics at the leaf level}
 
@@
 
-\section*{Numerical results}
+\section{Numerical results}
+
+\section*{Acknowledgements}
+
+The authors acknowledge valuable discussion with Randy LeVeque, Marsha Berger,
+and Hans-Petter Langtangen.  The leaf/patch paradigm was independently presented
+by B.\ as part of a talk at the SCI Institute, Utah, in July 2011.
 

### 2013-03-11 docs/parco13/abstract.tex
Some text on parallelization

@@
 \newcommand{\sR}{\mathbb{R}}
+\newcommand{\dt}{\mathrm{d}t}
 \newcommand{\forestclaw}{Forestclaw\xspace}
@@
 with Clawpack at the leaf level.  We comment on challenges that arise in
-enabling efficient parallelism and multiblock geometries and close with a range
+enabling multiblock geometries and efficient parallelism and close with a range
 of numerical examples that demonstrate the conceptual improvements in relation
@@
 the numerical data ($m^d$ floating point values per patch).
-\item The resulting parallel programming model is a hybrid and often referred
-to as MPI+X.
+\item The resulting parallel programming model is a hybrid (often referred
+to as MPI+X).
 \end{enumerate}
@@
 A main challenge is presented by the fact that the patch neighborhood is only
-known to the mesh backend, from which it needs to be propagated to the
+known to \pforest, from which it needs to be propagated to the
 numerical code in \forestclaw that implements the interaction with ghost
@@
 blocks).  The exchange of ghost data and parallel partitioning is provided
-transparently by the backend.  Mesh modification directives, such as local
-refinement and coarsening, are called from \forestclaw and relayed to the mesh
-backend.
+transparently by \pforest.  Mesh modification directives, such as local
+refinement and coarsening, are called from \forestclaw and relayed to \pforest.
+
+% Forestclaw would support swapping out p4est or Clawpack
 
@@
 
-$z$-order
+The MPI layer is addressed from within \pforest and not exposed to the
+\forestclaw code.  The order of leaves is maintained in \pforest according to a
+space filling curve.  Each MPI rank has a local view on its own partition,
+augmented where necessary with information about one layer of ghost patches.
 
-Iterators over all leafs, optionally restricted to a given level.
-Random access is possible.  Looping over the patches in the order prescribed
-by the forest leads to a high percentage of cache reuse, since the ghost
-patches are likely closeby with respect to the $z$-order.
+\forestclaw uses iterators over all leafs, optionally restricted to a given
+level.  Random access is possible and used when executing $O(1)$ time neighbor
+lookups.  Looping over the patches in the order prescribed by the forest and
+accessing neighbors only relative to the current patch leads to a high
+percentage of cache reuse.
+% since the ghost patches are likely closeby with respect to the $z$-order.
 
-same block, same rank: local copy.
+When \forestclaw accesses neighbor patches, they can be on the same or a
+different block.  In the latter case, coordinate transformations are carried
+out.  The structure of \forestclaw is oblivious to the fact that it only has a
+local view of the distributed mesh and data which relieves the developer from
+programming to the MPI interface.  Parallel neighbor exchanges are hidden
+inside \pforest and called by \forestclaw at well-defined synchronization
+points.  There is one such parallel data exchange per time step for a global
+value of $\dt$, or one exchange per discretization level per time step if $\dt$
+is chosen per-patch depending on its size (this is sometimes called
+subcycling).  When using subcycling, the load balance is attained per level as
+it is done in structured AMR or recent geometric adaptive multigrid schemes
+\cite{SundarBirosBursteddeEtAl12}.
 
-neighbor block: permutation of patch points due to non-aligned
-coordinate systems
+The threaded parallelism over the degrees of freedom of a patch is handled by
+\forestclaw alone without the need to involve \pforest.
 
-neighbor rank: data needs to be fetched over the network first.
+% Trees have different coordinate systems.
+% neighbor block: permutation of patch points due to non-aligned coordinate
+% systems
 
-Trees have different coordinate systems.
+% neighbor rank: data needs to be fetched over the network first.
 
-$L$ is 30 for 2D and 19 for 3D, so it allows for deep hierarchies.
+% $L$ is 30 for 2D and 19 for 3D, so it allows for deep hierarchies.
 
-fast (O(log N/P)) neighbor and parent/child lookups
-horizontal/vertical tree traversal
-
-subcycling: level-wise load balancing as done in structured AMR or
-geometric adaptive multigrid \cite{SundarBirosBursteddeEtAl12}
+% fast (O(log N/P)) neighbor and parent/child lookups
+% horizontal/vertical tree traversal
 

### 2013-03-14 applications/advection/2d/sphere/ghost_twopatch.f
Added some debugging commands

@@ -268,12 +268,9 @@ c     # Assume this is mapped.
 
 c     # 'iface' is relative to the coarse grid
 
-      write(6,*)'Mess with mb_average_face_ghost'
-
-      call set_bcs_to_zero(mx,my,mbc,meqn,qcoarse, qfine)
-
-
-      return
+c      write(6,*)'Mess with mb_average_face_ghost'
+c      call set_bcs_to_zero(mx,my,mbc,meqn,qcoarse, qfine)
+c      return
 
 
 
@@ -476,18 +473,18 @@ c     Average fine grid to coarse grid or copy neighboring coarse grid
       integer ifine, jfine
       double precision kf, qf, kc
 
-      write(6,*) 'WARNING : (mb_average_corner_ghost) : setting to 0'
-
-      do mq = 1,meqn
-         do ibc = 1,mbc
-            do jbc = 1,mbc
-               qcoarse(1-ibc,1-jbc,mq) = 0
-               qcoarse(mx+ibc,my+jbc,mq) = 0
-            enddo
-         enddo
-      enddo
-
-      return
+c       write(6,*) 'WARNING : (mb_average_corner_ghost) : setting to 0'
+c
+c       do mq = 1,meqn
+c          do ibc = 1,mbc
+c             do jbc = 1,mbc
+c                qcoarse(1-ibc,1-jbc,mq) = 0
+c                qcoarse(mx+ibc,my+jbc,mq) = 0
+c             enddo
+c          enddo
+c       enddo
+c
+c       return
 
 
       r2 = refratio*refratio
@@ -605,19 +602,19 @@ c     # Exchange ghost cells at block corner
       integer mq, ibc, jbc, ii, jj, r2, ifine, jfine
       double precision sum,qf,kf,kc
 
-      write(6,*) 'WARNING : (mb_average_block_corner_ghost) : ',
-     &      ' setting to 0'
-
-      do mq = 1,meqn
-         do ibc = 1,mbc
-            do jbc = 1,mbc
-               qcoarse(1-ibc,1-jbc,mq) = 0
-               qcoarse(mx+ibc,my+jbc,mq) = 0
-            enddo
-         enddo
-      enddo
-
-      return
+c       write(6,*) 'WARNING : (mb_average_block_corner_ghost) : ',
+c      &      ' setting to 0'
+c
+c       do mq = 1,meqn
+c          do ibc = 1,mbc
+c             do jbc = 1,mbc
[…]

### 2013-03-14 docs/parco13/abstract.tex
Some minor grammatical changes;  added bibliography info (see ../Biblio/amr.bib), several comments included in the text.

@@
 
+\newcommand{\comment}[1]{\textcolor{green}{[DAC: #1]}\xspace}
 \newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}\xspace}
@@
 \author[C]{\fnms{Andy R.} \snm{Terrel}}
-\runningauthor{C.\ Burstedde et.al.}
+\runningauthor{C.\ Burstedde et al.}
 \address[A]{Institut f\"ur Numerische Simulation, Universit\"at Bonn, Germany}
@@
 for speed, while we inherit the flexibility of adaptive meshes.  In our work we
-choose to work with the \pforest AMR library since it allows to compose the mesh
-from multiple mapped octrees, thus enabling the cubed sphere and other
+choose to work with the \pforest AMR library since it allows us to compose the mesh
+from multiple mapped octrees and enable the cubed sphere and other
 nontrivial multiblock geometries.
@@
 % adaptive mesh refinement,
+% wave propagation algorithms
 % clawpack,
@@
 
-Consequently, it is not clear a priori how to accelerate PDE solvers that use
-adaptive mesh refinement, especially when working with unstructured meshes.  Of
-course, it has been realized early that it helps to aggregate degrees of
-freedom (DOF) at the element level, as it has been done with high-order
-spectral element \cite{TufoFischer99}, finite volume
-\todo{add favorite citation}, or discontinuous Galerkin
+Consequently, it is not clear a priori how to accelerate PDE solvers
+that use adaptive mesh refinement, especially when working with
+unstructured meshes \comment{This sounds we are now going to talk
+  about unstructured meshes.}.  Of course, it has been realized early
+that it helps to aggregate degrees of freedom (DOF) at the element
+level, as it has been done with high-order spectral element
+\cite{TufoFischer99}, finite volume
+\cite{paramesh, chombo, samrai}, or discontinuous Galerkin
 \cite{BursteddeGhattasGurnisEtAl10} methods.
-%
+\comment{Are you looking for references to parallel AMR codes?  Or more general AMR codes based on the FVM? }
+
 % For example, SEs optimized for computational speed have originally been
@@
 
-To enable hardware acceleration for parallel dynamic AMR, we like to build upon
-the forest-of-octrees paradigm because of its low overhead and proven
-scalability \cite{BursteddeWilcoxGhattas11}.  This approach identifies each
-octree leaf with a mesh element.  In this work, we go
-beyond the traditional high-order element and define each element to be a
-dense compute patch with $m^d$ DOFs.  In fact, this approach resembles
-block-structured AMR \cite{ColellaGravesKeenEtAl07}
-\todo{favorite BoxLib citation}
-except that the patches are not overlapping, which enables us to capitalize on
-our previous experience with scalable FE solvers for PDEs
-\cite{BursteddeStadlerAlisicEtAl13}.  The Clawpack software \cite{LeVeque97}
-provides a popular implementation of such a patch.  It has
-been designed to solve hyperbolic conservation laws
+To enable hardware acceleration for parallel dynamic AMR, we like to
+build upon the forest-of-octrees paradigm because of its low overhead
+and proven scalability \cite{BursteddeWilcoxGhattas11}.  This approach
+identifies each octree leaf with a mesh element.  In this work, we go
+beyond the traditional high-order element and define each element to
+be a dense compute patch with $m^d$ DOFs.  In fact, this approach
+resembles block-structured AMR \cite{be-ol:1984, be-co:1989, be-le:1991,
+ColellaGravesKeenEtAl07} except that the
+patches are not overlapping, which enables us to capitalize on our
+previous experience with scalable FE solvers for PDEs
+\cite{BursteddeStadlerAlisicEtAl13}.  The Clawpack software
+\cite{LeVeque97} provides a popular implementation of such a patch.
+It has been designed to solve hyperbolic conservation laws
 % on a uniform compute patch
-and successfully used in the context of block-structured AMR \todo{AMRClaw
-citation}.
+and successfully used in the context of block-structured AMR
+\cite{be-le:1991, amrclaw}.
 
-In this paper we describe how we design the coupling of forest-of-octree AMR
+In this paper we describe our design for the coupling of forest-of-octree AMR
 with Clawpack at the leaf level.  We comment on challenges that arise in
-enabling multiblock geometries and efficient parallelism and close with a range
+enabling multiblock geometries and efficient parallelism and conclude with a range
 of numerical examples that demonstrate the conceptual improvements in relation
@@
 on the other.  Both are specialized codes with the following characteristics:
+\comment{I am not sure about the Python dependency;  the latest version of Clawpack relies on
+Python to generate an input file, but the code itself is F77 (or maybe some F90). }
 \begin{center}
@@
 The proposed 1:1 correspondence between a leaf and a patch thus combines two
-hithertwo disjoint models in a modular way:
+previously disjoint models in a modular way:
 \begin{enumerate}
@@
 discretization and solvers on the other.
-\item The AMR metadata (24 bytes per leaf) is insignificant compared to
+\item The AMR metadata (approx 100 bytes per leaf
+\comment{?? : in 3d: 5 ints,  9 floats; in 2d: 4 ints, 6 floats}) is insignificant compared to
 the numerical data ($m^d$ floating point values per patch).
@@
 its own geometric mapping.  The connectivity of the blocks is static and can be
-created by external hexahedral mesh generators, elimintating the need to encode
+created by external hexahedral mesh generators, eliminating the need to encode
 it by hand.
 
-A main challenge is presented by the fact that the patch neighborhood is only
-known to \pforest, from which it needs to be propagated to the
-numerical code in \forestclaw that implements the interaction with ghost
-patches.  To this end, we define an interface that allows distributed read-only
-access to the sequence of blocks, the list of patches for each, and a lookup of
-neighbor patches (and their relative orientation which is nontrivial between
-blocks).  The exchange of ghost data and parallel partitioning is provided
-transparently by \pforest.  Mesh modification directives, such as local
-refinement and coarsening, are called from \forestclaw and relayed to \pforest.
+A main challenge is presented by the fact that the patch neighborhood
+is only known to \pforest.  This patch connectivity information needs
+to be propagated to the numerical code in \forestclaw that implements
+the interaction with neighbor patches via the use of a layer of ghost
+cells surrounding each patch.  To this end, we define an interface
+that allows distributed read-only access to the sequence of blocks,
+the list of patches for each, and a lookup of neighbor patches (and
+their relative orientation which is nontrivial between blocks).  The
+exchange of ghost data and parallel partitioning is provided
+transparently by \pforest.  Mesh modification directives, such as
+local refinement and coarsening, are called from \forestclaw and
+relayed to \pforest.
 
@@
 level.  Random access is possible and used when executing $O(1)$ time neighbor
-lookups.  Looping over the patches in the order prescribed by the forest and
+lookups.  Looping over the patches in the order prescribed by the forest \comment{tree?} and
 accessing neighbors only relative to the current patch leads to a high
@@
 \bibliographystyle{unsrt}
-\bibliography{ccgo}
+\bibliography{ccgo,../Biblio/amr}
 

### 2013-03-16 docs/parco13/abstract.tex
Move my literature too and case change
Comments

@@
 \usepackage{color}
+\usepackage{url}
 \usepackage{xspace}
@@
 Consequently, it is not clear a priori how to accelerate PDE solvers
-that use adaptive mesh refinement, especially when working with
-unstructured meshes \comment{This sounds we are now going to talk
-  about unstructured meshes.}.  Of course, it has been realized early
+that use adaptive mesh refinement.
+%, especially when working with
+%unstructured meshes \comment{This sounds we are now going to talk
+%  about unstructured meshes.}.
+% CB: Good point, took it out.
+Of course, it has been realized early
 that it helps to aggregate degrees of freedom (DOF) at the element
@@
 \cite{BursteddeGhattasGurnisEtAl10} methods.
-\comment{Are you looking for references to parallel AMR codes?  Or more general AMR codes based on the FVM? }
+\comment{Are you looking for references to parallel AMR codes?  Or more general AMR codes based on the FVM?}
+\todo{No, I was looking for unstructured FV references.  Would you have any,
+and can we add the block-AMR references later?}
 
@@
 and proven scalability \cite{BursteddeWilcoxGhattas11}.  This approach
-identifies each octree leaf with a mesh element.  In this work, we go
+identifies each octree leaf with a mesh element.  In our work, we go
 beyond the traditional high-order element and define each element to
@@
 resembles block-structured AMR \cite{be-ol:1984, be-co:1989, be-le:1991,
-ColellaGravesKeenEtAl07} except that the
+ColellaGravesKeenEtAl07,Uintah} except that the
 patches are not overlapping, which enables us to capitalize on our
@@
 on the other.  Both are specialized codes with the following characteristics:
-\comment{I am not sure about the Python dependency;  the latest version of Clawpack relies on
-Python to generate an input file, but the code itself is F77 (or maybe some F90). }
+% \comment{I am not sure about the Python dependency;  the latest version of
+% Clawpack relies on
+%Python to generate an input file, but the code itself is F77 (or maybe some
+%F90). }
+% CB: I made it optional.
 \begin{center}
@@
 language & C & Fortran 77 \\
-dependencies & none & Python \\
+dependencies & none & Python (optional) \\
 %\hline
@@
 discretization and solvers on the other.
-\item The AMR metadata (approx 100 bytes per leaf
-\comment{?? : in 3d: 5 ints,  9 floats; in 2d: 4 ints, 6 floats}) is insignificant compared to
+\item The AMR metadata (24 bytes per \pforest leaf plus 92 bytes per
+forestclaw patch)
+% \comment{?? : in 3d: 5 ints,  9 floats; in 2d: 4 ints, 6 floats}
+% CB: thanks, I hav added it.
+is insignificant compared to
 the numerical data ($m^d$ floating point values per patch).
@@
 level.  Random access is possible and used when executing $O(1)$ time neighbor
-lookups.  Looping over the patches in the order prescribed by the forest \comment{tree?} and
+lookups.  Looping over the patches in the order prescribed by the forest
+% \comment{tree?}
+% CB: I think forest is fine here.
+and
 accessing neighbors only relative to the current patch leads to a high
-percentage of cache reuse.
+percentage of cache reuse
+\cite{BursteddeBurtscherGhattasEtAl09}.
 % since the ghost patches are likely closeby with respect to the $z$-order.
@@
 
-\todo{What is BearClaw (tree-based / S.\ Mitran)?}
+% \todo{What is BearClaw (tree-based / S.\ Mitran)?}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{ccgo,../Biblio/amr}
+\bibliography{../biblio/ccgo,../biblio/amr}
 

### 2013-03-16 docs/parco13/abstract.tex
Update ccgo bib file

@@
 resembles block-structured AMR \cite{be-ol:1984, be-co:1989, be-le:1991,
-ColellaGravesKeenEtAl07,Uintah} except that the
+ColellaGravesKeenEtAl07,
+BerzinsLuitjensMengEtAl10} except that the
 patches are not overlapping, which enables us to capitalize on our

### 2013-03-16 docs/parco13/abstract.tex
Be more explicit on distributed memory

@@
 parallelization & MPI & threads (Manyclaw variant) \\
-memory access & distributed & shared \\
+memory access & distributed & shared on each MPI rank \\
 data type & integers & floating point values \\
@@
 discretization and solvers on the other.
-\item The AMR metadata (24 bytes per \pforest leaf plus 92 bytes per
-forestclaw patch)
+\item The AMR metadata (\pforest:
+under 1k bytes per octree, 8 bytes per MPI rank,
+24 bytes per leaf; \forestclaw: $84 + 28d$ bytes per patch)
 % \comment{?? : in 3d: 5 ints,  9 floats; in 2d: 4 ints, 6 floats}
@@
 \item The resulting parallel programming model is a hybrid (often referred
-to as MPI+X).
+to as MPI+X).  Only rank-local leaves/patches are stored and computed.
 \end{enumerate}
@@
 cells surrounding each patch.  To this end, we define an interface
-that allows distributed read-only access to the sequence of blocks,
+that allows read-only access to the sequence of blocks,
 the list of patches for each, and a lookup of neighbor patches (and
-their relative orientation which is nontrivial between blocks).  The
+their relative orientation which is nontrivial between blocks).
+Just as \pforest, \forestclaw stores only the patches local to each
+MPI rank.  The
 exchange of ghost data and parallel partitioning is provided
@@
 
-\forestclaw uses iterators over all leafs, optionally restricted to a given
-level.  Random access is possible and used when executing $O(1)$ time neighbor
-lookups.  Looping over the patches in the order prescribed by the forest
+\forestclaw uses iterators over all rank-local leafs, optionally restricted to
+a given level.  Random access is possible and used when executing $O(1)$ time
+neighbor lookups.  Looping over the patches in the order prescribed by the
+forest
 % \comment{tree?}

### 2013-03-17 docs/parco13/abstract.tex
Added patch based numerics section

@@
 \newcommand{\dt}{\mathrm{d}t}
+\newcommand{\clawpack}{{\sc Clawpack\xspace}}
 \newcommand{\forestclaw}{Forestclaw\xspace}
@@
 % CB: Good point, took it out.
-Of course, it has been realized early
-that it helps to aggregate degrees of freedom (DOF) at the element
-level, as it has been done with high-order spectral element
-\cite{TufoFischer99}, finite volume
-\cite{paramesh, chombo, samrai}, or discontinuous Galerkin
-\cite{BursteddeGhattasGurnisEtAl10} methods.
-\comment{Are you looking for references to parallel AMR codes?  Or more general AMR codes based on the FVM?}
-\todo{No, I was looking for unstructured FV references.  Would you have any,
-and can we add the block-AMR references later?}
+Of course, was realized early that it helps to aggregate degrees of
+freedom (DOF) at the element level, as has been done with high-order
+spectral element \cite{TufoFischer99}, finite volume (element?)
+\cite{???}, or discontinuous Galerkin
+\cite{BursteddeGhattasGurnisEtAl10} methods.  The finite volume method
+has typically been implemented using a single degree of freedom per
+cell on either structured \cite{ppm, clawpack, weno} or unstructured
+meshes \cite{ddfv, ausm+, openfoam, }.  \comment{Are you looking for
+  references to parallel AMR codes?  Or more general AMR codes based
+  on the FVM?}  \todo{No, I was looking for unstructured FV
+  references.  Would you have any, and can we add the block-AMR
+  references later?} \comment{But most finite volume methods
+  (structured or unstructured) only have a single degree of freedom
+  per mesh cell (``element''?)  Only recently have people increased
+  the order by extending the stencil (without increase the DOFs per
+  mesh cell).}
 
@@
 
-To enable hardware acceleration for parallel dynamic AMR, we like to
+To enable hardware acceleration for parallel dynamic AMR, we
 build upon the forest-of-octrees paradigm because of its low overhead
@@
 BerzinsLuitjensMengEtAl10} except that the
-patches are not overlapping, which enables us to capitalize on our
+patches are not overlapping (except in ghost cell regions),
+which enables us to capitalize on our
 previous experience with scalable FE solvers for PDEs
-\cite{BursteddeStadlerAlisicEtAl13}.  The Clawpack software
+\cite{BursteddeStadlerAlisicEtAl13}.  The \clawpack software
 \cite{LeVeque97} provides a popular implementation of such a patch.
@@
 In this paper we describe our design for the coupling of forest-of-octree AMR
-with Clawpack at the leaf level.  We comment on challenges that arise in
+with \clawpack at the leaf level.  We comment on challenges that arise in
 enabling multiblock geometries and efficient parallelism and conclude with a range
@@
 The starting point of our work is defined by the \pforest algorithms for
-forest-of-octrees AMR on the one hand, and the Clawpack algorithms for the
+forest-of-octrees AMR on the one hand, and the \clawpack algorithms for the
 numerical solution of hyperbolic conservation laws
@@
 % CB: I made it optional.
+% DAC : Good.
 \begin{center}
@@
 Just as \pforest, \forestclaw stores only the patches local to each
-MPI rank.  The
-exchange of ghost data and parallel partitioning is provided
+MPI rank.  The exchange of ghost data and parallel partitioning is provided
 transparently by \pforest.  Mesh modification directives, such as
@@
 
-Patch: contains the local coordinates and copies of neighboring ghost degrees
-of freedom.  The layer of ghost copies is usually 2 points deep.
+For hyperbolic problems, we integrate the solution on a a single patch
+using the wave propagation algorithm described by R. J. LeVeque and
+implemented in {\sc Clawpack} \cite{le:2002, clawpack}.  We assume a
+single degree of freedom per mesh cell and reconstruct a piecewise
+constant solution to obtain left and right states at a mesh cell
+edges.  At each edge, we solve Riemann problems to obtain left and
+right going waves propagating at speeds determined from the solution
+to the Riemann problem.  For scalar advection, the speed of each wave
+is simply the local advection speed at the cell interface. For
+non-linear problems and systems, an approximate Riemann solver, such
+as a Roe solver, is typically used. Wave limiters are used to surpress
+spurious oscillations at sharp gradients in the flow.
+
+Data exchanges between neighboring patches are done via layers of
+ghost cells extending the dimensions along the edges of each patch.
+The interior edge values of a given patch overlap the ghost cell
+region of a neighboring patch.  For the second order wave propagation
+algorithm, two layers of ghost cells are sufficient.  Neighboring
+patches at the same level of refinement simply copy their interior
+edge values into a neighbors' ghost cells.  Neighboring fine grid
+patches average their interior edge data to a coarser neighbor's ghost
+cell values.  And neighboring coarse grid patches interpolate data
+from their interior edge cells to their fine grid neighbor's ghost
+cell values.  To avoid loss of conservation and the creation of
+spurious extrema, we use a standard conservative, limited
+interpolation scheme to interpolate values from the coarse grid to
+fine grid coarse cells \cite{amrclaw, chombo}.  When subcycling, time accurate
+data between coarse grids is used to fill in ghost cells for fine grids.
+
+Grid refinement and de-refinement also requires interpolation from coarse grids
+to newly created fine grids, and the averaging of fine grid data to a newly created
+underlying coarse grid.
 
 \section{Numerical results}
+
+* sphere results\\
+* swirl example
+
+To come...
 
@@
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha Berger,
-and Hans-Petter Langtangen.  The leaf/patch paradigm was independently presented
+and Hans-Petter Langtangen.  We also acknowledge David Ketchison and the KAUST sponsored HPC$^3$
+numerics workshop at which the initial phases of this project were first discussed.
+The leaf/patch paradigm was independently presented
 by B.\ as part of a talk at the SCI Institute, Utah, in July 2011.

### 2013-03-18 docs/parco13/abstract.tex
Go through parco13 abstract

@@
 \newcommand{\dt}{\mathrm{d}t}
-\newcommand{\clawpack}{{\sc Clawpack\xspace}}
+\newcommand{\clawpack}{{\sc Clawpack}\xspace}
 \newcommand{\forestclaw}{Forestclaw\xspace}
 \newcommand{\pforest}{\texttt{p4est}\xspace}
+\newcommand{\manyclaw}{Manyclaw\xspace}
 
@@
 % CB: Good point, took it out.
-Of course, was realized early that it helps to aggregate degrees of
+Of course, it was realized early that it helps to aggregate degrees of
 freedom (DOF) at the element level, as has been done with high-order
-spectral element \cite{TufoFischer99}, finite volume (element?)
-\cite{???}, or discontinuous Galerkin
-\cite{BursteddeGhattasGurnisEtAl10} methods.  The finite volume method
-has typically been implemented using a single degree of freedom per
-cell on either structured \cite{ppm, clawpack, weno} or unstructured
-meshes \cite{ddfv, ausm+, openfoam, }.  \comment{Are you looking for
-  references to parallel AMR codes?  Or more general AMR codes based
-  on the FVM?}  \todo{No, I was looking for unstructured FV
-  references.  Would you have any, and can we add the block-AMR
-  references later?} \comment{But most finite volume methods
-  (structured or unstructured) only have a single degree of freedom
-  per mesh cell (``element''?)  Only recently have people increased
-  the order by extending the stencil (without increase the DOFs per
-  mesh cell).}
+spectral element \cite{TufoFischer99}
+%finite volume (element?)
+or discontinuous Galerkin \cite{HesthavenWarburton02}
+methods.  GPU implementations of the latter have been proposed recently
+\cite{KlocknerWarburtonBridgeEtAl09, BursteddeGhattasGurnisEtAl10}.
+The finite volume method has typically been implemented using a single degree
+of freedom per cell on
+% either
+structured
+\cite{ppm, clawpack, weno}
+or unstructured meshes
+\cite{ddfv, ausm+, openfoam},
+where higher order methods have been constructed by widening the stencil
+\cite{needonehere}.
+% \comment{Are you looking for
+%   references to parallel AMR codes?  Or more general AMR codes based
+%   on the FVM?}  \todo{No, I was looking for unstructured FV
+%   references.  Would you have any, and can we add the block-AMR
+%   references later?} \comment{But most finite volume methods
+%   (structured or unstructured) only have a single degree of freedom
+%   per mesh cell (``element''?)  Only recently have people increased
+%   the order by extending the stencil (without increase the DOFs per
+%   mesh cell).}
+% CB: Can you fix some of the above citations?
 
@@
 
-To enable hardware acceleration for parallel dynamic AMR, we
+To facilitate hardware acceleration for parallel dynamic AMR, we
 build upon the forest-of-octrees paradigm because of its low overhead
@@
 be a dense compute patch with $m^d$ DOFs.  In fact, this approach
-resembles block-structured AMR \cite{be-ol:1984, be-co:1989, be-le:1991,
+resembles block-structured AMR \cite{be-ol:1984, be-co:1989,
+% be-le:1991 (cited below)
 ColellaGravesKeenEtAl07,
 BerzinsLuitjensMengEtAl10} except that the
-patches are not overlapping (except in ghost cell regions),
+patches are not overlapping,
+% (except in ghost cell regions),
+% CB: we comment on this later in Section 4.
 which enables us to capitalize on our
@@
 % \comment{I am not sure about the Python dependency;  the latest version of
-% Clawpack relies on
+% \clawpack relies on
 %Python to generate an input file, but the code itself is F77 (or maybe some
@@
 % CB: I made it optional.
-% DAC : Good.
+% DAC: Good.
 \begin{center}
 \begin{tabular}{l|l|l}
-& \multicolumn{1}{c|}{\pforest} & \multicolumn{1}{c}{Clawpack} \\
+& \multicolumn{1}{c|}{\pforest} & \multicolumn{1}{c}{\clawpack} \\
 \hline
 subject & hexahedral nonconforming mesh &  hyperbolic PDE on $[0, 1]^d$ \\
-structure & forest of octrees & patch of $m^d$ DOFs \\
-atomic unit & octree leaf & one DOF \\
-parallelization & MPI & threads (Manyclaw variant) \\
+toplevel unit & forest of octrees & patch of $m^d$ FV cells \\
+atomic unit & octree leaf & one DOF in each cell \\
+parallelization & MPI & threads (\manyclaw variant) \\
 memory access & distributed & shared on each MPI rank \\
@@
 \end{center}
-The proposed 1:1 correspondence between a leaf and a patch thus combines two
+Each leaf as the atomic unit of \pforest houses a toplevel unit of \clawpack.
+The term cell is used to identify a single DOF within a \clawpack patch.  The
+proposed 1:1 correspondence between a leaf and a patch thus combines two
 previously disjoint models in a modular way:
 \begin{enumerate}
-\item We permit the reuse of existing, verified and performant codes.
+\item We permit the reuse of existing, verified, and performant codes.
 % if the integration is done well
-\item We preserve the separation of the mesh on one hand from the
+\item We preserve the separation between the mesh on one hand and the
 discretization and solvers on the other.
@@
 under 1k bytes per octree, 8 bytes per MPI rank,
-24 bytes per leaf; \forestclaw: $84 + 28d$ bytes per patch)
+24 bytes per leaf. \forestclaw: $84 + 28d$ bytes per patch)
 % \comment{?? : in 3d: 5 ints,  9 floats; in 2d: 4 ints, 6 floats}
-% CB: thanks, I hav added it.
+% CB: thanks, I have added it.
 is insignificant compared to
@@
 \item The resulting parallel programming model is a hybrid (often referred
-to as MPI+X).  Only rank-local leaves/patches are stored and computed.
+to as MPI+X).  Only rank-local leaves/patches are stored and computed on.
 \end{enumerate}
@@
 A particular feature of \forestclaw is that the generic handling of multiblock
-geometries is inherited from \pforest, identifying each octree as a block, and
-each leaf as a patch.  Each block is understood as a reference unit cube with
-its own geometric mapping.  The connectivity of the blocks is static and can be
-created by external hexahedral mesh generators, eliminating the need to encode
-it by hand.
+geometries is inherited from \pforest, identifying each octree as a block.
+% and each leaf as a patch.
+Each block is understood as a reference unit cube with
+its own geometric mapping.  The connectivity of the blocks can be created by
+external hexahedral mesh generators, eliminating the need to encode it by hand.
 
@@
 their relative orientation which is nontrivial between blocks).
-Just as \pforest, \forestclaw stores only the patches local to each
-MPI rank.  The exchange of ghost data and parallel partitioning is provided
-transparently by \pforest.  Mesh modification directives, such as
-local refinement and coarsening, are called from \forestclaw and
-relayed to \pforest.
+Suitably informed by \pforest, \forestclaw stores only the patches local to
+each MPI rank.  The exchange of ghost data and parallel repartitioning is
+provided transparently by \pforest.  Mesh modification directives, such as
+adaptive refinement and coarsening, are called from \forestclaw and relayed to
+\pforest.
 
-% Forestclaw would support swapping out p4est or Clawpack
+% Forestclaw would support swapping out p4est or \clawpack
 
@@
 space filling curve.  Each MPI rank has a local view on its own partition,
-augmented where necessary with information about one layer of ghost patches.
+augmented where necessary with information about one layer of ghost leaves.
 
@@
 subcycling).  When using subcycling, the load balance is attained per level as
-it is done in structured AMR or recent geometric adaptive multigrid schemes
+it is done for example in Chombo \cite{ColellaGravesKeenEtAl07}
+or recent geometric adaptive multigrid schemes
 \cite{SundarBirosBursteddeEtAl12}.
 
-The threaded parallelism over the degrees of freedom of a patch is handled by
-\forestclaw alone without the need to involve \pforest.
+The threaded parallelism over the degrees of freedom of a patch can be handled
+by \forestclaw alone without the need to involve \pforest.
 
@@
 
-For hyperbolic problems, we integrate the solution on a a single patch
+% CB: I'm trying to avoid "mesh" cell here since in \pforest FE simulations
+% that would mean a whole leaf.
+
+For hyperbolic problems, we integrate the solution on a a single uniform
+patch, containing $m^d$ cells,
 using the wave propagation algorithm described by R. J. LeVeque and
-implemented in {\sc Clawpack} \cite{le:2002, clawpack}.  We assume a
-single degree of freedom per mesh cell and reconstruct a piecewise
-constant solution to obtain left and right states at a mesh cell
+implemented in \clawpack \cite{le:2002, clawpack}.  We assume a
+single degree of freedom per cell and reconstruct a piecewise
+constant solution to obtain left and right states at cell
 edges.  At each edge, we solve Riemann problems to obtain left and
@@
 to the Riemann problem.  For scalar advection, the speed of each wave
-is simply the local advection speed at the cell interface. For
-non-linear problems and systems, an approximate Riemann solver, such
-as a Roe solver, is typically used. Wave limiters are used to surpress
+is simply the local advection speed at the cell interface.  For non-linear
+problems and systems, an approximate Riemann solver, such as a Roe solver
+\cite{roesolver}, is typically used.  Wave limiters are used to surpress
 spurious oscillations at sharp gradients in the flow.
@@
 region of a neighboring patch.  For the second order wave propagation
-algorithm, two layers of ghost cells are sufficient.  Neighboring
+algorithm, two layers of ghost cells are sufficient.  This implies that
+one layer of ghost patches is sufficient for $m \ge 4$.  Neighboring
 patches at the same level of refinement simply copy their interior
@@
 data between coarse grids is used to fill in ghost cells for fine grids.
+As mentioned in the previous section, this procedure can be extended
+transparently to distributed parallelism by defining an abstract exchange
+routine for ghost patch data.
 
-Grid refinement and de-refinement also requires interpolation from coarse grids
-to newly created fine grids, and the averaging of fine grid data to a newly created
-underlying coarse grid.
+Grid refinement and coarsening requires interpolation from coarse grids
+to newly created fine grids, and the averaging of fine grid data to a newly
+created underlying coarse grid.  This operation is rank-local, analogous to the
+general dynamic AMR procedures used in \pforest-based FE codes \cite[Fig.\
+4]{BursteddeGhattasStadlerEtAl08}.
 
@@
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha Berger,
-and Hans-Petter Langtangen.  We also acknowledge David Ketchison and the KAUST sponsored HPC$^3$
-numerics workshop at which the initial phases of this project were first discussed.
+and Hans-Petter Langtangen.  We also acknowledge David Ketcheson and the KAUST
+sponsored HPC$^3$ numerics workshop at which the initial phases of this project
+were first discussed.
 The leaf/patch paradigm was independently presented

### 2013-03-18 docs/parco13/abstract.tex
Change some FV wording, add references

@@
 structured
-\cite{ppm, clawpack, weno}
+\cite{ppm, clawpack}
 or unstructured meshes
 \cite{ddfv, ausm+, openfoam},
-where higher order methods have been constructed by widening the stencil
-\cite{needonehere}.
+where higher order methods have been constructed by widening the stencil, for instance in WENO methods
+\cite{Shu:2009bi}.
 % \comment{Are you looking for
@@
 and successfully used in the context of block-structured AMR
-\cite{be-le:1991, amrclaw}.
+\cite{be-le:1991, amrclaw, Berger:2011du}.
 
@@
 
-For hyperbolic problems, we integrate the solution on a a single uniform
-patch, containing $m^d$ cells,
-using the wave propagation algorithm described by R. J. LeVeque and
-implemented in \clawpack \cite{le:2002, clawpack}.  We assume a
-single degree of freedom per cell and reconstruct a piecewise
-constant solution to obtain left and right states at cell
-edges.  At each edge, we solve Riemann problems to obtain left and
-right going waves propagating at speeds determined from the solution
-to the Riemann problem.  For scalar advection, the speed of each wave
-is simply the local advection speed at the cell interface.  For non-linear
-problems and systems, an approximate Riemann solver, such as a Roe solver
-\cite{roesolver}, is typically used.  Wave limiters are used to surpress
-spurious oscillations at sharp gradients in the flow.
+For hyperbolic problems, we integrate the solution on a a single uniform patch,
+containing $m^d$ cells, using the wave propagation algorithm described by R. J.
+LeVeque \cite{LeVeque:1997eg} and implemented in \clawpack \cite{le:2002,
+clawpack}.  We assume a single degree of freedom per cell and reconstruct a
+piecewise constant solution to obtain left and right states at cell edges.  At
+each edge, we solve Riemann problems to obtain left and right going waves
+propagating at speeds determined from the solution to the Riemann problem.  For
+scalar advection, the speed of each wave is simply the local advection speed at
+the cell interface.  For non- linear problems and systems, an approximate
+Riemann solver, such as a Roe solver \cite{roesolver}, is typically used.  Since
+much of the physics of an application can be contained in the Riemann solver,
+\forestclaw adopts \clawpack's interface to Riemann solvers effectively allowing
+problems solved with \clawpack to be solvable in \forestclaw.  In order to
+achieve second order accuracy wave limiters are used to suppress spurious
+oscillations at sharp gradients in the flow.
 
@@
 interpolation scheme to interpolate values from the coarse grid to
-fine grid coarse cells \cite{amrclaw, chombo}.  When subcycling, time accurate
+fine grid coarse cells \cite{amrclaw, chombo}.  When sub-cycling, time accurate
 data between coarse grids is used to fill in ghost cells for fine grids.

### 2013-03-18 docs/parco13/abstract.tex
Incorporate two comments of Andy and Kyle

--- 2013-03-18 docs/parco13/abstract.tex
@@
 language & C & Fortran 77 \\
-dependencies & none & Python (optional) \\
+% CB: dropped dependencies.
+% dependencies & none & Python (optional) \\
 %\hline
@@
 using the wave propagation algorithm described by R. J. LeVeque and
-implemented in \clawpack \cite{le:2002, clawpack}.  We assume a
+implemented in \clawpack \cite{le:2002, clawpack}.  Optionally, the coordinates
+of the patch can be geometrically mapped.  We assume a
 single degree of freedom per cell and reconstruct a piecewise

### 2013-03-18 docs/parco13/abstract.tex
Add a forest picture

@@
 \usepackage{color}
+\usepackage{graphicx}
 \usepackage{url}
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=.4\columnwidth]{../figs_misc/cubed3}
+\caption{The leaves in a forest of six octrees that serves as the computational
+domain for the cubed sphere.  An adhoc refinement pattern has been 2:1 balanced
+in terms of neighbor sizes and partitioned between five MPI processes (the
+three middle ones are shown from black to white).}
+\label{cubed3}
+\end{figure}
+
 The MPI layer is addressed from within \pforest and not exposed to the
@@
 space filling curve.  Each MPI rank has a local view on its own partition,
-augmented where necessary with information about one layer of ghost leaves.
+augmented where necessary with information about one layer of ghost leaves
+(see Figure~\ref{cubed3}).
 

### 2013-03-18 docs/parco13/abstract.tex
More bits
Add more delectable references and some text.

--- 2013-03-18 docs/parco13/abstract.tex
@@
 
-With the advent of high-throughput coprocessors, such as GPGPUs or the MIC
-chip, comes the opportunity to sustain unprecedented rates of floating point
+With the advent of many-core chips, such as GPUs and the MIC architecture, comes the opportunity to sustain unprecedented rates of floating point
 operations at comparably high integration density and low cost.  These
@@
 freedom (DOF) at the element level, as has been done with high-order
-spectral element \cite{TufoFischer99}
+spectral element \cite{TufoFischer99}, low order continuous Galerkin via methods that accumulate many elements simultaneously \cite{knepleyterrel:2013},
 %finite volume (element?)
@@
 beyond the traditional high-order element and define each element to
-be a dense compute patch with $m^d$ DOFs.  In fact, this approach
+be a dense computational patch with $m^d$ DOFs.  In fact, this approach
 resembles block-structured AMR \cite{be-ol:1984, be-co:1989,
@@
 language & C & Fortran 77 \\
-dependencies & none & Python (optional) \\
+% dependencies & none & Python (optional) \\
 %\hline
@@
 is chosen per-patch depending on its size (this is sometimes called
-subcycling).  When using subcycling, the load balance is attained per level as
+sub-cycling).  When using sub-cycling, the load balance is attained per level as
 it is done for example in Chombo \cite{ColellaGravesKeenEtAl07}
@@
 by \forestclaw alone without the need to involve \pforest.
+Additionally many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for the hyperbolic integration on a leaf-patch.
+The design of leaf-patches allows for the smart management of data local to accelerator and the host.
 
@@
 
-For hyperbolic problems, we integrate the solution on a a single uniform patch,
+For hyperbolic problems, we integrate the solution on a single uniform patch,
 containing $m^d$ cells, using the wave propagation algorithm described by R. J.

### 2013-03-18 docs/parco13/abstract.tex

@@
 
-% \begin{keyword}
-% adaptive mesh refinement,
-% wave propagation algorithms
-% clawpack,
-% HPC,
-% manycore
-% \end{keyword}
+\begin{keyword}
+adaptive mesh refinement,
+hyperbolic conservation laws,
+clawpack,
+HPC,
+manycore
+\end{keyword}
 
@@
 language & C & Fortran 77 \\
+% CB: dropped dependencies.
 % dependencies & none & Python (optional) \\
@@
 
+Parallelization is not completed yet but envisioned for the final version of
+the paper.
+
 \section*{Acknowledgements}

### 2013-03-18 docs/parco13/abstract.tex
Typo
Beef up figure

@@
 \begin{figure}
-\centering
+\begin{center}
+\includegraphics[width=.4\columnwidth]{../figs_misc/forest_stacked}
+\hspace{.05\columnwidth}
 \includegraphics[width=.4\columnwidth]{../figs_misc/cubed3}
-\caption{The leaves in a forest of six octrees that serves as the computational
-domain for the cubed sphere.  An adhoc refinement pattern has been 2:1 balanced
-in terms of neighbor sizes and partitioned between five MPI processes (the
-three middle ones are shown from black to white).}
+\end{center}
+\caption{Left: Forest of two quadtrees, partitioned among three MPI processes.
+Each quadtree has its own coordinate orientation.  The situation in 3D
+(octrees) is analogous.  Right: The leaves in a forest of six quadtrees that
+serves as the computational domain for the cubed sphere.  An adhoc refinement
+pattern has been 2:1 balanced in terms of neighbor sizes and partitioned
+between five MPI processes (the three middle ones are shown from black to
+white).}
 \label{cubed3}
@@
 scalar advection, the speed of each wave is simply the local advection speed at
-the cell interface.  For non- linear problems and systems, an approximate
+the cell interface.  For non-linear problems and systems, an approximate
 Riemann solver, such as a Roe solver \cite{roesolver}, is typically used.  Since

### 2013-03-18 docs/parco13/abstract.tex
Add text and figure for results section

@@
 
-* sphere results\\
-* swirl example
+\begin{figure}
+\begin{center}
+\includegraphics[width=.4\columnwidth]{quadtree}
+\hspace{.05\columnwidth}
+\includegraphics[width=.4\columnwidth]{sphere_plot167}
+\end{center}
+\caption{Numerical results for solving the advection equation.  Left:  Unit
+square (single quadtree) with a twisting velocity field.  Right: Spherical ball
+with a rotational velocity field constructed from two mapped quadtrees.  In
+both cases the concentration is color-coded with a sharp gradient shown in red.
+The adaptive mesh refinement follows the location of the gradient (the patches
+are not shown where they become too fine for display).  Here we use $m = 8$
+cells per \clawpack patch.}
+\label{results2d}
+\end{figure}
 
-To come...
+The results in this section are preliminary and will be extended in the final
+version.  We have so far exercised \forestclaw for two 2D manifolds, namely the
+unit square and a sphere composed of two mapped squares.  On both we solve the
+advection equation with a given velocity field.  We advect a concentration
+field with a sharp gradient which we track dynamically with adaptive mesh
+refinement.  We observe that the refinement follows the front and allows for an
+accurate representation of its location over time; see Figure~\ref{results2d}.
 
-Parallelization is not completed yet but envisioned for the final version of
-the paper.
+\pforest is parallelized with MPI and well tested on large supercomputers, both
+in 2D and 3D.  Parallelization of \forestclaw is not completed yet but
+envisioned for the final version of the paper.
 

### 2013-03-18 docs/parco13/abstract.tex
Remove two undefined references
Minor changes

@@
 \usepackage{xspace}
+\usepackage{graphicx}
 
@@
 \newcommand{\manyclaw}{Manyclaw\xspace}
+
+% \newcommand{\plotbox}[1]{\fbox{#1}}
+\newcommand{\plotbox}[1]{#1}
+
+\newcommand{\Fig}[1]{Figure \ref{fig:#1}}
 
@@
 or unstructured meshes
-\cite{ddfv, ausm+, openfoam},
+\cite{li:2010, he-hu:2008, do-om:2005, li-sh-sv-va:2007, openfoam},
 where higher order methods have been constructed by widening the stencil, for instance in WENO methods
@@
 \section{Numerical results}
+Below, we provide two examples that demonstrate the \forestclaw code.
+The numerical results to date have been designed to verify that
+interface layer between \pforest and \forestclaw is sufficiently
+flexible and robust.  Of particular importance was ensuring that all
+ghost cell ghost cell transfers (including averaging and
+interpolation) have been implemented correctly.
+The basic \clawpack algorithm and corresponding code have been
+thoroughly tested, and so needed no further verification here.
+
+In both sets of numerical results, we solve a scalar advection
+equation, given by
+\begin{equation*}
+q_t + ({\bf u}\cdot \nabla) q = 0
+\end{equation*}
+\noindent
+where the velocity field ${\bf u} = (u(\xi,\eta,t), v(\xi,\eta,t)$ is
+a prescribed function of computational coordinates.  The relevant
+numerical parameters that are set in each case include the size of the
+patch on each leaf ($m=8$ in each case), and the minimum and maximum
+levels, which in turn fix minimum and maximum effective grid
+resolutions.
+
+In both examples, a patch is tagged for refinement if the
+difference between its maximum and minimum values exceed a prescribed
+threshold.  A family of patches (4 sibling patches in 2d) is coarsened
+if a coarsened parent patch did not meet the criteria for refinement.
+
+\noindent
+{\bf Example 1 : } In this example, an initial concentration field $q$
+is set to 0 in the left half of a computational square and to 1 in the
+right half.  A time dependent flow field is prescribed that distorts
+the interface between the 0 and 1 concentration values.  The minimum
+and maximum levels of refinement were set to 3 and 6, respectively.
+This results in a minimum resolution of $64 \times 64$ and a maximum
+resolution of $512 \times 512$. \Fig{results2d} shows the results
+at an intermediate time step.
+
+% \begin{figure}
+% \begin{center}
+% \plotbox{\includegraphics[height=0.485\textwidth,clip=true,trim=3cm 0cm 2cm 0cm]{quadtree.png}} \hfil
+% \plotbox{\includegraphics[height=0.485\textwidth,clip=true,trim=6cm 0cm 7cm 0cm]{sphere_plot167.png}}
+% \end{center}
+% \caption{Results for scalar advection using \forestclaw.}
+% \label{fig:results}
+% \end{figure}
+
+\noindent {\bf Example 2 : } This example, we demonstrate the
+multiblock functionality of \forestclaw by considering flow on a
+sphere.  The sphere mapping we use consists of two quadtrees, each
+defined in the computational space $[-1,1] \times [-1,1]$
+\cite{ca-he-le:2008, be-ca-he-le:2009}.  Each quadtree is mapped to cover one hemisphere
+of the sphere.  The initial conditions were set to 0 in one half of
+the sphere, and 1 in the other half, where the halves are not
+necessarily aligned with the equator of the mapping.  The flow field
+${u}$ simulates rigid body rotation.  We show the results at an intermediate time in
+\Fig{results2d}.
 
@@
 \begin{center}
-\includegraphics[width=.4\columnwidth]{quadtree}
-\hspace{.05\columnwidth}
-\includegraphics[width=.4\columnwidth]{sphere_plot167}
+\plotbox{\includegraphics[width=.4\columnwidth]{quadtree}} \hfil
+% \hspace{.05\columnwidth}
+\plotbox{\includegraphics[width=.4\columnwidth]{sphere_plot167}}
 \end{center}
@@
 cells per \clawpack patch.}
-\label{results2d}
+\label{fig:results2d}
 \end{figure}
@@
 refinement.  We observe that the refinement follows the front and allows for an
-accurate representation of its location over time; see Figure~\ref{results2d}.
+accurate representation of its location over time; see \Fig{results2d}.
 
@@
 
-The authors acknowledge valuable discussion with Randy LeVeque, Marsha Berger,
-and Hans-Petter Langtangen.  We also acknowledge David Ketcheson and the KAUST
-sponsored HPC$^3$ numerics workshop at which the initial phases of this project
-were first discussed.
-The leaf/patch paradigm was independently presented
-by B.\ as part of a talk at the SCI Institute, Utah, in July 2011.
+The authors acknowledge valuable discussion with Randy LeVeque, Marsha
+Berger, and Hans-Petter Langtangen.  We also acknowledge David
+Ketcheson and the KAUST sponsored HPC$^3$ numerics workshop at which
+the initial phases of this project were first discussed.  The second
+author would like to also acknowledge the Isaac Newton Institute
+(Cambridge, UK), where much of the preliminary development work for
+\forestclaw was done. The leaf/patch paradigm was independently
+presented by B.\ as part of a talk at the SCI Institute, Utah, in July
+2011.
 

### 2013-03-18 docs/parco13/abstract.tex
Minor changes

@@
 \usepackage{color}
-\usepackage{graphicx}
 \usepackage{url}
@@
 
-\begin{keyword}
-adaptive mesh refinement,
-hyperbolic conservation laws,
-clawpack,
-HPC,
-manycore
-\end{keyword}
+% \begin{keyword}
+% adaptive mesh refinement,
+% wave propagation algorithms
+% clawpack,
+% HPC,
+% manycore
+% \end{keyword}
 
@@
 
-With the advent of many-core chips, such as GPUs and the MIC architecture, comes the opportunity to sustain unprecedented rates of floating point
+With the advent of high-throughput coprocessors, such as GPGPUs or the MIC
+chip, comes the opportunity to sustain unprecedented rates of floating point
 operations at comparably high integration density and low cost.  These
@@
 freedom (DOF) at the element level, as has been done with high-order
-spectral element \cite{TufoFischer99}, low order continuous Galerkin via methods that accumulate many elements simultaneously \cite{knepleyterrel:2013},
+spectral element \cite{TufoFischer99}
 %finite volume (element?)
@@
 structured
-\cite{ppm, clawpack}
+\cite{ppm, clawpack, weno, central schemes}
 or unstructured meshes
-\cite{li:2010, he-hu:2008, do-om:2005, li-sh-sv-va:2007, openfoam},
-where higher order methods have been constructed by widening the stencil, for instance in WENO methods
-\cite{Shu:2009bi}.
+\cite{ddfv, ausm+, openfoam},
+where higher order methods have been constructed by widening the stencil
+\cite{needonehere}.
 % \comment{Are you looking for
@@
 beyond the traditional high-order element and define each element to
-be a dense computational patch with $m^d$ DOFs.  In fact, this approach
-resembles block-structured AMR \cite{be-ol:1984, be-co:1989,
+be a dense compute patch with $m^d$ DOFs.  In fact, this approach
+resembles \comment{is essentially?} block-structured AMR \cite{be-ol:1984, be-co:1989,
 % be-le:1991 (cited below)
@@
 and successfully used in the context of block-structured AMR
-\cite{be-le:1991, amrclaw, Berger:2011du}.
+\cite{be-le:1991, amrclaw}.
 
@@
 language & C & Fortran 77 \\
-% CB: dropped dependencies.
-% dependencies & none & Python (optional) \\
+dependencies & none & Python (optional) \\
 %\hline
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=.4\columnwidth]{../figs_misc/forest_stacked}
-\hspace{.05\columnwidth}
-\includegraphics[width=.4\columnwidth]{../figs_misc/cubed3}
-\end{center}
-\caption{Left: Forest of two quadtrees, partitioned among three MPI processes.
-Each quadtree has its own coordinate orientation.  The situation in 3D
-(octrees) is analogous.  Right: The leaves in a forest of six quadtrees that
-serves as the computational domain for the cubed sphere.  An adhoc refinement
-pattern has been 2:1 balanced in terms of neighbor sizes and partitioned
-between five MPI processes (the three middle ones are shown from black to
-white).}
-\label{cubed3}
-\end{figure}
-
 The MPI layer is addressed from within \pforest and not exposed to the
@@
 space filling curve.  Each MPI rank has a local view on its own partition,
-augmented where necessary with information about one layer of ghost leaves
-(see Figure~\ref{cubed3}).
+augmented where necessary with information about one layer of ghost leaves.
 
@@
 is chosen per-patch depending on its size (this is sometimes called
-sub-cycling).  When using sub-cycling, the load balance is attained per level as
+subcycling).  When using subcycling, the load balance is attained per level as
 it is done for example in Chombo \cite{ColellaGravesKeenEtAl07}
@@
 by \forestclaw alone without the need to involve \pforest.
-Additionally many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for the hyperbolic integration on a leaf-patch.
-The design of leaf-patches allows for the smart management of data local to accelerator and the host.
 
@@
 
-For hyperbolic problems, we integrate the solution on a single uniform patch,
-containing $m^d$ cells, using the wave propagation algorithm described by R. J.
-LeVeque \cite{LeVeque:1997eg} and implemented in \clawpack \cite{le:2002,
-clawpack}.  We assume a single degree of freedom per cell and reconstruct a
-piecewise constant solution to obtain left and right states at cell edges.  At
-each edge, we solve Riemann problems to obtain left and right going waves
-propagating at speeds determined from the solution to the Riemann problem.  For
-scalar advection, the speed of each wave is simply the local advection speed at
-the cell interface.  For non-linear problems and systems, an approximate
-Riemann solver, such as a Roe solver \cite{roesolver}, is typically used.  Since
-much of the physics of an application can be contained in the Riemann solver,
-\forestclaw adopts \clawpack's interface to Riemann solvers effectively allowing
-problems solved with \clawpack to be solvable in \forestclaw.  In order to
-achieve second order accuracy wave limiters are used to suppress spurious
-oscillations at sharp gradients in the flow.
+For hyperbolic problems, we integrate the solution on a a single uniform
+patch, containing $m^d$ cells,
+using the wave propagation algorithm described by R. J. LeVeque and
+implemented in \clawpack \cite{le:2002, clawpack}.  We assume a
+single degree of freedom per cell and reconstruct a piecewise
+constant solution to obtain left and right states at cell
+edges.  At each edge, we solve Riemann problems to obtain left and
+right going waves propagating at speeds determined from the solution
+to the Riemann problem.  For scalar advection, the speed of each wave
+is simply the local advection speed at the cell interface.  For non-linear
+problems and systems, an approximate Riemann solver, such as a Roe solver
+\cite{roesolver}, is typically used.  Wave limiters are used to surpress
+spurious oscillations at sharp gradients in the flow.
 
@@
 interpolation scheme to interpolate values from the coarse grid to
-fine grid coarse cells \cite{amrclaw, chombo}.  When sub-cycling, time accurate
+fine grid coarse cells \cite{amrclaw, chombo}.  When subcycling, time accurate
 data between coarse grids is used to fill in ghost cells for fine grids.
@@
 ghost cell ghost cell transfers (including averaging and
-interpolation) have been implemented correctly.
+interpolation) have been implemented correctly.  \comment{Parallel performance results will be demonstrated in the full paper?}
 The basic \clawpack algorithm and corresponding code have been
@@
 This results in a minimum resolution of $64 \times 64$ and a maximum
-resolution of $512 \times 512$. \Fig{results2d} shows the results
+resolution of $512 \times 512$. Figure \Fig{swirl} shows the results
 at an intermediate time step.
 
-% \begin{figure}
-% \begin{center}
-% \plotbox{\includegraphics[height=0.485\textwidth,clip=true,trim=3cm 0cm 2cm 0cm]{quadtree.png}} \hfil
-% \plotbox{\includegraphics[height=0.485\textwidth,clip=true,trim=6cm 0cm 7cm 0cm]{sphere_plot167.png}}
-% \end{center}
-% \caption{Results for scalar advection using \forestclaw.}
-% \label{fig:results}
-% \end{figure}
+\begin{figure}
+\begin{center}
+\plotbox{\includegraphics[height=0.485\textwidth,clip=true,trim=3cm 0cm 2cm 0cm]{quadtree.png}} \hfil
+\plotbox{\includegraphics[height=0.485\textwidth,clip=true,trim=6cm 0cm 7cm 0cm]{sphere_plot167.png}}
+\end{center}
+\caption{Results for scalar advection using \forestclaw.}
+\label{fig:results}
+\end{figure}
 
@@
 ${u}$ simulates rigid body rotation.  We show the results at an intermediate time in
-\Fig{results2d}.
-
-\begin{figure}
-\begin{center}
-\plotbox{\includegraphics[width=.4\columnwidth]{quadtree}} \hfil
-% \hspace{.05\columnwidth}
-\plotbox{\includegraphics[width=.4\columnwidth]{sphere_plot167}}
-\end{center}
-\caption{Numerical results for solving the advection equation.  Left:  Unit
-square (single quadtree) with a twisting velocity field.  Right: Spherical ball
-with a rotational velocity field constructed from two mapped quadtrees.  In
-both cases the concentration is color-coded with a sharp gradient shown in red.
-The adaptive mesh refinement follows the location of the gradient (the patches
-are not shown where they become too fine for display).  Here we use $m = 8$
-cells per \clawpack patch.}
-\label{fig:results2d}
-\end{figure}
-
-The results in this section are preliminary and will be extended in the final
-version.  We have so far exercised \forestclaw for two 2D manifolds, namely the
-unit square and a sphere composed of two mapped squares.  On both we solve the
-advection equation with a given velocity field.  We advect a concentration
-field with a sharp gradient which we track dynamically with adaptive mesh
-refinement.  We observe that the refinement follows the front and allows for an
-accurate representation of its location over time; see \Fig{results2d}.
-
-\pforest is parallelized with MPI and well tested on large supercomputers, both
-in 2D and 3D.  Parallelization of \forestclaw is not completed yet but
-envisioned for the final version of the paper.
+\Fig{results}.
 

### 2013-03-19 docs/parco13/abstract.tex
Added ausm and ddfv refs (for future purposes?)
Included description of numerical examples that didn't get committed this morning

--- 2013-03-18 docs/parco13/abstract.tex
@@
 or unstructured meshes
-\cite{li:2010, he-hu:2008, do-om:2005, li-sh-sv-va:2007, openfoam},
+\cite{%ddfv, ausm+,
+openfoam},
 where higher order methods have been constructed by widening the stencil, for instance in WENO methods

### 2013-04-08 docs/parco13/abstract.tex
Use the submitted abstract.tex file

@@
 \usepackage{xspace}
-\usepackage{graphicx}
 
@@
 
-\newcommand{\Fig}[1]{Figure \ref{fig:#1}}
+\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
 
@@
 or unstructured meshes
-\cite{%ddfv, ausm+,
-openfoam},
+\cite{openfoam},
 where higher order methods have been constructed by widening the stencil, for instance in WENO methods
@@
 white).}
-\label{cubed3}
+\label{fig:cubed3}
 \end{figure}
@@
 augmented where necessary with information about one layer of ghost leaves
-(see Figure~\ref{cubed3}).
+(see \Fig{cubed3}).
 
@@
 \section{Numerical results}
-Below, we provide two examples that demonstrate the \forestclaw code.
+
+\begin{figure}
+\begin{center}
+\plotbox{\includegraphics[width=.4\columnwidth]{quadtree}}
+\hspace{.05\columnwidth}
+\plotbox{\includegraphics[width=.4\columnwidth]{sphere_plot167}}
+\end{center}
+\caption{Numerical results for solving the advection equation.  Left:  Unit
+square (single quadtree) with a twisting velocity field.  Right: Spherical ball
+with a rotational velocity field constructed from two mapped quadtrees.  In
+both cases the concentration is color-coded with a sharp gradient shown in red.
+The adaptive mesh refinement follows the location of the gradient (the patches
+are not shown where they become too fine for display).  Here we use $m = 8$
+cells per \clawpack patch.}
+\label{fig:results2d}
+\end{figure}
+
+We provide two examples that demonstrate the \forestclaw code.
 The numerical results to date have been designed to verify that
@@
 flexible and robust.  Of particular importance was ensuring that all
-ghost cell ghost cell transfers (including averaging and
+ghost cell transfers (including averaging and
 interpolation) have been implemented correctly.
 The basic \clawpack algorithm and corresponding code have been
-thoroughly tested, and so needed no further verification here.
+thoroughly tested and need no further verification in our context.
 
 In both sets of numerical results, we solve a scalar advection
-equation, given by
-\begin{equation*}
-q_t + ({\bf u}\cdot \nabla) q = 0
-\end{equation*}
-\noindent
+equation,
+\begin{equation}
+q_t + ({\bf u}\cdot \nabla) q = 0 ,
+\end{equation}
 where the velocity field ${\bf u} = (u(\xi,\eta,t), v(\xi,\eta,t)$ is
@@
 patch on each leaf ($m=8$ in each case), and the minimum and maximum
-levels, which in turn fix minimum and maximum effective grid
+refinement levels, which in turn fix minimum and maximum effective grid
 resolutions.
@@
 In both examples, a patch is tagged for refinement if the
-difference between its maximum and minimum values exceed a prescribed
-threshold.  A family of patches (4 sibling patches in 2d) is coarsened
-if a coarsened parent patch did not meet the criteria for refinement.
+difference between its maximum and minimum values exceeds a prescribed
+threshold.  A family of $2^d$ patches is coarsened to a parent patch
+if this patch would not meet the criteria for refinement.
 
-\noindent
-{\bf Example 1 : } In this example, an initial concentration field $q$
+\begin{description}
+\item{\bf Example 1:} An initial concentration field $q$
 is set to 0 in the left half of a computational square and to 1 in the
@@
 the interface between the 0 and 1 concentration values.  The minimum
-and maximum levels of refinement were set to 3 and 6, respectively.
+and maximum levels of refinement are set to 3 and 6, respectively.
 This results in a minimum resolution of $64 \times 64$ and a maximum
@@
 
-\noindent {\bf Example 2 : } This example, we demonstrate the
+\item{\bf Example 2:} We demonstrate the
 multiblock functionality of \forestclaw by considering flow on a
@@
 defined in the computational space $[-1,1] \times [-1,1]$
-\cite{ca-he-le:2008, be-ca-he-le:2009}.  Each quadtree is mapped to cover one hemisphere
-of the sphere.  The initial conditions were set to 0 in one half of
-the sphere, and 1 in the other half, where the halves are not
+\cite{ca-he-le:2008, be-ca-he-le:2009}.  Each quadtree is mapped to cover one
+hemisphere.  The initial conditions is $q = 0$ in one half of
+the sphere, and $q = 1$ in the other half, where the halves are not
 necessarily aligned with the equator of the mapping.  The flow field
-${u}$ simulates rigid body rotation.  We show the results at an intermediate time in
-\Fig{results2d}.
-
-\begin{figure}
-\begin{center}
-\plotbox{\includegraphics[width=.4\columnwidth]{quadtree}} \hfil
-% \hspace{.05\columnwidth}
-\plotbox{\includegraphics[width=.4\columnwidth]{sphere_plot167}}
-\end{center}
-\caption{Numerical results for solving the advection equation.  Left:  Unit
-square (single quadtree) with a twisting velocity field.  Right: Spherical ball
-with a rotational velocity field constructed from two mapped quadtrees.  In
-both cases the concentration is color-coded with a sharp gradient shown in red.
-The adaptive mesh refinement follows the location of the gradient (the patches
-are not shown where they become too fine for display).  Here we use $m = 8$
-cells per \clawpack patch.}
-\label{fig:results2d}
-\end{figure}
+${u}$ simulates rigid body rotation.  We show the results at an intermediate
+time in \Fig{results2d}.
+\end{description}
 
 The results in this section are preliminary and will be extended in the final
-version.  We have so far exercised \forestclaw for two 2D manifolds, namely the
-unit square and a sphere composed of two mapped squares.  On both we solve the
-advection equation with a given velocity field.  We advect a concentration
-field with a sharp gradient which we track dynamically with adaptive mesh
-refinement.  We observe that the refinement follows the front and allows for an
-accurate representation of its location over time; see \Fig{results2d}.
+version.  We have so far exercised \forestclaw for two 2D manifolds,
+% namely the
+% unit square and a sphere composed of two mapped squares.
+% On both we solve the
+% advection equation with a given velocity field.
+advecting a concentration field with a sharp gradient that we track dynamically
+with adaptive mesh refinement.  We observe that the refinement follows the
+front and allows for an accurate representation of its location over time.
+% \Fig{results2d}.
 
@@
 \forestclaw was done. The leaf/patch paradigm was independently
-presented by B.\ as part of a talk at the SCI Institute, Utah, in July
-2011.
+presented by B.\ as part of a talk at the SCI Institute, Salt Lake City, Utah,
+USA in July 2011.
 

### 2013-06-07 docs/parco13/paper.tex
Add initial revisions and todo list to PARCO13

--- 2013-04-08 docs/parco13/abstract.tex
@@
 
+\newcommand{\alert}[1]{{\color{red}#1}}
+
 %%%%%%%%%%% End of definitions
@@
 \begin{abstract}
-%
-We present a new hybrid paradigm for parallel adaptive mesh refinement (AMR)
-that combines the scalability and lightweight architecture of tree-based AMR
-with the computational efficiency of patch-based solvers for hyperbolic
-conservation laws.  The key idea is to interpret each leaf of the AMR hierarchy
-as one uniform compute patch in $\sR^d$ of dimensions $m^d$, where $m$ is
-customarily between 8 and 32.  Thus, computation on each patch can be optimized
-for speed, while we inherit the flexibility of adaptive meshes.  In our work we
-choose to work with the \pforest AMR library since it allows us to compose the mesh
-from multiple mapped octrees and enable the cubed sphere and other
-nontrivial multiblock geometries.
-%
+We present a new hybrid paradigm for parallel adaptive mesh
+refinement (AMR) that combines the scalability and lightweight architecture of
+tree-based AMR with the computational efficiency of patch-based solvers for
+hyperbolic conservation laws.  The key idea is to interpret each leaf of the AMR
+hierarchy as one uniform compute patch in $\sR^d$ with $m^d$ degrees of freedom,
+where $m$ is customarily between 8 and 32.  Thus, computation on each patch can
+be optimized for speed, while we inherit the flexibility of adaptive meshes.  In
+our work we choose to work with the \pforest AMR library since it allows us to
+compose the mesh from multiple mapped octrees and enables the cubed sphere and
+other nontrivial multiblock geometries.
 \end{abstract}
@@
 freedom (DOF) at the element level, as has been done with high-order
-spectral element \cite{TufoFischer99}, low order continuous Galerkin via methods that accumulate many elements simultaneously \cite{knepleyterrel:2013},
+spectral element \cite{TufoFischer99}, low order continuous Galerkin methods that accumulate many elements simultaneously \cite{knepleyterrel:2013},
 %finite volume (element?)
@@
 or unstructured meshes
-\cite{openfoam},
-where higher order methods have been constructed by widening the stencil, for instance in WENO methods
+\cite{openfoam};
+higher order methods have also been constructed by widening the stencil, for instance in WENO methods
 \cite{Shu:2009bi}.
@@
 and proven scalability \cite{BursteddeWilcoxGhattas11}.  This approach
-identifies each octree leaf with a mesh element.  In our work, we go
-beyond the traditional high-order element and define each element to
-be a dense computational patch with $m^d$ DOFs.  In fact, this approach
+identifies each octree leaf with a mesh element.  \alert{The presented work goes beyond the traditional high-order element and defines each element to
+be a dense computational patch with $m^d$ DOFs.}  In fact, this approach
 resembles block-structured AMR \cite{be-ol:1984, be-co:1989,
@@
 \cite{be-le:1991, amrclaw, Berger:2011du}.
+\comment{Review 1 points out that the statement in red may not be what we are intending to say here.
+Mainly the distinction between claiming better convergence then a high-order method (which I think is not what we are saying) and just stating that our elements are patches themselves as opposed to a single DOF.}
 
@@
 the interaction with neighbor patches via the use of a layer of ghost
-cells surrounding each patch.  To this end, we define an interface
+cells surrounding each patch.  \alert{To this end}, we define an interface
 that allows read-only access to the sequence of blocks,
@@
 Suitably informed by \pforest, \forestclaw stores only the patches local to
-each MPI rank.  The exchange of ghost data and parallel repartitioning is
+each MPI rank.  \alert{The exchange of ghost} data and parallel repartitioning is
 provided transparently by \pforest.  Mesh modification directives, such as
@@
 \pforest.
+
+\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.}
 
@@
 
-\forestclaw uses iterators over all rank-local leafs, optionally restricted to
+\forestclaw uses iterators over all rank-local leaves, optionally restricted to
 a given level.  Random access is possible and used when executing $O(1)$ time
@@
 forest
+\alert{Review 2 - Explain the preceding paragraph in more detail}
 % \comment{tree?}
@@
 points.  There is one such parallel data exchange per time step for a global
-value of $\dt$, or one exchange per discretization level per time step if $\dt$
+value of $\dt$, the time step length, or one exchange per discretization level per time step if $\dt$
 is chosen per-patch depending on its size (this is sometimes called
@@
 by \forestclaw alone without the need to involve \pforest.
-Additionally many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for the hyperbolic integration on a leaf-patch.
-The design of leaf-patches allows for the smart management of data local to accelerator and the host.
+Additionally many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for the integration of the hyperbolic system on a leaf-patch thereby allowing for hybrid parallelism.
+The design of leaf-patches then can enable efficient management of data local to many-core architectures (e.g.
+the Xeon Phi), and the host.
 

### 2013-07-29 applications/manyclaw/2d/simple/simple_user.cpp
Remove pointer bug in manyclaw code.

@@ -122,7 +122,7 @@ void simple_patch_setup(fclaw2d_domain_t *domain,
     amr_manyclaw_setaux(domain,this_patch,this_block_idx,this_patch_idx);
     advection_rp_aux_global_t* aux_global = new advection_rp_aux_global_t;
     aux_global->u[0] = 1.0; aux_global->u[1] = 1.0;
-    manyclaw_set_riemann_solvers(this_patch,advection_rp_grid_eval_serial, &aux_global,
+    manyclaw_set_riemann_solvers(this_patch,advection_rp_grid_eval_serial, aux_global,
                                  updater_first_order_dimensional_splitting);
 }
 

### 2013-07-29 src/solvers/amr_manyclaw/amr_manyclaw.cpp
Remove pointer bug in manyclaw code.

@@ -571,7 +571,6 @@ static
 void amr_manyclaw_patch_data_delete(void **wp)
 {
     amr_manyclaw_patch_data_t *manyclaw_patch_data = (amr_manyclaw_patch_data_t*) *wp;
-    delete manyclaw_patch_data->aux_global;
     delete manyclaw_patch_data;
 
     // or?

### 2013-07-29 src/amr_regrid.cpp
Move a code line to the proper place

@@ -226,9 +226,6 @@ void regrid(fclaw2d_domain_t **domain)
     fclaw2d_domain_iterate_families(*domain, cb_tag4coarsening,
                                     (void*) NULL);
 
-    // Domain data may go out of scope now.
-    ddata = NULL;
-
     // Then refine.
     fclaw2d_domain_iterate_patches(*domain, cb_tag4refinement,
                                    (void *) NULL);
@@ -237,6 +234,9 @@ void regrid(fclaw2d_domain_t **domain)
     // Will return be NULL if no refining was done
     fclaw2d_domain_t *new_domain = fclaw2d_domain_adapt(*domain);
     fclaw_bool have_new_refinement = new_domain != NULL;
+
+    // Domain data may go out of scope now.
+    ddata = NULL;
 
     if (have_new_refinement)
     {

### 2013-07-29 src/amr_exchangecoarse.cpp
Two possible comment typos

@@ -674,7 +674,7 @@ void exchange_with_coarse(fclaw2d_domain_t *domain,
     fclaw2d_domain_iterate_level(domain,coarser_level,cb_face_interpolate,
                                  (void *) &e_info);
 
-    /* Interpolate from coarse grids to finer grids. */
+    /* Interpolate from coarse grids to finer grids. TODO: comment correct?? */
     e_info.is_coarse = fclaw_false;
     e_info.is_fine = fclaw_true;
     fclaw2d_domain_iterate_level(domain,finer_level,cb_face_interpolate,
@@ -686,7 +686,7 @@ void exchange_with_coarse(fclaw2d_domain_t *domain,
     fclaw2d_domain_iterate_level(domain,coarser_level, cb_corner_interpolate,
                                  (void *) &e_info);
 
-    /* Interpolate coarse grid to fine grid ghost cells. */
+    /* Interpolate coarse grid to fine grid ghost cells. TODO: correct?? */
     e_info.is_coarse = fclaw_false;
     e_info.is_fine = fclaw_true;
     fclaw2d_domain_iterate_level(domain,finer_level, cb_corner_interpolate,

### 2013-07-29 src/amr_regrid.cpp
Count number of exchange and regrid operations

@@ -276,4 +276,7 @@ void regrid(fclaw2d_domain_t **domain)
     // Stop timer
     ddata = get_domain_data(*domain);
     fclaw2d_timer_stop (&ddata->timers[FCLAW2D_TIMER_REGRID]);
+
+    // Count calls to this function
+    ++ddata->count_amr_regrid;
 }

### 2013-07-29 src/amrreset.cpp

@@ -98,15 +98,21 @@ void amrreset(fclaw2d_domain_t **domain)
         sc_stats_compute ((*domain)->mpicomm, FCLAW2D_TIMER_COUNT, stats);
         sc_stats_print (sc_package_id, SC_LP_PRODUCTION, FCLAW2D_TIMER_COUNT,
                         stats, 1, 0);
-        SC_GLOBAL_PRODUCTIONF ("Procs %d advance %d %g exchange %g regrid %g\n",
-                               (*domain)->mpisize, ddata->count_amr_advance,
+        SC_GLOBAL_PRODUCTIONF ("Procs %d advance %d %g exchange %d %g "
+                               "regrid %d %g\n", (*domain)->mpisize,
+                               ddata->count_amr_advance,
                                stats[FCLAW2D_TIMER_ADVANCE].average,
+                               ddata->count_ghost_exchange,
                                stats[FCLAW2D_TIMER_EXCHANGE].average,
+                               ddata->count_amr_regrid,
                                stats[FCLAW2D_TIMER_REGRID].average);
-        SC_GLOBAL_PRODUCTIONF ("Max/P %d advance %d %g exchange %g regrid %g\n",
-                               (*domain)->mpisize, ddata->count_amr_advance,
+        SC_GLOBAL_PRODUCTIONF ("Max/P %d advance %d %g exchange %d %g "
+                               "regrid %d %g\n", (*domain)->mpisize,
+                               ddata->count_amr_advance,
                                stats[FCLAW2D_TIMER_ADVANCE].max,
+                               ddata->count_ghost_exchange,
                                stats[FCLAW2D_TIMER_EXCHANGE].max,
+                               ddata->count_amr_regrid,
                                stats[FCLAW2D_TIMER_REGRID].max);
     }
 

### 2013-07-29 src/fclaw2d_typedefs.h

@@ -68,7 +68,7 @@ typedef struct fclaw2d_domain_data
 {
     /* Debug counters and timers */
     int count_set_clawpatch, count_delete_clawpatch;
-    int count_amr_advance;
+    int count_amr_advance, count_ghost_exchange, count_amr_regrid;
     int is_latest_domain;
     fclaw2d_timer_t timers[FCLAW2D_TIMER_COUNT];
 

### 2013-07-29 src/amr_repartition_domain.cpp
Count number of exchange and regrid operations

@@ -125,6 +125,7 @@ static
 /* This is called anytime we need to update ghost patch data */
 void exchange_ghost_patch_data(fclaw2d_domain_t* domain, fclaw_bool time_interp)
 {
+    fclaw2d_domain_data_t *ddata = get_domain_data (domain);
     fclaw2d_domain_exchange_t *e = get_domain_exchange_data(domain);
 
     /* Store pointers to local boundary data.  We do this here
@@ -152,6 +153,9 @@ void exchange_ghost_patch_data(fclaw2d_domain_t* domain, fclaw_bool time_interp)
     /* Store newly updated e->ghost_patch_data into ghost patches constructed
        locally */
     unpack_ghost_patches(domain,e, time_interp);
+
+    /* Count calls to this function */
+    ++ddata->count_ghost_exchange;
 }
 
 

### 2013-07-29 src/amr_utils.cpp

@@ -57,6 +57,8 @@ void init_domain_data(fclaw2d_domain_t *domain)
 
     ddata->count_set_clawpatch = ddata->count_delete_clawpatch = 0;
     ddata->count_amr_advance = 0;
+    ddata->count_ghost_exchange = 0;
+    ddata->count_amr_regrid = 0;
     ddata->is_latest_domain = 0;        /* set 1 by amrinit or rebuild_domain */
 
     ddata->domain_exchange = NULL;
@@ -211,6 +213,8 @@ void copy_domain_data(fclaw2d_domain_t *old_domain, fclaw2d_domain_t *new_domain
             sizeof (fclaw2d_timer_t) * FCLAW2D_TIMER_COUNT);
     ddata_new->is_latest_domain = 1;
     ddata_new->count_amr_advance = ddata_old->count_amr_advance;
+    ddata_new->count_ghost_exchange = ddata_old->count_ghost_exchange;
+    ddata_new->count_amr_regrid = ddata_old->count_amr_regrid;
 
 
     /*

### 2013-07-30 docs/stampede/analyze.pl
Adapt analyze.pl to the new output convention

@@ -4,7 +4,7 @@ use strict;
 use Getopt::Std;
 
 my (%opt, $average, $startproc, $found, $filestr);
-my ($procs, $numad, $tad, $tex, $trg);
+my ($procs, $numad, $numex, $numrg, $tad, $tex, $trg);
 my (%proclist, %proclevels, %procad, %procex, %procrg);
 my ($level, $adarr, $exarr, $rgarr, $step);
 
@@ -29,23 +29,27 @@ while (<>) {
 
 	$found = 0;
 	if ($average) {
-if (m:Procs (\d+) advance (\d+) ([0-9.e+-]+) exchange ([0-9.e+-]+) regrid ([0-9.e+-]+):) {
+if (m:Procs (\d+) advance (\d+) ([0-9.e+-]+) exchange (\d+) ([0-9.e+-]+) regrid (\d+) ([0-9.e+-]+):) {
 		$found = 1;
 		$procs = $1;
 		$numad = $2;
 		$tad = $3;
-		$tex = $4;
-		$trg = $5;
+		$numex = $4;
+		$tex = $5;
+		$numrg = $6;
+		$trg = $7;
 }
 	}
 	else {
-if (m:Max/P (\d+) advance (\d+) ([0-9.e+-]+) exchange ([0-9.e+-]+) regrid ([0-9.e+-]+):) {
+if (m:Max/P (\d+) advance (\d+) ([0-9.e+-]+) exchange (\d+) ([0-9.e+-]+) regrid (\d+) ([0-9.e+-]+):) {
 		$found = 1;
 		$procs = $1;
 		$numad = $2;
 		$tad = $3;
-		$tex = $4;
-		$trg = $5;
+		$numex = $4;
+		$tex = $5;
+		$numrg = $6;
+		$trg = $7;
 }
 	}
 	if ($found) {

### 2013-07-30 src/amr_regrid.cpp
Start using global min/maxlevels in regrid

@@ -212,6 +212,8 @@ void regrid(fclaw2d_domain_t **domain)
     fclaw2d_domain_data_t* ddata = get_domain_data(*domain);
     fclaw2d_timer_start (&ddata->timers[FCLAW2D_TIMER_REGRID]);
 
+    /* These values are updated below after iterate_adapted. */
+    /* TODO: can we use global min/maxlevels here? */
     int minlevel = gparms->minlevel;
     int maxlevel = gparms->maxlevel;
 
@@ -250,6 +252,8 @@ void regrid(fclaw2d_domain_t **domain)
 
         // Do a level exchange to make sure all data is valid;
         // Coarse and fine grid exchanges will happen during time stepping as needed.
+        minlevel = new_domain->global_minlevel;
+        maxlevel = new_domain->global_maxlevel;
         for (int level = minlevel; level <= maxlevel; level++)
         {
             level_exchange(new_domain,level, FCLAW2D_TIMER_REGRID);

### 2013-07-30 docs/stampede/analyze.pl
Make analyze.pl recognize both conventions

@@ -27,6 +27,7 @@ else {
 
 while (<>) {
 
+	# Try the latest output convention first
 	$found = 0;
 	if ($average) {
 if (m:Procs (\d+) advance (\d+) ([0-9.e+-]+) exchange (\d+) ([0-9.e+-]+) regrid (\d+) ([0-9.e+-]+):) {
@@ -52,6 +53,32 @@ if (m:Max/P (\d+) advance (\d+) ([0-9.e+-]+) exchange (\d+) ([0-9.e+-]+) regrid 
 		$trg = $7;
 }
 	}
+
+	# If the latest convention didn't match, try the older one
+	if (!$found) {
+	if ($average) {
+if (m:Procs (\d+) advance (\d+) ([0-9.e+-]+) exchange ([0-9.e+-]+) regrid ([0-9.e+-]+):) {
+		$found = 1;
+		$procs = $1;
+		$numad = $2;
+		$tad = $3;
+		$tex = $4;
+		$trg = $5;
+}
+	}
+	else {
+if (m:Max/P (\d+) advance (\d+) ([0-9.e+-]+) exchange ([0-9.e+-]+) regrid ([0-9.e+-]+):) {
+		$found = 1;
+		$procs = $1;
+		$numad = $2;
+		$tad = $3;
+		$tex = $4;
+		$trg = $5;
+}
+	}
+	}
+
+	# Suppose we have identified relevant output
 	if ($found) {
 
 		if (!$proclist{$procs}) {

### 2013-07-30 docs/parco13/paper.tex
Rewrap to reduce the size of diff to abstract.tex
Copy some reviewer's comments

@@
 \begin{abstract}
-We present a new hybrid paradigm for parallel adaptive mesh
-refinement (AMR) that combines the scalability and lightweight architecture of
-tree-based AMR with the computational efficiency of patch-based solvers for
-hyperbolic conservation laws.  The key idea is to interpret each leaf of the AMR
-hierarchy as one uniform compute patch in $\sR^d$ with $m^d$ degrees of freedom,
-where $m$ is customarily between 8 and 32.  Thus, computation on each patch can
-be optimized for speed, while we inherit the flexibility of adaptive meshes.  In
-our work we choose to work with the \pforest AMR library since it allows us to
-compose the mesh from multiple mapped octrees and enables the cubed sphere and
-other nontrivial multiblock geometries.
+We present a new hybrid paradigm for parallel adaptive mesh refinement (AMR)
+that combines the scalability and lightweight architecture of tree-based AMR
+with the computational efficiency of patch-based solvers for hyperbolic
+conservation laws.  The key idea is to interpret each leaf of the AMR hierarchy
+as one uniform compute patch in $\sR^d$ with $m^d$ degrees of freedom, where $m$ is
+customarily between 8 and 32.  Thus, computation on each patch can be optimized
+for speed, while we inherit the flexibility of adaptive meshes.  In our work we
+choose to work with the \pforest AMR library since it allows us to compose the mesh
+from multiple mapped octrees and enables the cubed sphere and other
+nontrivial multiblock geometries.
 \end{abstract}
@@
 
-\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.}
+\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
+%
+Quote: How is AMR working in detail. For example: Figure 1 shows a
+2:1 refinement, thus refinement of a single patch could affect a whole region
+of the computational domain. + The underlying equations are known to
+Forestclaw: How is is the information shared with p4est (equation dependent
+refinement criteria, refinement and coarsening strategies (interpolation,
+integration).
+}
 

### 2013-07-30 docs/parco13/paper.tex
Address two concerns

@@
 \usepackage{graphicx}
+\usepackage{tikz}
 \usepackage{url}
@@
 and proven scalability \cite{BursteddeWilcoxGhattas11}.  This approach
-identifies each octree leaf with a mesh element.  \alert{The presented work goes beyond the traditional high-order element and defines each element to
-be a dense computational patch with $m^d$ DOFs.}  In fact, this approach
+identifies each octree leaf with a mesh element.  The presented work does not
+construct a traditional high-order element but defines each element to be a
+dense computational patch with $m^d$ DOFs.  In fact, this approach
 resembles block-structured AMR \cite{be-ol:1984, be-co:1989,
@@
 \cite{be-le:1991, amrclaw, Berger:2011du}.
-\comment{Review 1 points out that the statement in red may not be what we are intending to say here.
-Mainly the distinction between claiming better convergence then a high-order method (which I think is not what we are saying) and just stating that our elements are patches themselves as opposed to a single DOF.}
+% \comment{Review 1 points out that the statement in red may not be what we are
+% intending to say here.  Mainly the distinction between claiming better
+% convergence then a high-order method (which I think is not what we are saying)
+% and just stating that our elements are patches themselves as opposed to a
+% single DOF.}
+% CB: reworded the criticized statement above.
 
@@
 \hspace{.05\columnwidth}
+\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},
+                    x=1mm, y=1mm]
+  \node at (0,0) {
     \includegraphics[width=.4\columnwidth]{../figs_misc/cubed3}
+  };
+  \node [fill=white, inner sep=2pt] at (28, 10) {1};
+  \node [fill=white, inner sep=2pt] at (5, 27) {2};
+  \node [fill=white, inner sep=2pt] at (41, 27) {2};
+  \node [fill=white, inner sep=2pt] at (28, 36) {3};
+\end{tikzpicture}
+%\includegraphics[width=.4\columnwidth]{../figs_misc/cubed3}
 \end{center}
@@
 pattern has been 2:1 balanced in terms of neighbor sizes and partitioned
-between five MPI processes (the three middle ones are shown from black to
-white).}
+between five MPI processes (the three middle ones out of $0 \ldots 4$
+are shown from black to white).}
 \label{fig:cubed3}

### 2013-07-30 docs/parco13/paper.tex
Clarify refinement procedure

@@
 \newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}\xspace}
+\newcommand{\cO}{\mathcal{O}}
 \newcommand{\sR}{\mathbb{R}}
@@
 to be propagated to the numerical code in \forestclaw that implements
-the interaction with neighbor patches via the use of a layer of ghost
-cells surrounding each patch.  \alert{To this end}, we define an interface
-that allows read-only access to the sequence of blocks,
-the list of patches for each, and a lookup of neighbor patches (and
-their relative orientation which is nontrivial between blocks).
+the interaction with neighbor patches via the use of a layer of (typically two)
+ghost cells surrounding each patch.  To this end, we have designed an
+interface that provides read-only access to the sequence of blocks, the list of
+patches for each, and a constant-time lookup of neighbor patches (and their
+relative orientation which is nontrivial between blocks).
+% Since a neighbor patch has $m \ge 4$ cells in each direction, the required
+% ghost cells are available by construction.
 Suitably informed by \pforest, \forestclaw stores only the patches local to
-each MPI rank.  \alert{The exchange of ghost} data and parallel repartitioning is
-provided transparently by \pforest.  Mesh modification directives, such as
-adaptive refinement and coarsening, are called from \forestclaw and relayed to
-\pforest.
+each MPI rank.
 
-\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
-%
-Quote: How is AMR working in detail. For example: Figure 1 shows a
-2:1 refinement, thus refinement of a single patch could affect a whole region
-of the computational domain. + The underlying equations are known to
-Forestclaw: How is is the information shared with p4est (equation dependent
-refinement criteria, refinement and coarsening strategies (interpolation,
-integration).
-}
+Mesh modification directives, such as adaptive refinement and coarsening, are
+called from \forestclaw and executed as black-box operations inside \pforest.
+The only information that flows back to \pforest is a set of per-patch
+refinement and coarsening flags, which are computed from the numerical state in
+\forestclaw.  Our current algorithm requires neighbor patches to have a size
+difference of at most a factor of two (the 2:1 balance condition).  This condition
+can have nonlocal effects on the refinement pattern and is enforced by
+special-purpose parallel algorithms in \pforest \cite{IsaacBursteddeGhattas12}.
 
@@
 \forestclaw uses iterators over all rank-local leaves, optionally restricted to
-a given level.  Random access is possible and used when executing $O(1)$ time
+a given level.  Random access is possible and used when executing $\cO(1)$ time
 neighbor lookups.  Looping over the patches in the order prescribed by the
@@
 % since the ghost patches are likely closeby with respect to the $z$-order.
+
+\alert{The exchange of ghost} data and parallel repartitioning is
+provided transparently by \pforest.
+\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
+%
+Quote: How is AMR working in detail. For example: Figure 1 shows a
+2:1 refinement, thus refinement of a single patch could affect a whole region
+of the computational domain. + The underlying equations are known to
+Forestclaw: How is is the information shared with p4est (equation dependent
+refinement criteria, refinement and coarsening strategies (interpolation,
+integration).
+}
 
@@
 The threaded parallelism over the degrees of freedom of a patch can be handled
-by \forestclaw alone without the need to involve \pforest.
-Additionally many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for the integration of the hyperbolic system on a leaf-patch thereby allowing for hybrid parallelism.
-The design of leaf-patches then can enable efficient management of data local to many-core architectures (e.g.
-the Xeon Phi), and the host.
+by \forestclaw alone without the need to involve \pforest.  Additionally,
+many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for
+the integration of the hyperbolic system on a leaf-patch thereby allowing for
+hybrid parallelism.  The design of leaf-patches then can enable efficient
+management of data local to many-core architectures (for example the Xeon Phi
+coprocessor) and the host.
 

### 2013-07-30 docs/parco13/paper.tex
Work through section 3 in paper

@@
 called from \forestclaw and executed as black-box operations inside \pforest.
-The only information that flows back to \pforest is a set of per-patch
+The only information that flows into \pforest is a set of per-patch
 refinement and coarsening flags, which are computed from the numerical state in
@@
 
-\forestclaw uses iterators over all rank-local leaves, optionally restricted to
-a given level.  Random access is possible and used when executing $\cO(1)$ time
-neighbor lookups.  Looping over the patches in the order prescribed by the
-forest
-\alert{Review 2 - Explain the preceding paragraph in more detail}
+\forestclaw uses iterators over all rank-local leaves to execute numerical
+tasks one patch at a time, optionally restricted to a given refinement level.
+Random access to patches is possible and used when dereferencing the results
+from
+% executing
+% $\cO(1)$ time
+neighbor lookups.  Looping over the patches in the order
+prescribed by the forest
+% \alert{Review 2 - Explain the preceding paragraph in more detail}
 % \comment{tree?}
@@
 
-\alert{The exchange of ghost} data and parallel repartitioning is
-provided transparently by \pforest.
-\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
-%
-Quote: How is AMR working in detail. For example: Figure 1 shows a
-2:1 refinement, thus refinement of a single patch could affect a whole region
-of the computational domain. + The underlying equations are known to
-Forestclaw: How is is the information shared with p4est (equation dependent
-refinement criteria, refinement and coarsening strategies (interpolation,
-integration).
-}
-
 When \forestclaw accesses neighbor patches, they can be on the same or a
@@
 local view of the distributed mesh and data which relieves the developer from
-programming to the MPI interface.  Parallel neighbor exchanges are hidden
-inside \pforest and called by \forestclaw at well-defined synchronization
-points.  There is one such parallel data exchange per time step for a global
-value of $\dt$, the time step length, or one exchange per discretization level per time step if $\dt$
-is chosen per-patch depending on its size (this is sometimes called
-sub-cycling).  When using sub-cycling, the load balance is attained per level as
-it is done for example in Chombo \cite{ColellaGravesKeenEtAl07}
-or recent geometric adaptive multigrid schemes
+programming to the MPI interface.
+% Parallel neighbor exchanges are hidden
+% inside \pforest and called by \forestclaw
+% % at well-defined synchronization points
+% as follows.
+
+With MPI parallelism, neighbor patches can be either local or assigned to a
+different process (ghost patches).  Since this must make no difference
+numerically, we need to ensure that the values of ghost patches are current
+whenever they are accessed.  To this end, we allocate storage for a layer of
+ghost patches in \forestclaw, which we can pass to a general-purpose \pforest
+routine that communicates local leaf data to all processes that view a
+particular leaf as a ghost.  If we call this routine before we go through the
+local neighbor interactions, we can handle ghost values implicitly by neighbor
+lookups without querying if they are local or remote.
+
+% \alert{The exchange of ghost} data and parallel repartitioning is
+% provided transparently by \pforest.
+% \alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
+% %
+% Quote: How is AMR working in detail. For example: Figure 1 shows a
+% 2:1 refinement, thus refinement of a single patch could affect a whole region
+% of the computational domain. + The underlying equations are known to
+% Forestclaw: How is is the information shared with p4est (equation dependent
+% refinement criteria, refinement and coarsening strategies (interpolation,
+% integration).
+% }
+% CB: Is this addressed sufficiently now?
+
+% count of exchanges with and without sub-cycling
+In the context of finite-element or finite-difference methods, there should be
+one such parallel data exchange per time step for a global value of the time
+step length $\dt$, or one exchange per discretization level per time step if
+$\dt$ is chosen per-patch depending on its size (this is sometimes called
+sub-cycling).  In the tradition of block-structured AMR codes, interaction
+between neighbor patches is done in a hierarchy from coarse to fine levels, and
+then correction factors are propagated back from fine to coarse levels, requiring
+an exchange at each level.  With sub-cycling, this entails a recursion with an
+operation count that is exponential in the difference between the largest and
+the smallest refinement level, with the benefit that $\dt$ matches the CFL
+condition at each level.  While a large number of exchanges per time step can
+present a scaling bottleneck due to the inherent synchronization and latency
+losses, the expectation is still that the time to solution improves when
+switching from uniform to adaptive meshes, since we are computing with fewer
+patches, and improves again when enabling sub-cycling, since we take larger
+time steps on the coarser levels.
+
+% load balance / (weighted) partition
+The space-filling curve paradigm allows for lightweight repartitioning
+algorithms.  Even uniform meshes benefit from this approach since the number of
+processors does not need to be commensurable with the number of patches in each
+space dimension.  For adaptive meshes, we repartiton after every refinement
+operation and transfer the numerical data accordingly; see \cite[Fig.\
+4]{BursteddeGhattasStadlerEtAl08} for the overall procedure.  When using
+sub-cycling, we have the option to assign a weight to each patch based on its
+level, determined by the expected number of sub-cycles per coarse-level time
+step, and partition to equidistribute the cumulative weight between the
+processors.  Currently, we are not enforcing a load balance that is attained
+separately for each level, as it is done for example in Chombo
+\cite{ColellaGravesKeenEtAl07} or recent geometric adaptive multigrid schemes
 \cite{SundarBirosBursteddeEtAl12}.
@@
 ghost cell transfers (including averaging and
-interpolation) have been implemented correctly.
-The basic \clawpack algorithm and corresponding code have been
+interpolation) are implemented correctly.
+The basic \clawpack algorithm and corresponding code are
 thoroughly tested and need no further verification in our context.

### 2013-07-30 src/amr_exchangecoarse.cpp
Finally got logic for the parallel ghost patch exchange for the sphere working

@@ -189,7 +189,42 @@ void cb_corner_average(fclaw2d_domain_t *domain,
                     }
                     else
                     {
-                        int icorner_coarse = icorner;
+                        int icorner_coarse;
+                        if (is_block_corner)
+                        {
+                            icorner_coarse = icorner;
+                        }
+                        else
+                        {
+                            if (icorner == 0)
+                            {
+                                if (intersects_block[0])
+                                    icorner_coarse = 2;
+                                else if (intersects_block[2])
+                                    icorner_coarse = 1;
+                            }
+                            else if (icorner == 1)
+                            {
+                                if (intersects_block[1])
+                                    icorner_coarse = 3;
+                                else if (intersects_block[2])
+                                    icorner_coarse = 0;
+                            }
+                            else if (icorner == 2)
+                            {
+                                if (intersects_block[0])
+                                    icorner_coarse = 0;
+                                else if (intersects_block[3])
+                                    icorner_coarse = 3;
+                            }
+                            else if (icorner == 3)
+                            {
+                                if (intersects_block[1])
+                                    icorner_coarse = 1;
+                                else if (intersects_block[3])
+                                    icorner_coarse = 2;
+                            }
+                        }
                         coarse_cp->mb_average_corner_ghost(icorner_coarse,
                                                            refratio,corner_cp,time_interp,
                                                            is_block_corner,intersects_block);
@@ -296,11 +331,11 @@ void cb_corner_interpolate(fclaw2d_domain_t *domain,
             }
             else if (ref_flag == -1 && is_fine)
             {
-                /* Corner is at at coarser level.  If the corner patch is a parallel
+                /* Corner is at a coarser level.  If the corner patch is a parallel
                    ghost patch, then we should average onto it as well (even if it
                    is technically read only)  Otherwise, it won't have good data for
                    doing interpolation to this fine grid later */
-                if (fclaw2d_patch_is_ghost (neighbor_patch))
+                if (fclaw2d_patch_is_ghost(neighbor_patch))
                 {
                     /* Corner is a parallel ghost patch.  Do the same as above, but
                        but from the perspective of the ghost (coarser) patch.*/
@@ -309,7 +344,6 @@ void cb_corner_interpolate(fclaw2d_domain_t *domain,
 
                     if (this_block_idx == corner_block_idx)
                     {
-                        /* Works only for the sphere grid */
                         int icorner_coarse = 3-icorner;
[…]

### 2013-07-30 applications/advection/2d/sphere/ghost_twopatch.f
Added comments clarifying the corner exchange at block boundaries

@@ -412,9 +412,11 @@ c                 # Average fine grid corners onto coarse grid ghost corners
                         ifine = (ibc-1)*refratio + ii
                         jfine = (jbc-1)*refratio + jj
                         if (is_block_bdry(0) .eq. 1) then
+c                          # get info from fine block corner 2
                            kf = areafine(ifine,my+1-jfine)
                            qf =   qfine(ifine,my+1-jfine,mq)
                         elseif (is_block_bdry(2) .eq. 1) then
+c                          # get info from fine block corner 1
                            kf = areafine(mx+1-ifine,jfine)
                            qf =   qfine(mx+1-ifine,jfine,mq)
                         endif
@@ -435,9 +437,11 @@ c                 # Average fine grid corners onto coarse grid ghost corners
                         ifine = (ibc-1)*refratio + ii
                         jfine = (jbc-1)*refratio + jj
                         if (is_block_bdry(1) .eq. 1) then
+c                          # get info from fine block corner 3
                            kf = areafine(mx+1-ifine,my+1-jfine)
                            qf =   qfine(mx+1-ifine,my+1-jfine,mq)
                         elseif (is_block_bdry(2) .eq. 1) then
+c                          # get info from fine block corner 0
                            kf = areafine(ifine,jfine)
                            qf =   qfine(ifine,jfine,mq)
                         endif
@@ -458,9 +462,11 @@ c                 # Average fine grid corners onto coarse grid ghost corners
                         ifine = (ibc-1)*refratio + ii
                         jfine = (jbc-1)*refratio + jj
                         if (is_block_bdry(0) .eq. 1) then
+c                          # get info from fine block corner 0
                            kf = areafine(ifine,jfine)
                            qf =   qfine(ifine,jfine,mq)
                         elseif (is_block_bdry(3) .eq. 1) then
+c                          # get info from fine block corner 3
                            kf = areafine(mx+1-ifine,my+1-jfine)
                            qf =   qfine(mx+1-ifine,my+1-jfine,mq)
                         endif
@@ -481,9 +487,11 @@ c                 # Average fine grid corners onto coarse grid ghost corners
                         ifine = (ibc-1)*refratio + ii
                         jfine = (jbc-1)*refratio + jj
                         if (is_block_bdry(1) .eq. 1) then
+c                          # get info from fine block corner 1
                            kf = areafine(mx+1-ifine,jfine)
                            qf =   qfine(mx+1-ifine,jfine,mq)
[…]

### 2013-07-30 applications/advection/2d/sphere/afterframe.m
Add patch boundary properties - thicker line for plotting

@@ -24,6 +24,10 @@ if (mq == 1)
   set(o,'ylim',[qmin+1 qmax+2]);
   set(o,'ticklength',[0 0])
   set(o,'fontsize',16,'fontweight','bold')
+
+  setpatchborderprops(1:10,'linewidth',2);
+  showpatchborders;
+
 else
   cm_index = false;
 

### 2013-07-30 docs/parco13/paper.tex

@@
 
+\alert{Thought: Should we swap sections 3. and 4.?}
+
 For hyperbolic problems, we integrate the solution on a single uniform patch,
@@
 numerical parameters that are set in each case include the size of the
-patch on each leaf ($m=8$ in each case), and the minimum and maximum
+patch on each leaf ($m=8$ or $m=16$), and the minimum and maximum
 refinement levels, which in turn fix minimum and maximum effective grid

### 2013-07-30 docs/parco13/paper.tex
Start with strong scaling for uniform mesh

@@
 
-Grid refinement and coarsening requires interpolation from coarse grids
+Mesh refinement and coarsening requires interpolation from coarse grids
 to newly created fine grids, and the averaging of fine grid data to a newly
@@
 patch on each leaf ($m=8$ or $m=16$), and the minimum and maximum
-refinement levels, which in turn fix minimum and maximum effective grid
+refinement levels, which in turn fix minimum and maximum effective mesh
 resolutions.
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_advance_strong-crop}
+  \\[1ex]
+  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_exchange_strong-crop}
+\end{center}
+\caption{Strong scaling of MPI parallelism for advection on a uniform mesh.
+  Top: run time of time integration.  Bottom: time spent in local and parallel
+  neighbor exchanges.  The number of patches is $2^{2 \ell}$ at level $\ell$.
+  The flat lines for the smaller runs are caused by having
+  less patches than MPI processes, leaving some of them idle.  For each data point
+  we use the maximum time required across all processes.}
+\label{fig:uniscale}
+\end{figure}%
+We begin our analysis of Example 1 with uniform-mesh experiments run on TACC's
+\emph{Stampede} supercomputer at different levels of resolution.  First of all,
+we use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
+scaling efficiency for the time integration between 97\%--103\% depending on
+the resolution (from levels 1 and 8).  Then, we examine processor counts from
+16 to 256 using 16 cores per node.  The scaling behavior remains linear for
+the time integration itself, and mostly linear for the local and parallel
+neighbor exchanges which require about 10\% of the time integration run time;
+see Figure~\ref{fig:uniscale}.  Due to space constraints we omit the
+corresponding weak scaling results, which are nearly optimal as well.
+
 The results in this section are preliminary and will be extended in the final
@@
 USA in July 2011.
+\alert{Add TACC Stampede grants}
 

### 2013-07-30 docs/parco13/paper.tex
More on the paper

@@
 \newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
+\newcommand{\Tab}[1]{Table~\ref{tab:#1}}
 
@@
 neighbor exchanges which require about 10\% of the time integration run time;
-see Figure~\ref{fig:uniscale}.  Due to space constraints we omit the
+see \Fig{uniscale}.  Due to space constraints we omit the
 corresponding weak scaling results, which are nearly optimal as well.
 
-The results in this section are preliminary and will be extended in the final
-version.  We have so far exercised \forestclaw for two 2D manifolds,
-% namely the
-% unit square and a sphere composed of two mapped squares.
-% On both we solve the
-% advection equation with a given velocity field.
-advecting a concentration field with a sharp gradient that we track dynamically
-with adaptive mesh refinement.  We observe that the refinement follows the
-front and allows for an accurate representation of its location over time.
-% \Fig{results2d}.
+\begin{table}
+  \begin{center}
+    \begin{tabular}{cccc|rr}
+      \multicolumn{4}{c|}{strategy} & \multicolumn{2}{c}{wall clock time} \\
+      mesh & remesh & partition & time step & $P = 16$ & $P = 256$ \\
+      \hline
+      uniform & none       & by count  & global   & 3961.  & 256.  \\
+      \hline
+          AMR & every step & by count  & global   &  252.  &  54.6 \\
+          AMR & every 4    & by count  & global   &  178.  &  39.7 \\
+      \hline
+          AMR & every step & by count  & subcycle &   99.9 &  17.3 \\
+          AMR & every 4    & by count  & subcycle &   87.2 &  14.0 \\
+      \hline
+          AMR & every step & by weight & subcycle &   95.7 &  18.2 \\
+          AMR & every 4    & by weight & subcycle &   84.4 & 14.2 \\
+      \hline
+    \end{tabular}
+  \end{center}
+\caption{Comparison of different meshing strategies for the advection example
+  with a fixed maximum level 8, run on 16 and 256 cores, respectively.  The
+  wall clock time includes the whole run of the program beginning to end (file
+  I/O disabled).  The adaptive runs converge to a minimum level of 3.}
+\label{tab:swirlwall}
+\end{table}
+Next, we examine how we can reduce the overall wall clock time of a simulation
+by switching from uniform refinement to AMR with the same maximum level,
+effectively coarsening where high resolution is not beneficial.  For an example
+with a sharp front as depicted in \Fig{results2d} we expect considerable
+savings by AMR, which we confirm in \Tab{swirlwall}.  The wall clock times can
+be reduced by up to a factor of 50 by combining AMR with subcycling.  The
+adaptive runs are still very well load balanced overall, which we infer from
+the linear scaling of the time integration alone (not shown) and from the fact
+that the weighted partition does not influence the run times.  There is the
+tradeoff however, at least in the current implementation, that the adaptive
+runs suffer from the loss in scalability associated with the neighbor exchanges
+that we discussed earlier.
 
-\pforest is parallelized with MPI and well tested on large supercomputers, both
-in 2D and 3D.  Parallelization of \forestclaw is not completed yet but
-envisioned for the final version of the paper.
+% The results in this section are preliminary and will be extended in the final
+% version.  We have so far exercised \forestclaw for two 2D manifolds,
+% % namely the
+% % unit square and a sphere composed of two mapped squares.
+% % On both we solve the
+% % advection equation with a given velocity field.
+% advecting a concentration field with a sharp gradient that we track dynamically
+% with adaptive mesh refinement.  We observe that the refinement follows the
+% front and allows for an accurate representation of its location over time.
+% % \Fig{results2d}.
+% 
+% \pforest is parallelized with MPI and well tested on large supercomputers, both
+% in 2D and 3D.  Parallelization of \forestclaw is not completed yet but
+% envisioned for the final version of the paper.
 

### 2013-07-30 docs/parco13/paper.tex

@@
 
+\begin{figure}
+  \begin{center}
+  \begin{minipage}{.4\columnwidth}
+    \begin{tabular}{cc}
+      timing table for sphere & 123, \\
+    \end{tabular}
+  \end{minipage}
+  \hspace{.05\columnwidth}
+  \includegraphics[width=.4\columnwidth]{../figs_misc/sphere_partition5}
+\end{center}
+\caption{Partition of the spherical Example 2 between five MPI processors, indicated
+  by the color scale, at simulated time $T = 1$.  The partition shows the mesh equator,
+  while the refinement is invariant under the number of processors and the partition
+and follows the front, which is not aligned with the equator.}
+\label{fig:spherical}
+\end{figure}
+To demonstrate a multiblock connectivity, we include results for Example 2 in
+\Fig{spherical}.
+
 % The results in this section are preliminary and will be extended in the final

### 2013-07-31 applications/manyclaw/2d/swirl/swirl_user.cpp
Adding routines for swirl example.

@@ -109,7 +109,7 @@ void swirl_patch_setup(fclaw2d_domain_t *domain,
     /* This is called once when a new patch is created. */
     manyclaw_set_solver(domain,this_patch,this_block_idx,this_patch_idx);
     amr_manyclaw_setaux(domain,this_patch,this_block_idx,this_patch_idx);
-    manyclaw_set_riemann_solvers(this_patch,advection_rp_grid_eval_serial,
+    manyclaw_set_riemann_solvers(this_patch,advection_var_rp_grid_eval_serial, NULL,
                                  updater_first_order_dimensional_splitting);
 }
 

### 2013-07-31 src/amr_utils.cpp
Use mpi_init_threads.

@@ -365,7 +365,12 @@ fclaw_mpi_init (int * argc, char *** argv, MPI_Comm mpicomm, int lp)
 {
     int mpiret;
 
-    mpiret = MPI_Init (argc, argv);
+    //mpiret = MPI_Init (argc, argv);
+    //SC_CHECK_MPI (mpiret);
+
+    int provided;
+    mpiret = MPI_Init_thread (argc, argv, MPI_THREAD_FUNNELED, &provided);
+    if (provided != MPI_THREAD_FUNNELED) printf("Recieved mpi_init_thread level %d\n", provided);
     SC_CHECK_MPI (mpiret);
 
     sc_init (mpicomm, 0, 0, NULL, lp);

### 2013-07-31 applications/manyclaw/2d/swirl/swirl_user.cpp
Use omp version of advection kernel

@@ -109,7 +109,7 @@ void swirl_patch_setup(fclaw2d_domain_t *domain,
     /* This is called once when a new patch is created. */
     manyclaw_set_solver(domain,this_patch,this_block_idx,this_patch_idx);
     amr_manyclaw_setaux(domain,this_patch,this_block_idx,this_patch_idx);
-    manyclaw_set_riemann_solvers(this_patch,advection_var_rp_grid_eval_serial, NULL,
+    manyclaw_set_riemann_solvers(this_patch,advection_var_rp_grid_eval_void_omp, NULL,
                                  updater_first_order_dimensional_splitting);
 }
 

### 2013-07-31 docs/parco13/paper.tex
Comment on latest weak scaling results

@@
 neighbor exchanges which require about 10\% of the time integration run time;
-see \Fig{uniscale}.  Due to space constraints we omit the
-corresponding weak scaling results, which are nearly optimal as well.
+see \Fig{uniscale}.
+
+We have examined weak scalability as well by using four times as many processes
+for each level inclease, from level 6 at 64 cores to level 10 at 4096.  The
+exchange times stay consistently below 16\% of the time integration.  The time
+in seconds per time integration step varies between .0578 and .0602,
+which yields a parallel efficiency of 96\%.
 

### 2013-07-31 docs/parco13/paper.tex
Comment on Stampede allocation

@@
 We begin our analysis of Example 1 with uniform-mesh experiments run on TACC's
-\emph{Stampede} supercomputer at different levels of resolution.  First of all,
+Stampede supercomputer at different levels of resolution.  First of all,
 we use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
@@
 
+We would like to thank the Texas Advanced Computing Center (TACC) for access to
+the Stampede supercomputer under allocation TG-DPP130002.
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha
@@
 USA in July 2011.
-\alert{Add TACC Stampede grants}
 

### 2013-07-31 src/amr_regrid.H
Bump p4est/wrap to enable level-wise exchange

@@ -113,6 +113,8 @@ void rebuild_domain(fclaw2d_domain_t* old_domain, fclaw2d_domain_t* new_domain);
  */
 void repartition_domain(fclaw2d_domain_t** domain, int mode);
 void exchange_ghost_patch_data(fclaw2d_domain_t* domain,fclaw_bool time_interp);
+void exchange_ghost_patch_data_levels(fclaw2d_domain_t* domain,fclaw_bool time_interp,
+                                      int exchange_minlevel, int exchange_maxlevel);
 void delete_ghost_patches(fclaw2d_domain_t *domain);
 
 

### 2013-07-31 src/amr_repartition_domain.cpp

@@ -60,24 +60,29 @@ void delete_ghost_patches(fclaw2d_domain_t* domain)
 }
 
 
-static
-    void unpack_ghost_patches(fclaw2d_domain_t* domain, fclaw2d_domain_exchange_t *e,
-                              fclaw_bool time_interp)
+static void
+unpack_ghost_patches(fclaw2d_domain_t* domain, fclaw2d_domain_exchange_t *e,
+                     fclaw_bool time_interp,
+                     int exchange_minlevel, int exchange_maxlevel)
 {
     for(int i = 0; i < domain->num_ghost_patches; i++)
     {
         fclaw2d_patch_t* ghost_patch = &domain->ghost_patches[i];
-        int blockno = ghost_patch->u.blockno;
+        if (exchange_minlevel <= ghost_patch->level &&
+            ghost_patch->level <= exchange_maxlevel)
+        {
+            int blockno = ghost_patch->u.blockno;
 
-        /* not clear how useful this patchno is.  In any case, it isn't
-           used in defining the ClawPatch, so probably doesn't
-           need to be passed in */
-        int patchno = i;
+            /* not clear how useful this patchno is.  In any case, it isn't
+               used in defining the ClawPatch, so probably doesn't
+               need to be passed in */
+            int patchno = i;
 
-        /* access data stored on remote procs.  This might be time interpolated data. */
-        double *q = (double*) e->ghost_data[patchno];
+            /* access data stored on remote procs.  This might be time interpolated data. */
+            double *q = (double*) e->ghost_data[patchno];
 
-        unpack_clawpatch(domain, ghost_patch,blockno, patchno, q, time_interp);
+            unpack_clawpatch(domain, ghost_patch,blockno, patchno, q, time_interp);
+        }
     }
 }
 
@@ -125,6 +130,14 @@ static
 /* This is called anytime we need to update ghost patch data */
 void exchange_ghost_patch_data(fclaw2d_domain_t* domain, fclaw_bool time_interp)
 {
+    exchange_ghost_patch_data_levels (domain, time_interp,
+                                      domain->global_minlevel, domain->global_maxlevel);
+}
+
+/* This is called anytime we need to update ghost patch data for certain levels */
+void exchange_ghost_patch_data_levels(fclaw2d_domain_t* domain, fclaw_bool time_interp,
+                                      int exchange_minlevel, int exchange_maxlevel)
+{
[…]

### 2013-07-31 src/forestclaw2d.c

@@ -802,7 +802,8 @@ fclaw2d_domain_allocate_before_exchange (fclaw2d_domain_t * domain,
 
 void
 fclaw2d_domain_ghost_exchange (fclaw2d_domain_t * domain,
-                               fclaw2d_domain_exchange_t * e)
+                               fclaw2d_domain_exchange_t * e,
+                               int exchange_minlevel, int exchange_maxlevel)
 {
     p4est_wrap_t *wrap = (p4est_wrap_t *) domain->pp;
     p4est_ghost_t *ghost = wrap->match_aux ? wrap->ghost_aux : wrap->ghost;
@@ -815,9 +816,21 @@ fclaw2d_domain_ghost_exchange (fclaw2d_domain_t * domain,
 
     P4EST_ASSERT (e->num_exchange_patches == (int) ghost->mirrors.elem_count);
     P4EST_ASSERT (e->num_ghost_patches == (int) ghost->ghosts.elem_count);
-    p4est_ghost_exchange_custom_data (wrap->p4est, ghost, e->data_size,
-                                      e->patch_data,
-                                      e->ghost_contiguous_memory);
+    if (exchange_minlevel <= domain->global_minlevel &&
+        domain->global_maxlevel <= exchange_maxlevel)
+    {
+        p4est_ghost_exchange_custom (wrap->p4est, ghost, e->data_size,
+                                     e->patch_data,
+                                     e->ghost_contiguous_memory);
+    }
+    else
+    {
+        p4est_ghost_exchange_custom_levels (wrap->p4est, ghost,
+                                            exchange_minlevel,
+                                            exchange_maxlevel,
+                                            e->data_size, e->patch_data,
+                                            e->ghost_contiguous_memory);
+    }
 }
 
 void

### 2013-07-31 src/forestclaw2d.h

@@ -443,9 +443,12 @@ fclaw2d_domain_exchange_t
  *                              #(received patches) is domain->num_ghost_patches.
  * \param [in] e                Allocated buffers whose e->patch_data[i] pointers
  *                              must have been set properly by forestclaw.
+ * \param [in] exchange_minlevel The minimum quadrant level that is exchanged.
+ * \param [in] exchange_maxlevel The maximum quadrant level that is exchanged.
  */
 void fclaw2d_domain_ghost_exchange (fclaw2d_domain_t * domain,
-                                    fclaw2d_domain_exchange_t * e);
+                                    fclaw2d_domain_exchange_t * e,
+                                    int exchange_minlevel, int exchange_maxlevel);
 
 /** Free buffers used in exchanging off-processor data during time stepping.
  * This should be done just before regridding.

### 2013-07-31 src/amr_advance.cpp
added comments as to where EGPD is not needed

@@ -75,6 +75,7 @@ double advance_level(fclaw2d_domain_t *domain,
     double t_level = a_time_stepper->level_time(a_level);
     fclaw_bool time_interp_is_false = fclaw_false;
 
+    /* do an EGPD here, based on whether input flag do_egpd has been set */
 
     double maxcfl = 0;
     if (verbose)
@@ -117,6 +118,7 @@ double advance_level(fclaw2d_domain_t *domain,
                    be used in setting corners on coarser grids, if needed. */
                 set_phys_bc(domain,a_level,t_level,time_interp_is_false);
 
+                /* No EGDP needed here */
                 exchange_with_coarse(domain,a_level,t_level,alpha,
                                      FCLAW2D_TIMER_ADVANCE);
 
@@ -138,6 +140,7 @@ double advance_level(fclaw2d_domain_t *domain,
                         cout << " ----> Making recursive call to advance_level for level "
                              << a_level-1 << endl;
                     }
+                    /* set do_egpd = false */
                     maxcfl = advance_level(domain,a_level-1,last_coarse_step,a_time_stepper);
                     if (verbose)
                     {
@@ -165,6 +168,7 @@ double advance_level(fclaw2d_domain_t *domain,
                         cout << " ----> Making recursive call to advance_level for level "
                              << a_level-1 << endl;
                     }
+                    /* do_egpd = false */
                     maxcfl = advance_level(domain,a_level-1,last_coarse_step,a_time_stepper);
                     if (verbose)
                     {
@@ -200,6 +204,7 @@ double advance_level(fclaw2d_domain_t *domain,
                         cout << " --> Doing time interpolatation from coarse grid at level "
                              << a_level-1 << " using alpha = " << alpha << endl;
                     }
+                    /* Make call to EGPD */
                     exchange_with_coarse(domain,a_level,t_level,alpha,
                                          FCLAW2D_TIMER_ADVANCE);
 

### 2013-07-31 docs/parco13/paper.tex
Swap sections 3 and 4
Minor changes to text.

@@
 \newcommand{\clawpack}{{\sc Clawpack}\xspace}
-\newcommand{\forestclaw}{Forestclaw\xspace}
+\newcommand{\forestclaw}{ForestClaw\xspace}
 \newcommand{\pforest}{\texttt{p4est}\xspace}
 \newcommand{\manyclaw}{Manyclaw\xspace}
+
+\newcommand{\ignore}[1]{}
 
@@
 
-With the advent of many-core chips, such as GPUs and the MIC architecture, comes the opportunity to sustain unprecedented rates of floating point
+With the advent of many-core chips such as GPUs and the MIC architecture
+ comes the opportunity to sustain unprecedented rates of floating point
 operations at comparably high integration density and low cost.  These
@@
 and proven scalability \cite{BursteddeWilcoxGhattas11}.  This approach
-identifies each octree leaf with a mesh element.  The presented work does not
+identifies each octree leaf with a mesh element.  The present work does not
 construct a traditional high-order element but defines each element to be a
@@
 the numerical data ($m^d$ floating point values per patch).
+% \comment{It is probably more like M*m^d floating point values - with metric terms, velocity
+% fields, multiple copies of the solution (needed for time stepping, re-taking times steps, etc).
+% but the idea here is the same. What is a good way to say this? }
+
 \item The resulting parallel programming model is a hybrid (often referred
@@
 ghost cells surrounding each patch.  To this end, we have designed an
-interface that provides read-only access to the sequence of blocks, the list of
-patches for each, and a constant-time lookup of neighbor patches (and their
-relative orientation which is nontrivial between blocks).
+interface that provides \ignore{read-only} access to the sequence of blocks, the list of
+patches per-block, and a constant-time lookup of neighbor patches (and their
+relative and, in general, nontrivial orientation between blocks).
+% \comment{DAC: What do you mean by read-only?  Neighbors get updated by the ghost cell
+% exchange process...}
 % Since a neighbor patch has $m \ge 4$ cells in each direction, the required
@@
 % \cite{Forum94, SnirOttoHuss-LedermanEtAl96}.
+
+\section{Patch-based numerics at the leaf level}
+
+% CB: I'm trying to avoid "mesh" cell here since in \pforest FE simulations
+% that would mean a whole leaf.
+
+\alert{I have swapped sections 3. and 4.  Please check/smooth the text.}
+
+For hyperbolic problems, we integrate the solution on a single uniform patch,
+containing $m^d$ cells, using the wave propagation algorithm described by R. J.
+LeVeque \cite{LeVeque:1997eg} and implemented in \clawpack \cite{le:2002,
+clawpack}.  We assume a single degree of freedom per cell and reconstruct a
+piecewise constant solution to obtain left and right states at cell edges.  At
+each edge, we solve Riemann problems to obtain left and right going waves
+propagating at speeds determined from the solution to the Riemann problem.  For
+scalar advection, the speed of each wave is simply the local advection speed at
+the cell interface.  For non-linear problems and systems, an approximate
+Riemann solver, such as a Roe solver \cite{roesolver}, is typically used.  Since
+much of the physics of an application can be contained in the Riemann solver,
+\forestclaw adopts \clawpack's interface to Riemann solvers effectively allowing
+problems solved with \clawpack to be solvable in \forestclaw.  Wave limiters
+are used to supress spurious oscillations resulting from truncation error in a typical
+second order method.
+\comment{DAC: Limiters formally reduce the order of accuracy.}
+\ignore{In order to
+achieve second order accuracy wave limiters are used to suppress spurious
+oscillations at sharp gradients in the flow.}
+
+Data exchanges between neighboring patches are done via layers of
+ghost cells extending the dimensions along the edges of each patch.
+The interior edge values of a given patch overlap the ghost cell
+region of a neighboring patch.  For the second order wave propagation
+algorithm, two layers of ghost cells are sufficient.  This implies that
+one layer of ghost patches is sufficient for $m \ge 4$.  Neighboring
+patches at the same level of refinement simply copy their interior
+edge values into a neighbors' ghost cells.  Neighboring fine grid
+patches average their interior edge data to a coarser neighbor's ghost
+cell values.  And neighboring coarse grid patches interpolate data
+from their interior edge cells to their fine grid neighbor's ghost
+cell values.  To avoid loss of conservation and the creation of
+spurious extrema, we use a standard conservative, limited
+interpolation scheme to interpolate values from the coarse grid to
+fine grid coarse cells \cite{amrclaw, chombo}.  When sub-cycling, time accurate
+data between coarse grids is used to fill in ghost cells for fine grids.
+As mentioned in the previous section, this procedure can be extended
+transparently to distributed parallelism by defining an abstract exchange
+routine for ghost patch data.
+
+Mesh refinement and coarsening requires interpolation from coarse grids
+to newly created fine grids, and the averaging of fine grid data to a newly
+created underlying coarse grid.  This operation is rank-local, analogous to the
+general dynamic AMR procedures used in \pforest-based FE codes \cite[Fig.\
+4]{BursteddeGhattasStadlerEtAl08}.
 
@@
 The MPI layer is addressed from within \pforest and not exposed to the
-\forestclaw code.  The order of leaves is maintained in \pforest according to a
+\forestclaw code.  The \ignore{order of leaves} leaf ordering
+is maintained in \pforest according to a
 space filling curve.  Each MPI rank has a local view on its own partition,
@@
 
-\section{Patch-based numerics at the leaf level}
-
-% CB: I'm trying to avoid "mesh" cell here since in \pforest FE simulations
-% that would mean a whole leaf.
-
-\alert{Thought: Should we swap sections 3. and 4.?}
-
-For hyperbolic problems, we integrate the solution on a single uniform patch,
-containing $m^d$ cells, using the wave propagation algorithm described by R. J.
-LeVeque \cite{LeVeque:1997eg} and implemented in \clawpack \cite{le:2002,
-clawpack}.  We assume a single degree of freedom per cell and reconstruct a
-piecewise constant solution to obtain left and right states at cell edges.  At
-each edge, we solve Riemann problems to obtain left and right going waves
-propagating at speeds determined from the solution to the Riemann problem.  For
-scalar advection, the speed of each wave is simply the local advection speed at
-the cell interface.  For non-linear problems and systems, an approximate
-Riemann solver, such as a Roe solver \cite{roesolver}, is typically used.  Since
-much of the physics of an application can be contained in the Riemann solver,
-\forestclaw adopts \clawpack's interface to Riemann solvers effectively allowing
-problems solved with \clawpack to be solvable in \forestclaw.  In order to
-achieve second order accuracy wave limiters are used to suppress spurious
-oscillations at sharp gradients in the flow.
-
-Data exchanges between neighboring patches are done via layers of
-ghost cells extending the dimensions along the edges of each patch.
-The interior edge values of a given patch overlap the ghost cell
-region of a neighboring patch.  For the second order wave propagation
-algorithm, two layers of ghost cells are sufficient.  This implies that
-one layer of ghost patches is sufficient for $m \ge 4$.  Neighboring
-patches at the same level of refinement simply copy their interior
-edge values into a neighbors' ghost cells.  Neighboring fine grid
-patches average their interior edge data to a coarser neighbor's ghost
-cell values.  And neighboring coarse grid patches interpolate data
-from their interior edge cells to their fine grid neighbor's ghost
-cell values.  To avoid loss of conservation and the creation of
-spurious extrema, we use a standard conservative, limited
-interpolation scheme to interpolate values from the coarse grid to
-fine grid coarse cells \cite{amrclaw, chombo}.  When sub-cycling, time accurate
-data between coarse grids is used to fill in ghost cells for fine grids.
-As mentioned in the previous section, this procedure can be extended
-transparently to distributed parallelism by defining an abstract exchange
-routine for ghost patch data.
-
-Mesh refinement and coarsening requires interpolation from coarse grids
-to newly created fine grids, and the averaging of fine grid data to a newly
-created underlying coarse grid.  This operation is rank-local, analogous to the
-general dynamic AMR procedures used in \pforest-based FE codes \cite[Fig.\
-4]{BursteddeGhattasStadlerEtAl08}.
-
 \section{Numerical results}
@@
 \end{center}
-\caption{Strong scaling of MPI parallelism for advection on a uniform mesh.
-  Top: run time of time integration.  Bottom: time spent in local and parallel
-  neighbor exchanges.  The number of patches is $2^{2 \ell}$ at level $\ell$.
-  The flat lines for the smaller runs are caused by having
-  less patches than MPI processes, leaving some of them idle.  For each data point
-  we use the maximum time required across all processes.}
+\caption{Strong scaling of MPI parallelism for advection on a
+  uniformly refined mesh.  Top: run time of time integration.  Bottom:
+  time spent in local and parallel neighbor exchanges.  The number of
+  patches is $2^{2 \ell}$ at level $\ell$.  The flat lines for the
+  smaller runs are caused by having fewer patches than MPI processes,
+  leaving some of the processors idle.  For each data point we use the
+  maximum time required across all processes. \comment{This last
+    sentence is a bit unclear.  What do you mean exactly?}}
 \label{fig:uniscale}
 \end{figure}%
-We begin our analysis of Example 1 with uniform-mesh experiments run on TACC's
-Stampede supercomputer at different levels of resolution.  First of all,
-we use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
+We begin our analysis of Example 1 with uniformly-refined mesh experiments run on TACC's
+Stampede supercomputer at different levels of resolution.
+We use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
 scaling efficiency for the time integration between 97\%--103\% depending on
@@
 We have examined weak scalability as well by using four times as many processes
-for each level inclease, from level 6 at 64 cores to level 10 at 4096.  The
+for each level increase, from level 6 at 64 cores to level 10 at 4096.  The
 exchange times stay consistently below 16\% of the time integration.  The time
@@
 Next, we examine how we can reduce the overall wall clock time of a simulation
-by switching from uniform refinement to AMR with the same maximum level,
+by switching from uniformly refined meshes to adaptive refinement with the same maximum level,
 effectively coarsening where high resolution is not beneficial.  For an example
@@
 be reduced by up to a factor of 50 by combining AMR with subcycling.  The
-adaptive runs are still very well load balanced overall, which we infer from
+adaptive runs are still  load balanced overall, which we infer from
 the linear scaling of the time integration alone (not shown) and from the fact
@@
 tradeoff however, at least in the current implementation, that the adaptive
-runs suffer from the loss in scalability associated with the neighbor exchanges
-that we discussed earlier.
+runs suffer from the loss in scalability associated with the ghost patch exchanges
+described earlier. Currently, ghost patch exchanges are made across all levels, even
+if only two levels are involved in the exchange.
 

### 2013-07-31 docs/stampede/manyclaw/scripta.sh
Add two plot scripts

@@ -1,0 +1,22 @@
+#! /bin/sh
+
+MINL=1
+MAXL=5
+
+NODES="1 2 4"
+
+
+for NODE in $NODES ; do
+	# This defines the processor count
+	echo "# nodes $NODE"
+        perl -e "print 16 * $NODE"
+
+for LEVEL in `seq $MINL $MAXL` ; do
+	# This defines the problem size
+	#echo "# level $LEVEL"
+
+        grep "Max/P" "mx32_omp_16_level_${LEVEL}_mpi_${NODE}.out" | \
+		cut -d ' ' -f 5,6 | awk "{ printf (\" %g\", \$2 / \$1 / (4**$LEVEL)); }"
+done
+	echo
+done

### 2013-07-31 docs/stampede/manyclaw/scripte.sh
Remove level factor from scripts

@@ -15,8 +15,9 @@ for LEVEL in `seq $MINL $MAXL` ; do
 	# This defines the problem size
 	#echo "# level $LEVEL"
 
+	# Not computing the level factor
         grep "Max/P" "mx32_omp_16_level_${LEVEL}_mpi_${NODE}.out" | \
-		cut -d ' ' -f 8,9 | awk "{ printf (\" %g\", \$2 / \$1 / (4**$LEVEL)); }"
+		cut -d ' ' -f 8,9 | awk "{ printf (\" %g\", \$2 / \$1 / (1**$LEVEL)); }"
 done
 	echo
 done

### 2013-07-31 docs/parco13/paper.tex
Go through paper up to and including sec 3

@@
 for speed, while we inherit the flexibility of adaptive meshes.  In our work we
-choose to work with the \pforest AMR library since it allows us to compose the mesh
-from multiple mapped octrees and enables the cubed sphere and other
-nontrivial multiblock geometries.
+choose to integrate with the \pforest AMR library since it allows us to compose
+the mesh from multiple mapped octrees and enables the cubed sphere and other
+nontrivial multiblock geometries.  We describe aspects of the parallel
+implementation and close with scalings for both MPI-only and OpenMP/MPI hybrid
+runs, where the largest MPI-only run uses 4,096 CPU cores.
 \end{abstract}
@@
 
-In this paper we describe our design for the coupling of forest-of-octree AMR
+In this paper we describe our design for the coupling of forest-of-octrees AMR
 with \clawpack at the leaf level.  We comment on challenges that arise in
 enabling multiblock geometries and efficient parallelism and conclude with a range
-of numerical examples that demonstrate the conceptual improvements in relation
-to other approaches.
+of numerical examples that demonstrate the conceptual advantages.
+% improvements in relation to other approaches.
 
@@
 
-\alert{I have swapped sections 3. and 4.  Please check/smooth the text.}
-
 For hyperbolic problems, we integrate the solution on a single uniform patch,
@@
 second order method.
-\comment{DAC: Limiters formally reduce the order of accuracy.}
+% at the cost of a localized order reduction.
+% \comment{DAC: Limiters formally reduce the order of accuracy.}
+% wauld be an awkward senctence if order reduction is mentioned
 \ignore{In order to
@@
 data between coarse grids is used to fill in ghost cells for fine grids.
-As mentioned in the previous section, this procedure can be extended
+As we detail in the following section, this procedure can be extended
 transparently to distributed parallelism by defining an abstract exchange
@@
 between five MPI processes (the three middle ones out of $0 \ldots 4$
-are shown from black to white).}
+are shown with a color scale from black to white).}
 \label{fig:cubed3}

### 2013-07-31 src/amr_advance.cpp
Changes to selectively do ghost patch exchanges

@@ -66,16 +66,15 @@ static void update_level_solution(fclaw2d_domain_t *domain,
 
 static
 double advance_level(fclaw2d_domain_t *domain,
-                   const int a_level,
-                   const int a_curr_fine_step,
-                   subcycle_manager* a_time_stepper)
+                     const int a_level,
+                     const int a_curr_fine_step,
+                     subcycle_manager* a_time_stepper,
+                     fclaw_bool do_egpd)
 {
     const amr_options_t *gparms = get_domain_parms(domain);
     fclaw_bool verbose = (fclaw_bool) a_time_stepper->verbosity();
     double t_level = a_time_stepper->level_time(a_level);
     fclaw_bool time_interp_is_false = fclaw_false;
-
-    /* do an EGPD here, based on whether input flag do_egpd has been set */
 
     double maxcfl = 0;
     if (verbose)
@@ -83,6 +82,15 @@ double advance_level(fclaw2d_domain_t *domain,
         cout << endl;
         cout << "Advancing level " << a_level << " from step " <<
             a_curr_fine_step << " at time " << t_level << endl;
+    }
+
+    if (do_egpd)
+    {
+        /* Do a global exchange.  In the non-subcycled case, this is only true
+           if called directly from advance_level;  otherwise it is false.
+           The efficiency may still not be there in the subcycled case
+        */
+        exchange_ghost_patch_data(domain,time_interp_is_false);
     }
 
     if (!a_time_stepper->can_advance(a_level,a_curr_fine_step))
@@ -118,8 +126,9 @@ double advance_level(fclaw2d_domain_t *domain,
                    be used in setting corners on coarser grids, if needed. */
                 set_phys_bc(domain,a_level,t_level,time_interp_is_false);
 
-                /* No EGDP needed here */
-                exchange_with_coarse(domain,a_level,t_level,alpha,
+                /* This can still be improved */
+                fclaw_bool do_egpd = fclaw_true;
+                exchange_with_coarse(domain,a_level,t_level,alpha,do_egpd,
                                      FCLAW2D_TIMER_ADVANCE);
 
                 /* Set physicals BCs on the finer level, using any newly updated
@@ -141,7 +150,9 @@ double advance_level(fclaw2d_domain_t *domain,
                              << a_level-1 << endl;
                     }
                     /* set do_egpd = false */
-                    maxcfl = advance_level(domain,a_level-1,last_coarse_step,a_time_stepper);
[…]

### 2013-07-31 src/amr_exchangecoarse.cpp

@@ -694,7 +694,8 @@ void cb_setup_time_interp(fclaw2d_domain_t *domain,
 
 void exchange_with_coarse(fclaw2d_domain_t *domain,
                           int level, double t_level,
-                          double alpha, fclaw2d_timer_names_t running)
+                          double alpha, fclaw_bool do_egpd,
+                          fclaw2d_timer_names_t running)
 {
     // Use separate timer for all exchanges
     fclaw2d_domain_data_t *ddata = get_domain_data(domain);
@@ -703,7 +704,6 @@ void exchange_with_coarse(fclaw2d_domain_t *domain,
     }
     fclaw2d_timer_start (&ddata->timers[FCLAW2D_TIMER_EXCHANGE]);
 
-    // Simple exchange - no time interpolation needed
     fclaw_bool time_interp = alpha > 0; //
     int finer_level = level;
     int coarser_level = level - 1;
@@ -712,16 +712,19 @@ void exchange_with_coarse(fclaw2d_domain_t *domain,
     e_info.time_interp = time_interp;
     e_info.level = level;
 
-    /* copy griddata to griddata_time_sync so that parallel exchange will work */
     if (time_interp)
     {
+        /* Store time interpolated data into m_griddata_time_sync. */
         fclaw2d_domain_iterate_level(domain, coarser_level,cb_setup_time_interp,
                                      (void *) &alpha);
     }
 
-    /* Do parallel ghost exchange; Pointers to data get reassigned here, in case
-       we are in the time interpolated case */
-    exchange_ghost_patch_data(domain,time_interp);
+    if (do_egpd)
+    {
+        exchange_ghost_patch_data_levels(domain,time_interp,coarser_level,coarser_level);
+        /* Do parallel ghost exchange; Pointers to data get reassigned here, in case
+           we are in the time interpolated case */
+    }
 
 
     /* Iterate over coarser level and average from finer neighbors to coarse. */

### 2013-07-31 src/amr_forestclaw.H

@@ -76,6 +76,7 @@ void level_exchange (fclaw2d_domain_t * domain, int a_level,
 // Pass in the enum value for a running timer or FCLAW2D_TIMER_NONE.
 void exchange_with_coarse (fclaw2d_domain_t * domain, int a_level,
                            double t_level, double alpha,
+                           fclaw_bool do_egpd,
                            fclaw2d_timer_names_t running);
 
 // amr_physbc.cpp

### 2013-07-31 docs/stampede/manyclaw/scripta.sh
adding figs

@@ -1,9 +1,9 @@
 #! /bin/sh
 
 MINL=1
-MAXL=5
+MAXL=8
 
-NODES="1 2 4"
+NODES="1 2 4 8"
 
 
 for NODE in $NODES ; do

### 2013-07-31 docs/parco13/paper.tex
More text changes

@@
 implementation and close with scalings for both MPI-only and OpenMP/MPI hybrid
-runs, where the largest MPI-only run uses 4,096 CPU cores.
+runs, where the largest MPI-only run uses 16,384 CPU cores.
 \end{abstract}
@@
 different process (ghost patches).  Since this must make no difference
-numerically, we need to ensure that the values of ghost patches are current
+numerically, we need to ensure that the values of parallel neighbor patches are current
 whenever they are accessed.  To this end, we allocate storage for a layer of
@@
 The threaded parallelism over the degrees of freedom of a patch can be handled
-by \forestclaw alone without the need to involve \pforest.  Additionally,
-many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for
+by \forestclaw alone without the need to involve \pforest.  For instance, a
+many-core implementation, such as \manyclaw \cite{manyclaw}, can be used for
 the integration of the hyperbolic system on a leaf-patch thereby allowing for
@@
 numerical parameters that are set in each case include the size of the
-patch on each leaf ($m=8$ or $m=16$), and the minimum and maximum
+patch on each leaf ($m=8$, $16$, or $32$), and the minimum and maximum
 refinement levels, which in turn fix minimum and maximum effective mesh
@@
 right half.  A time dependent flow field is prescribed that distorts
-the interface between the 0 and 1 concentration values.  The minimum
-and maximum levels of refinement are set to 3 and 6, respectively.
+the interface between the 0 and 1 concentration values.  
+\Fig{results2d} shows the results at an intermediate time step, where
+the minimum and maximum levels of refinement are set to 3 and 6, respectively.
 This results in a minimum resolution of $64 \times 64$ and a maximum
-resolution of $512 \times 512$.  \Fig{results2d} shows the results
-at an intermediate time step.
+resolution of $512 \times 512$.
 
@@
 \cite{ca-he-le:2008, be-ca-he-le:2009}.  Each quadtree is mapped to cover one
-hemisphere.  The initial conditions is $q = 0$ in one half of
+hemisphere.  The initial condition is $q = 0$ in one half of
 the sphere, and $q = 1$ in the other half, where the halves are not
@@
   leaving some of the processors idle.  For each data point we use the
-  maximum time required across all processes. \comment{This last
-    sentence is a bit unclear.  What do you mean exactly?}}
+  time required by the slowest processor to verify that the load is
+  balanced equally.
+%  all processes. \comment{This last
+%    sentence is a bit unclear.  What do you mean exactly?}
+}
 \label{fig:uniscale}
@@
 We have examined weak scalability as well by using four times as many processes
-for each level increase, from level 6 at 64 cores to level 10 at 4096.  The
-exchange times stay consistently below 16\% of the time integration.  The time
-in seconds per time integration step varies between .0578 and .0602,
-which yields a parallel efficiency of 96\%.
+for each level increase, from level 6 at 64 cores to level 11 at 16,384.  Up to
+4,096 cores, the exchange times stay consistently below 16\% of the time
+integration and the time in seconds per time integration step varies between
+.0578 and .0602, which yields a parallel efficiency of 96\%.  For the
+data point at 16,384 cores, the efficiency is still 82\%.
 
@@
 by switching from uniformly refined meshes to adaptive refinement with the same maximum level,
-effectively coarsening where high resolution is not beneficial.  For an example
+effectively coarsening where high resolution is not needed.  For an example
 with a sharp front as depicted in \Fig{results2d} we expect considerable
@@
 tradeoff however, at least in the current implementation, that the adaptive
-runs suffer from the loss in scalability associated with the ghost patch exchanges
-described earlier. Currently, ghost patch exchanges are made across all levels, even
-if only two levels are involved in the exchange.
+runs exhibit reduced scalability associated with the ghost patch exchanges
+described earlier.  Presently, ghost patch exchanges are made across all levels, even
+if only two levels are involved in the exchange.  One enhancement of the code that is
+currently being developed is to rewrite the time stepping loop with the
+objective of minimizing the parallel neighbor exchanges.
 
@@
 To demonstrate a multiblock connectivity, we include results for Example 2 in
-\Fig{spherical}.
+\Fig{spherical}, together with a strong scaling table that shows that the spherical
+example scales as well.
 

### 2013-07-31 docs/parco13/paper.tex
Comment
Add sphere table

@@
   \begin{center}
-  \begin{minipage}{.4\columnwidth}
-    \begin{tabular}{cc}
-      timing table for sphere & 123, \\
+    \hspace{.1\columnwidth}
+    \begin{minipage}[c]{.4\columnwidth}
+      \includegraphics[width=.99\columnwidth]{../figs_misc/sphere_partition5}
+    \end{minipage}
+    \hspace{.05\columnwidth}
+    \begin{minipage}[c]{.4\columnwidth}
+    \begin{tabular}{r|r}
+       $P$ & time int./step \\
+      \hline
+       16 & 1.115 \\
+       32 & 0.557 \\
+       64 & 0.297 \\
+      256 & 0.092 \\
     \end{tabular}
   \end{minipage}
-  \hspace{.05\columnwidth}
-  \includegraphics[width=.4\columnwidth]{../figs_misc/sphere_partition5}
 \end{center}
-\caption{Partition of the spherical Example 2 between five MPI processors, indicated
+\caption{Left: Partition of the spherical Example 2 between five MPI processors, indicated
   by the color scale, at simulated time $T = 1$.  The partition shows the mesh equator,
   while the refinement is invariant under the number of processors and the partition
-and follows the front, which is not aligned with the equator.}
+  and follows the front, which is not aligned with the equator.  Right: Strong scaling
+  for a uniform run at level 8.}
 \label{fig:spherical}
@@
 We would like to thank the Texas Advanced Computing Center (TACC) for access to
-the Stampede supercomputer under allocation TG-DPP130002.
+the Stampede supercomputer under allocation TG-DPP130002 granted by the NSF.
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha

### 2013-07-31 docs/parco13/paper.tex
Edit
Adding omp results.

--- 2013-07-31 docs/parco13/paper.tex
@@
 
+\begin{figure}
+\begin{center}
+  \includegraphics[angle=-90,width=.8\columnwidth]{../figs_misc/omp.eps}
+  \\[1ex]
+  \includegraphics[angle=-90,width=.8\columnwidth]{../figs)misc/advance_strong_omp.eps}
+\end{center}
+\caption{Strong scaling of OpenMP parallelism for advection on a
+  uniformly refined mesh.
+Top: run time of advection Riemann problem kernel in both an optimized inlined interface and a generic function pointer interface via \manyclaw.
+Bottom: MPI + OpenMp scaling results of advance kernel as above.
+}
+\label{fig:ompscale}
+\end{figure}%
+
+In \Fig{ompscale}, we scale intranode parallel calculations and in future iterations on accerelrators the \manyclaw set of solvers were compuled with \forestclaw.
+\manyclaw provides a generic set of interators for Riemann problems and related algorithms.
+In this work we show good scaling results across a single node and coupled to the MPI scaling as well.
+
 We have examined weak scalability as well by using four times as many processes
@@
 We would like to thank the Texas Advanced Computing Center (TACC) for access to
-the Stampede supercomputer under allocation TG-DPP130002.
+the Stampede supercomputer under allocation TG-DPP130002 and TG-ASC130001 granted by the NSF XSEDE Program. 
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha
@@
 presented by B.\ as part of a talk at the SCI Institute, Salt Lake City, Utah,
-USA in July 2011.
+USA in July 2011. The fourth authors recognize Simula National Lab for funding.
 

### 2013-07-31 docs/parco13/paper.tex

@@
 \Fig{spherical}, together with a strong scaling table that shows that the spherical
-example scales as well.
+example scales reasonably well too.
 

### 2013-07-31 docs/parco13/conc.tex
Start conc

\section{Conclusion}

We have presented the integration of an MPI-based forest-of-octrees adaptive
meshing code, \pforest, with a numerical solver for hyperbolic
conservation laws, \clawpack/\manyclaw,
that implements threaded parallelism for a single compute patch
on the other.  All topological operations on the mesh are handled by \pforest,
while all numerical operations are handled by the \clawpack implementation.
This combination naturally enables a hybrid MPI/threaded parallelization.
We demonstrate scalability of this approach in various ways, where the MPI-only variant
scales up to 16,384 codes

### 2013-07-31 docs/parco13/paper.tex
Start conc
hacking conclusion.

@@
 
+\section{Conclusion}
+
+In this work, we have shown a novel new hybrid paradigm for AMR that achieves good scaling both in distributed and shared memory aspects of current machines.
+By using each leaf in the AMR hierarchy, we are able to use many different low level solvers and manage the exchanges between them efficiently.
+The flexibility and scalability of this approach provides an ideal workspace to explore further algorithms and physics problems requiring new solvers at high resolutions.
+
 % The results in this section are preliminary and will be extended in the final

### 2013-07-31 docs/parco13/conc.tex
Make figures fit, improvise conclusion

@@
 We demonstrate scalability of this approach in various ways, where the MPI-only variant
-scales up to 16,384 codes
+scales weakly up to 16,384 codes, and strong scalability is linear for up to 256 cores
+in the MPI-only and 128 in the threaded implementation

### 2013-07-31 docs/parco13/paper.tex
Make figures fit, improvise conclusion

--- 2013-07-31 docs/parco13/paper.tex
@@
 \begin{center}
-\plotbox{\includegraphics[width=.4\columnwidth]{quadtree}}
+\plotbox{\includegraphics[width=.37\columnwidth]{quadtree}}
 \hspace{.05\columnwidth}
-\plotbox{\includegraphics[width=.4\columnwidth]{sphere_plot167}}
+\plotbox{\includegraphics[width=.37\columnwidth]{sphere_plot167}}
 \end{center}
@@
 \begin{center}
-  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_advance_strong-crop}
+  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_advance_strong-crop}
   \\[1ex]
-  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_exchange_strong-crop}
+  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_exchange_strong-crop}
 \end{center}
@@
 \begin{center}
-  \includegraphics[angle=-90,width=.8\columnwidth]{../figs_misc/omp.eps}
-  \\[1ex]
-  \includegraphics[angle=-90,width=.8\columnwidth]{../figs)misc/advance_strong_omp.eps}
+%  \includegraphics[angle=-90,width=.8\columnwidth]{../figs_misc/omp}
+%  \\[1ex]
+  \includegraphics[angle=-90,width=.7\columnwidth]{../figs_misc/advance_strong_omp}
 \end{center}
 \caption{Strong scaling of OpenMP parallelism for advection on a
-  uniformly refined mesh.
-Top: run time of advection Riemann problem kernel in both an optimized inlined interface and a generic function pointer interface via \manyclaw.
-Bottom: MPI + OpenMp scaling results of advance kernel as above.
+  uniformly refined mesh:
+  %Top: run time of advection Riemann problem kernel in both an optimized inlined interface and a generic function pointer interface via \manyclaw.  Bottom:
+  MPI + OpenMp hybrid parallelization of the time integration.
 }
@@
 
-In \Fig{ompscale}, we scale intranode parallel calculations and in future iterations on accerelrators the \manyclaw set of solvers were compuled with \forestclaw.
-\manyclaw provides a generic set of interators for Riemann problems and related algorithms.
-In this work we show good scaling results across a single node and coupled to the MPI scaling as well.
+In \Fig{ompscale}, we scale multi-node parallel calculations.
+%and in future
+%iterations on accelerators 
+We use the \manyclaw set of patch-based PDE solvers.
+%with
+%\forestclaw.  
+\manyclaw provides a generic set of interators for Riemann
+problems and related algorithms.
+
+%$In this work we show good scaling results
+%across a single node and coupled to the MPI scaling as well.
 
@@
 
+\input{conc}
+
 % The results in this section are preliminary and will be extended in the final

### 2013-07-31 docs/parco13/paper.tex

@@
 \begin{center}
-\plotbox{\includegraphics[width=.37\columnwidth]{quadtree}}
+\plotbox{\includegraphics[width=.35\columnwidth]{quadtree}}
 \hspace{.05\columnwidth}
-\plotbox{\includegraphics[width=.37\columnwidth]{sphere_plot167}}
+\plotbox{\includegraphics[width=.35\columnwidth]{sphere_plot167}}
 \end{center}
@@
 \begin{center}
-  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_advance_strong-crop}
+  \includegraphics[angle=-90,width=.7\columnwidth]{pdf/uni_advance_strong-crop}
   \\[1ex]
-  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_exchange_strong-crop}
+  \includegraphics[angle=-90,width=.7\columnwidth]{pdf/uni_exchange_strong-crop}
 \end{center}
@@
     \hspace{.1\columnwidth}
-    \begin{minipage}[c]{.4\columnwidth}
+    \begin{minipage}[c]{.35\columnwidth}
       \includegraphics[width=.99\columnwidth]{../figs_misc/sphere_partition5}
@@
 To demonstrate a multiblock connectivity, we include results for Example 2 in
-\Fig{spherical}, together with a strong scaling table that shows that the spherical
-example scales reasonably well too.
+\Fig{spherical}, together with a strong scaling table.
 
-\input{conc}
+\section{Conclusion}
+
+We have presented a novel hybrid paradigm for AMR that achieves good
+scaling both in the distributed and shared memory aspect of current
+supercomputers.
+%
+% By
+% abstracting a leaf in the AMR hierarchy as the atomic unit of computation, we
+% are able to use almost arbitrary low level solvers.
+%
+Our implementation integrates an MPI-based forest-of-octrees adaptive
+meshing code, \pforest, with a numerical solver for hyperbolic
+conservation laws, \clawpack/\manyclaw,
+that implements threaded parallelism for a single compute patch.
+%All topological operations on the mesh are handled by \pforest,
+%while all numerical operations are handled by the \clawpack instance.
+This combination naturally enables MPI/threaded hybridization,
+where the MPI version scales to 16,384 CPU cores.
+
+%
+% We demonstrate scalability of this approach in various ways, where the MPI-only variant
+% scales weakly up to 16,384 codes, and strong scalability is linear for up to 256 cores
+% in the MPI-only and 128 in the threaded implementation, and our results that higher
+% core counts are easily possible.
+%
+% d to manage the exchanges between them efficiently.  The
+% flexibility and scalability of this approach provides an ideal workspace to
+% explore further algorithms and physics problems requiring new solvers at high
+% resolutions.
+%
+%the versatility of this approach the modular development of
+%scalable codes.
 
@@
 (Cambridge, UK), where much of the preliminary development work for
-\forestclaw was done. The leaf/patch paradigm was independently
+\forestclaw was done.
+The fourth author recognizes Simula National Lab for funding.
+The leaf/patch paradigm was independently
 presented by B.\ as part of a talk at the SCI Institute, Salt Lake City, Utah,
-USA in July 2011. The fourth authors recognize Simula National Lab for funding.
+USA in July 2011.
 

### 2013-07-31 docs/parco13/paper.tex
Edits to the paper

@@
 \begin{center}
-  \includegraphics[angle=-90,width=.7\columnwidth]{pdf/uni_advance_strong-crop}
+  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_advance_strong-crop}
   \\[1ex]
-  \includegraphics[angle=-90,width=.7\columnwidth]{pdf/uni_exchange_strong-crop}
+  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_exchange_strong-crop}
 \end{center}
@@
 %  \\[1ex]
-  \includegraphics[angle=-90,width=.7\columnwidth]{../figs_misc/advance_strong_omp}
+  \includegraphics[angle=-90,width=.75\columnwidth]{../figs_misc/advance_strong_omp}
 \end{center}
-\caption{Strong scaling of OpenMP parallelism for advection on a
-  uniformly refined mesh:
-  %Top: run time of advection Riemann problem kernel in both an optimized inlined interface and a generic function pointer interface via \manyclaw.  Bottom:
-  MPI + OpenMp hybrid parallelization of the time integration.
+\caption{Strong scaling of MPI/OpenMP parallelism for advection on a
+  uniformly refined mesh with 16 threads per node.
+  %Top: run time of advection Riemann problem kernel in both an optimized
+  %inlined interface and a generic function pointer interface via \manyclaw.
+  %Bottom:
+  %MPI + OpenMp hybrid parallelization of the time integration.
 }
@@
 
-In \Fig{ompscale}, we scale multi-node parallel calculations.
+In \Fig{ompscale}, we show multi-node calculations using 16 OpenMP
+threads per node executed by the
 %and in future
 %iterations on accelerators 
-We use the \manyclaw set of patch-based PDE solvers.
+\manyclaw set of patch-based PDE solvers.
 %with
 %\forestclaw.  
-\manyclaw provides a generic set of interators for Riemann
-problems and related algorithms.
+%\manyclaw provides a generic set of interators for Riemann
+%problems and related algorithms.
 
@@
 \end{figure}
-To demonstrate a multiblock connectivity, we include results for Example 2 in
-\Fig{spherical}, together with a strong scaling table.
+To showcase a multiblock connectivity, we include results for the spherical Example 2 in
+\Fig{spherical}, together with a strong scaling table from 16 to 256 MPI ranks.
 
@@
 We have presented a novel hybrid paradigm for AMR that achieves good
-scaling both in the distributed and shared memory aspect of current
+scaling both in the distributed and shared memory aspects of current
 supercomputers.
@@
 This combination naturally enables MPI/threaded hybridization,
-where the MPI version scales to 16,384 CPU cores.
+where our biggest MPI run scales to 16,384 CPU cores.
 
@@
 We would like to thank the Texas Advanced Computing Center (TACC) for access to
-the Stampede supercomputer under allocation TG-DPP130002 and TG-ASC130001 granted by the NSF XSEDE Program. 
+the Stampede supercomputer under allocations TG-DPP130002 and TG-ASC130001
+granted by the NSF XSEDE Program.
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha

### 2013-08-01 docs/parco13/paper.tex
Crop a pdf
Tweak the paper for resubmission

@@
 implementation and close with scalings for both MPI-only and OpenMP/MPI hybrid
-runs, where the largest MPI-only run uses 16,384 CPU cores.
+runs, where the largest MPI run executes on 16,384 CPU cores.
 \end{abstract}
@@
 %  \\[1ex]
-  \includegraphics[angle=-90,width=.75\columnwidth]{../figs_misc/advance_strong_omp}
+  \includegraphics[angle=-90,width=.75\columnwidth]{../figs_misc/advance_strong_omp-crop}
 \end{center}
@@
 
-We have examined weak scalability as well by using four times as many processes
+We have examined weak scalability as well by using four times as many MPI
+processes
 for each level increase, from level 6 at 64 cores to level 11 at 16,384.  Up to
@@
     \hspace{.1\columnwidth}
-    \begin{minipage}[c]{.35\columnwidth}
-      \includegraphics[width=.99\columnwidth]{../figs_misc/sphere_partition5}
+    \begin{minipage}[c]{.30\columnwidth}
+      \includegraphics[width=\columnwidth]{../figs_misc/sphere_partition5}
     \end{minipage}
@@
 \end{center}
-\caption{Left: Partition of the spherical Example 2 between five MPI processors, indicated
-  by the color scale, at simulated time $T = 1$.  The partition shows the mesh equator,
-  while the refinement is invariant under the number of processors and the partition
+\caption{Left: Partition of the spherical Example 2 between five
+  MPI processes, indicated by the color scale, at simulated time $T = 1$.
+  % The  partition shows the mesh equator,
+  The refinement is invariant under the parallel partition
   and follows the front, which is not aligned with the equator.  Right: Strong scaling
@@
 
-We have presented a novel hybrid paradigm for AMR that achieves good
-scaling both in the distributed and shared memory aspects of current
-supercomputers.
-%
-% By
-% abstracting a leaf in the AMR hierarchy as the atomic unit of computation, we
-% are able to use almost arbitrary low level solvers.
-%
-Our implementation integrates an MPI-based forest-of-octrees adaptive
-meshing code, \pforest, with a numerical solver for hyperbolic
-conservation laws, \clawpack/\manyclaw,
-that implements threaded parallelism for a single compute patch.
-%All topological operations on the mesh are handled by \pforest,
-%while all numerical operations are handled by the \clawpack instance.
-This combination naturally enables MPI/threaded hybridization,
-where our biggest MPI run scales to 16,384 CPU cores.
+We have presented the integration of an MPI-based forest-of-octrees adaptive
+meshing code, \pforest, with a numerical solver for hyperbolic conservation
+laws, \clawpack/\manyclaw, that implements threaded parallelism for a single
+compute patch.
+%All mesh operations are handled by \pforest,
+%while numerics are handled by the \clawpack implementation.
+We abstract an interface to the parallel meshing code in order to derive the
+schedule of neighbor exchanges, averaging/interpolation, and time integration
+on the local patches.
+This approach naturally lends itself to MPI/thread hybridization.
+Apart from its parallel scalability, we favor the presented strategy
+for its modularity, encapsulation, and versality.  Future work
+will continue to optimize the parallel neighbor exchange patterns
+and investigate a generic handling of arbitrary multi-block geometries.
 
+% We have presented a novel hybrid paradigm for AMR that achieves good
+% scaling both in the distributed and shared memory aspects of current
+% supercomputers.
+% %
+% % By
+% % abstracting a leaf in the AMR hierarchy as the atomic unit of computation, we
+% % are able to use almost arbitrary low level solvers.
+% %
+% Our implementation integrates an MPI-based forest-of-octrees adaptive
+% meshing code, \pforest, with a numerical solver for hyperbolic
+% conservation laws, \clawpack/\manyclaw,
+% that implements threaded parallelism for a single compute patch.
+% %All topological operations on the mesh are handled by \pforest,
+% %while all numerical operations are handled by the \clawpack instance.
+% This combination naturally enables MPI/threaded hybridization,
+% where our biggest MPI run scales to 16,384 CPU cores.
 %
@@
 % resolutions.
-%
-%the versatility of this approach the modular development of
-%scalable codes.
 
@@
 the Stampede supercomputer under allocations TG-DPP130002 and TG-ASC130001
-granted by the NSF XSEDE Program.
+granted by the NSF XSEDE program.
 The authors acknowledge valuable discussion with Randy LeVeque, Marsha
@@
 \forestclaw was done.
-The fourth author recognizes Simula National Lab for funding.
+The fourth author recognizes Simula Research Lab, Norway, for funding.
 The leaf/patch paradigm was independently

### 2013-10-31 docs/parco13/paper.tex
Tweak section on parallelization

@@
 (see \Fig{cubed3}).
+The structure of \forestclaw is oblivious to the fact that it only has this
+local view, which relieves the developer from programming to the MPI interface.
 
@@
 different block.  In the latter case, coordinate transformations are carried
-out.  The structure of \forestclaw is oblivious to the fact that it only has a
-local view of the distributed mesh and data which relieves the developer from
-programming to the MPI interface.
+out.
 % Parallel neighbor exchanges are hidden
@@
 % as follows.
-
-With MPI parallelism, neighbor patches can be either local or assigned to a
-different process (ghost patches).  Since this must make no difference
+Irrespective of their block,
+% With MPI parallelism,
+neighbor patches can be either local or assigned to a different MPI process (in
+which case we call them ghost patches).  Since this must make no difference
 numerically, we need to ensure that the values of parallel neighbor patches are current
@@
 then correction factors are propagated back from fine to coarse levels, requiring
-an exchange at each level.  With sub-cycling, this entails a recursion with an
+at most two exchanges per level.  With sub-cycling, this entails a recursion with an
 operation count that is exponential in the difference between the largest and

### 2013-10-31 docs/parco13/paper.tex
Tweaks and new weak scaling table

@@
 
-We provide two examples that demonstrate the \forestclaw code.
+We provide two examples that exercise the \forestclaw code.
 The numerical results to date have been designed to verify that
@@
 interpolation) are implemented correctly.
-The basic \clawpack algorithm and corresponding code are
+The basic \clawpack algorithms and the corresponding code are
 thoroughly tested and need no further verification in our context.
@@
 %iterations on accelerators 
-\manyclaw set of patch-based PDE solvers.
+\manyclaw set of patch-based PDE solvers \cite{manyclaw}.
 %with
@@
 
+\begin{table}
+  \begin{center}
+   \begin{tabular}{l|rrrrrr}
+      $P$ (\# total MPI ranks) & 16 & 64 & 256 & 1,024 & 4,096 & 16,384 \\
+      \hline
+      par.\ efficiency (\%) & 100 & 96 & 96 & 103 & 103 & 82 \\
+    \end{tabular}
+  \end{center}
+  \caption{Weak scalability on a uniform mesh from level 6 at 64 cores to level
+    11 at 16,384
+    % using
+    (16 MPI ranks/node).}%
+  \label{tab:weakscale}
+\end{table}%
+%
 We have examined weak scalability as well by using four times as many MPI
-processes
-for each level increase, from level 6 at 64 cores to level 11 at 16,384.  Up to
-4,096 cores, the exchange times stay consistently below 16\% of the time
-integration and the time in seconds per time integration step varies between
-.0578 and .0602, which yields a parallel efficiency of 96\%.  For the
-data point at 16,384 cores, the efficiency is still 82\%.
+processes for each level increase.
+% , from level 6 at 64 cores to level 11 at 16,384.
+Up to 4,096 cores, the exchange times stay consistently below 16\% of the time
+integration,
+% and the time in seconds per time integration step varies between
+% .0578 and .0602,
+and we obtain optimal parallel efficiency.  For
+% the data point at
+16,384 cores, the efficiency reduces to 82\%
+with the current code (see \Tab{weakscale}).
 
@@
 runs exhibit reduced scalability associated with the ghost patch exchanges
-described earlier.  Presently, ghost patch exchanges are made across all levels, even
-if only two levels are involved in the exchange.  One enhancement of the code that is
-currently being developed is to rewrite the time stepping loop with the
-objective of minimizing the parallel neighbor exchanges.
+described earlier.
+% One enhancement of the code that is currently being developed is to rewrite
+% the time stepping loop with the objective of minimizing the parallel neighbor
+% exchanges.
+%
+% Presently, ghost patch exchanges are made across all levels, even
+% if only two levels are involved in the exchange.  One enhancement of the code that is
+% currently being developed is to rewrite the time stepping loop with the
+% objective of minimizing the parallel neighbor exchanges.
 

### 2013-10-31 docs/parco13/paper.tex
Use latest sphere strong scaling results
Two tweaks

@@
 %
-We have examined weak scalability as well by using four times as many MPI
+We examine weak scalability as well by using four times as many MPI
 processes for each level increase.
@@
   \begin{center}
-    \hspace{.1\columnwidth}
-    \begin{minipage}[c]{.30\columnwidth}
+    %\hspace{.1\columnwidth}
+    \begin{minipage}[c]{.3\columnwidth}
       \includegraphics[width=\columnwidth]{../figs_misc/sphere_partition5}
     \end{minipage}
-    \hspace{.05\columnwidth}
-    \begin{minipage}[c]{.4\columnwidth}
-    \begin{tabular}{r|r}
-       $P$ & time int./step \\
+    \hspace{3ex}
+    \begin{minipage}[c]{.44\columnwidth}
+    \begin{tabular}{rcr}
+      \multicolumn{1}{c}{$P$}
+      & time int./step (s) & eff.\ (\%) \\
       \hline
-       16 & 1.115 \\
-       32 & 0.557 \\
-       64 & 0.297 \\
-      256 & 0.092 \\
+         16 & 3.539  & 100.0 \\
+       % 32 & 1.763  &  99.6 \\
+         64 & 0.889  & 100.5 \\
+      % 128 & 0.445  & 100.6 \\
+        256 & 0.223  & 100.7 \\
+      1,024 & 0.055 & 100.2 \\
+      4,096 & 0.014 &  99.4 \\
+% old results
+%        16 & 1.115 \\
+%        32 & 0.557 \\
+%        64 & 0.297 \\
+%       256 & 0.092 \\
     \end{tabular}
@@
 To showcase a multiblock connectivity, we include results for the spherical Example 2 in
-\Fig{spherical}, together with a strong scaling table from 16 to 256 MPI ranks.
+\Fig{spherical}, together with a strong scaling table from 16 to 4,096 MPI ranks.
 

### 2017-07-20 src/patches/clawpatch/fclaw2d_clawpatch_output_netcdf.c
finish serial version implementation

@@ -32,28 +32,48 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 #include <fclaw2d_global.h>
 #include <fclaw2d_options.h>
 
+#include <netcdf.h>
 
-void cb_clawpatch_output_netcdf (fclaw2d_domain_t * domain,
-                                fclaw2d_patch_t * this_patch,
-                                int this_block_idx, int this_patch_idx,
-                                void *user)
+#define ERRCODE 2
+#define ERR(e) {printf("Error: %s\n", nc_strerror(e)); exit(ERRCODE);}
+
+typedef struct netcdf_data
 {
-#if 0
+    int ncid;
+    int step_inc;
+    int total_steps;  /* This level takes total_steps of step_inc each */
+
+    double initial_time;
+    double current_time;
+    double dt_step;
+} fclaw2d_level_data_t;
+
+void cb_clawpatch_output_netcdf_defpatch (fclaw2d_domain_t * domain,
+                                          fclaw2d_patch_t * this_patch,
+                                          int this_block_idx, int this_patch_idx,
+                                          void *user)
+{
     fclaw2d_global_iterate_t* s = (fclaw2d_global_iterate_t*) user;
     fclaw2d_global_t *glob = (fclaw2d_global_t*) s->glob;
-
-    fclaw2d_clawpatch_vtable_t *clawpatch_vt = fclaw2d_clawpatch_vt();
-    const fclaw_options_t *gparms = fclaw2d_get_options(glob);
+    int ncid = *((int *) s->user);
+    int subgroupid;
+    int retval;
+    int x_dimid, y_dimid, meqn_dimid, qid;
+    int dimids[3];
 
     int patch_num;
     int level;
     int mx,my,mbc,meqn;
-    double xlower,ylower,dx,dy;
+    double xlower,ylower,xupper,yupper,dx,dy;
     double *q;
-    // char fname[11];
 
-    int iframe = *((int *) s->user);
+    char patchname[10];
+    sprintf(patchname, "patch%04d", this_patch_idx);
+    /* Create subgroup */
+    if ((retval = nc_def_grp(ncid, patchname, &subgroupid)))
+    ERR(retval);
 
[…]

### 2017-07-20 src/patches/clawpatch/fclaw2d_clawpatch_output_netcdf.c
use patch_num instead of patch_idx to define subgroup

@@ -67,13 +67,6 @@ void cb_clawpatch_output_netcdf_defpatch (fclaw2d_domain_t * domain,
     double xlower,ylower,xupper,yupper,dx,dy;
     double *q;
 
-    char patchname[10];
-    sprintf(patchname, "patch%04d", this_patch_idx);
-    /* Create subgroup */
-    if ((retval = nc_def_grp(ncid, patchname, &subgroupid)))
-    ERR(retval);
-
-    const char* dim_name[2] = {"x","y"};    
     /* Get info not readily available to user */
     fclaw2d_patch_get_info(glob->domain,this_patch,
                            this_block_idx,this_patch_idx,
@@ -83,6 +76,15 @@ void cb_clawpatch_output_netcdf_defpatch (fclaw2d_domain_t * domain,
                                 &xlower,&ylower,&dx,&dy);
 
     fclaw2d_clawpatch_soln_data(glob,this_patch,&q,&meqn);
+
+
+    char patchname[10];
+    sprintf(patchname, "patch%04d", patch_num);
+    /* Create subgroup */
+    if ((retval = nc_def_grp(ncid, patchname, &subgroupid)))
+    ERR(retval);
+
+    const char* dim_name[2] = {"x","y"}; 
 
     xupper = xlower + mx*dx;
     yupper = ylower + my*dy;
@@ -147,9 +149,15 @@ void cb_clawpatch_output_netcdf_writeq (fclaw2d_domain_t * domain,
     int ncid = *((int *) s->user);
     int groupid,qid;
     int retval;
+    int patch_num, level;
+
+    /* Get info not readily available to user */
+    fclaw2d_patch_get_info(glob->domain,this_patch,
+                           this_block_idx,this_patch_idx,
+                           &patch_num,&level);
     
     char patchname[10];
-    sprintf(patchname, "patch%04d", this_patch_idx);
+    sprintf(patchname, "patch%04d", patch_num);
     if ((retval = nc_inq_grp_ncid(ncid, patchname, &groupid)))
         ERR(retval);
     if ((retval = nc_inq_varid(groupid, "q", &qid)))

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/clawpack46_setaux.f
(advection/swirl-conservative form) Add conservative Riemann solvers to 4.6 subdirectory

@@ -1,0 +1,31 @@
+      subroutine clawpack46_setaux(maxmx,maxmy,mbc,mx,my,
+     &      xlower,ylower,dx,dy,maux,aux)
+      implicit none
+
+      integer mbc, mx, my, maux, maxmx, maxmy
+      double precision xlower, ylower, dx, dy
+      double precision  aux(1-mbc:maxmx+mbc,1-mbc:maxmy+mbc,maux)
+
+      integer i, j
+      double precision xc, yc, ucc, pi
+
+      common /compi/ pi
+
+      do i = 1-mbc,mx+mbc
+         do j = 1-mbc,my+mbc
+c           # coordinates of lower left corner of grid cell:
+            xc = xlower + (i-0.5)*dx
+            yc = ylower + (j-0.5)*dy
+
+c           ucc = 2*((sin(pi*xp))**2 * sin(pi*yp) * cos(pi*yp))
+c           ucc = cos(2*pi*xp)
+
+            ucc = 0.1*sin(2*pi*xc)*sin(16*pi*xc)
+
+            aux(i,j,1) = ucc
+            aux(i,j,2) = 0.d0
+         enddo
+      enddo
+
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_new.f

@@ -1,0 +1,64 @@
+      subroutine rpn2cons(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
+     &                       auxl,auxr,wave,s,amdq,apdq)
+
+      implicit none
+
+      integer maxm, mbc,mwaves,meqn,maux, mx
+      integer ixy
+
+      double precision wave(1-mbc:maxm+mbc, meqn, mwaves)
+      double precision    s(1-mbc:maxm+mbc, mwaves)
+      double precision   ql(1-mbc:maxm+mbc, meqn)
+      double precision   qr(1-mbc:maxm+mbc, meqn)
+      double precision amdq(1-mbc:maxm+mbc, meqn)
+      double precision apdq(1-mbc:maxm+mbc, meqn)
+      double precision auxl(1-mbc:maxm+mbc,*)
+      double precision auxr(1-mbc:maxm+mbc,*)
+
+
+      integer i, iface, m, mw, version
+      double precision ur,ul,qrr,qll, uhat,qhat, tau,uavg
+
+
+      double precision dtcom, dxcom, dycom, tcom
+      integer icom, jcom
+
+      common/comxyt/dtcom,dxcom,dycom,tcom,icom,jcom
+
+      if ((mwaves .ne. 1) .or. (meqn .ne. 1)) then
+         write(6,*) 'rpn2cons : mwaves .ne. 1 or meqn .ne. 1'
+         stop
+      endif
+
+
+      iface = ixy
+      do i = 2-mbc, mx+mbc
+
+         ur = auxl(i,iface)
+         ul = auxr(i-1,iface)
+
+         qrr = ql(i,1)
+         qll = qr(i-1,1)
+
+         qhat = (qrr + qll)/2.d0
+
+         uavg = (ur+ul)/2.d0
+         tau = 1e-8
+         uhat = sign(1.d0,uavg)*max(abs(uavg),tau) 
+
+         wave(i,1,1) = qrr-qll + (ur-ul)*qhat/uhat         
+         if (uavg .eq. 0) then
+C           # lim(uavg --> 0) uavg*W = (ur-ul)*qhat
+C           #             
+            amdq(i,1) = -ul*qhat
+            apdq(i,1) = ur*qhat
+            s(i,1) = uavg
+         else
+         endif
+         amdq(i,1) = min(uhat,0.d0)*wave(i,1,1)
+         apdq(i,1) = max(uhat,0.d0)*wave(i,1,1)
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_qstar.f

@@ -1,0 +1,70 @@
+      subroutine rpn2cons_cc(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
+     &                       auxl,auxr,wave,s,amdq,apdq)
+
+      implicit none
+
+      integer maxm, mbc,mwaves,meqn,maux, mx
+      integer ixy
+
+      double precision wave(1-mbc:maxm+mbc, meqn, mwaves)
+      double precision    s(1-mbc:maxm+mbc, mwaves)
+      double precision   ql(1-mbc:maxm+mbc, meqn)
+      double precision   qr(1-mbc:maxm+mbc, meqn)
+      double precision amdq(1-mbc:maxm+mbc, meqn)
+      double precision apdq(1-mbc:maxm+mbc, meqn)
+      double precision auxl(1-mbc:maxm+mbc,*)
+      double precision auxr(1-mbc:maxm+mbc,*)
+
+
+      integer i, iface
+      double precision qstar, qrr,qll,ur,ul, uhat
+
+
+      double precision dtcom, dxcom, dycom, tcom
+      integer icom, jcom
+
+      common/comxyt/dtcom,dxcom,dycom,tcom,icom,jcom
+
+C     # We need to impose flux continuity at the "zero-wave"
+C     # ur > 0 : us*qs - ul*qll = 0 -->  us = ur; qs = ul*qll/ur;
+C     # ur > 0 : us*qs - ul*qll = 0 -->  us = ur; qs = ul*qll/ur;
+
+
+      iface = ixy
+      do i = 2-mbc, mx+mbc
+
+          ur = auxl(i,iface)
+          ul = auxr(i-1,iface)
+
+          qrr = ql(i,1)
+          qll = qr(i-1,1)
+
+          uhat = (ur+ul)/2.d0
+
+c         # ur and ul have the same sign
+          if (ur .gt. 0 .and. ul .gt. 0) then
+              qstar = ul*qll/ur
+              wave(i,1,1) = qrr - qstar
+              s(i,1) = ur
+              amdq(i,1) = 0.d0
+              apdq(i,1) = ur*qrr - ul*qll
+          elseif (ur .lt. 0 .and. ul .lt. 0) then
+              qstar = ur*qrr/ul
+              wave(i,1,1) = (ur*qrr - ul*qll)/ur
+              s(i,1) = ur
+              amdq(i,1) = ur*qrr - ul*qll
+              apdq(i,1) = 0.d0
+          else
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2fwave.f

@@ -1,0 +1,51 @@
+      subroutine rpn2fwave(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
+     &			auxl,auxr,wave,s,amdq,apdq)
+      implicit none
+
+      integer maxm, mbc,mwaves,meqn,maux, mx
+      integer ixy
+
+      double precision wave(1-mbc:maxm+mbc, meqn, mwaves)
+      double precision    s(1-mbc:maxm+mbc, mwaves)
+      double precision   ql(1-mbc:maxm+mbc, meqn)
+      double precision   qr(1-mbc:maxm+mbc, meqn)
+      double precision amdq(1-mbc:maxm+mbc, meqn)
+      double precision apdq(1-mbc:maxm+mbc, meqn)
+      double precision auxl(1-mbc:maxm+mbc,*)
+      double precision auxr(1-mbc:maxm+mbc,*)
+
+
+      integer i, iface, m
+      double precision uhat,qll,qrr,ul,ur, tol
+
+
+      double precision dtcom, dxcom, dycom, tcom
+      integer icom, jcom
+
+      common/comxyt/dtcom,dxcom,dycom,tcom,icom,jcom
+
+      iface = ixy
+      do i = 2-mbc, mx+mbc
+
+         ur = auxl(i,iface)
+         ul = auxr(i-1,iface)
+
+         qrr = ql(i,1)
+         qll = qr(i-1,1)
+
+         uhat = (ul + ur)/2.d0
+
+         wave(i,1,1) = ur*qrr - ul*qll
+         s(i,1) = uhat
+         if (uhat .lt. 0) then
+            amdq(i,1) = wave(i,1,1)
+            apdq(i,1) = 0.d0
+         else
+            amdq(i,1) = 0.d0
+            apdq(i,1) = wave(i,1,1)
+         endif
+      enddo
+
+
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/tag4coarsening.f

@@ -1,0 +1,67 @@
+      subroutine clawpack46_tag4coarsening(mx,my,mbc,meqn,
+     &      xlower,ylower,dx,dy, blockno, q0, q1, q2, q3,
+     &      coarsen_threshold, tag_patch)
+      implicit none
+
+      integer mx,my, mbc, meqn, tag_patch
+      integer blockno
+      double precision xlower(0:3), ylower(0:3), dx, dy
+      double precision coarsen_threshold
+      double precision q0(1-mbc:mx+mbc,1-mbc:my+mbc,meqn)
+      double precision q1(1-mbc:mx+mbc,1-mbc:my+mbc,meqn)
+      double precision q2(1-mbc:mx+mbc,1-mbc:my+mbc,meqn)
+      double precision q3(1-mbc:mx+mbc,1-mbc:my+mbc,meqn)
+
+      integer i,j, mq
+      double precision qmin, qmax
+
+c     # Assume that we will coarsen a family unless we find a grid
+c     # that doesn't pass the coarsening test.
+      tag_patch = 1
+      mq = 1
+      qmin = q0(1,1,mq)
+      qmax = q0(1,1,mq)
+
+c     # If we find that (qmax-qmin > coarsen_threshold) on any
+c     # grid, we return immediately, since the family will then
+c     # not be coarsened.
+
+      call user46_get_minmax(mx,my,mbc,meqn,mq,q0,qmin,qmax,
+     &      coarsen_threshold,tag_patch)
+      if (tag_patch == 0) return
+
+      call user46_get_minmax(mx,my,mbc,meqn,mq,q1,qmin,qmax,
+     &      coarsen_threshold,tag_patch)
+      if (tag_patch == 0) return
+
+      call user46_get_minmax(mx,my,mbc,meqn,mq,q2,qmin,qmax,
+     &      coarsen_threshold,tag_patch)
+      if (tag_patch == 0) return
+
+      call user46_get_minmax(mx,my,mbc,meqn,mq,q3,qmin,qmax,
+     &      coarsen_threshold,tag_patch)
+
+      end
+
+      subroutine user46_get_minmax(mx,my,mbc,meqn,mq,q,
+     &      qmin,qmax,coarsen_threshold,tag_patch)
+
+      implicit none
+      integer mx,my,mbc,meqn,mq,tag_patch
+      double precision coarsen_threshold
+      double precision qmin,qmax
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/tag4refinement.f

@@ -1,0 +1,30 @@
+      subroutine clawpack46_tag4refinement(mx,my,mbc,
+     &      meqn, xlower,ylower,dx,dy,blockno,
+     &      q, tag_threshold, init_flag,tag_patch)
+      implicit none
+
+      integer mx,my, mbc, meqn, tag_patch, init_flag
+      integer blockno
+      double precision xlower, ylower, dx, dy
+      double precision tag_threshold
+      double precision q(1-mbc:mx+mbc,1-mbc:my+mbc,meqn)
+
+      integer i,j, mq
+      double precision qmin, qmax
+
+      tag_patch = 0
+
+c     # We check only internal cells, because there is a problem with the
+c     # corner values on the cubed sphere. Possible bad values in (unused)
+c     # corners at block corners are triggering refinement where three blocks meet.
+      mq = 1
+      do j = 1,my
+         do i = 1,mx
+            if (q(i,j,mq) .gt. tag_threshold) then
+               tag_patch = 1
+               return
+            endif
+         enddo
+      enddo
+
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/1dadv/driver.f
(clawpack/advection/swirl - conservative) Add 1d reference solution that uses advection+source term

@@ -1,0 +1,32 @@
+      program driver
+c
+c  Generic driver routine for claw1ez
+c
+c  Author: Randall J. LeVeque
+c  Version of March, 1999 --  CLAWPACK Version 4.0
+c
+c
+      implicit double precision (a-h,o-z)
+
+c     # set parameters for maximum array sizes used in declarations
+c     # these must be increased for larger problems.
+c
+c
+      parameter (maxmx =   4000)
+      parameter (mwork =   122864)
+
+      parameter (mbc = 2)
+      parameter (meqn = 1)
+      parameter (mwaves = 1)
+      parameter (maux = 2)
+c
+      dimension q(1-mbc:maxmx+mbc, meqn)
+      dimension  aux(1-mbc:maxmx+mbc,maux)   !# dummy variable since no aux arrays used
+      dimension work(mwork)
+      dimension mthlim(mwaves)
+
+      call claw1ez(maxmx,meqn,mwaves,mbc,maux,mwork,mthlim,
+     &             q,work,aux)
+
+      stop 
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/1dadv/lib/inlinelimiter.f

@@ -1,0 +1,108 @@
+c
+c
+c     =====================================================
+      subroutine limiter(maxm,meqn,mwaves,mbc,mx,wave,s,mthlim)
+c     =====================================================
+c
+c     # Apply a limiter to the waves.  
+c
+c     # Version of December, 2002.
+c     # Modified from the original CLAWPACK routine to eliminate calls 
+c     # to philim.  Since philim was called for every wave at each cell
+c     # interface, this was adding substantial overhead in some cases.
+c
+c     # The limiter is computed by comparing the 2-norm of each wave with
+c     # the projection of the wave from the interface to the left or
+c     # right onto the current wave.  For a linear system this would
+c     # correspond to comparing the norms of the two waves.  For a 
+c     # nonlinear problem the eigenvectors are not colinear and so the 
+c     # projection is needed to provide more limiting in the case where the
+c     # neighboring wave has large norm but points in a different direction
+c     # in phase space.
+c
+c     # The specific limiter used in each family is determined by the
+c     # value of the corresponding element of the array mthlim.
+c     # Note that a different limiter may be used in each wave family.
+c
+c     # dotl and dotr denote the inner product of wave with the wave to
+c     # the left or right.  The norm of the projections onto the wave are then
+c     # given by dotl/wnorm2 and dotr/wnorm2, where wnorm2 is the 2-norm
+c     # of wave.
+c
+      implicit real*8(a-h,o-z)
+      dimension mthlim(mwaves)
+      dimension wave(1-mbc:maxm+mbc, meqn, mwaves)
+      dimension    s(1-mbc:maxm+mbc, mwaves)
+c
+c
+      do 200 mw=1,mwaves
+         if (mthlim(mw) .eq. 0) go to 200
+         dotr = 0.d0
+         do 190 i = 0, mx+1
+            wnorm2 = 0.d0
+            dotl = dotr
+            dotr = 0.d0
+            do 5 m=1,meqn
+               wnorm2 = wnorm2 + wave(i,m,mw)**2
+               dotr = dotr + wave(i,m,mw)*wave(i+1,m,mw)
+    5          continue
+            if (i.eq.0) go to 190
+            if (wnorm2.eq.0.d0) go to 190
+c
+            if (s(i,mw) .gt. 0.d0) then
+                r = dotl / wnorm2
+              else
+                r = dotr / wnorm2
+              endif
+c
+            go to (10,20,30,40,50) mthlim(mw)
+c
+   10       continue
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/1dadv/qinit.f

@@ -1,0 +1,20 @@
+c
+c
+c =========================================================
+       subroutine qinit(maxmx,meqn,mbc,mx,xlower,dx,q,maux,aux)
+c =========================================================
+c
+c     # Set initial conditions for q.
+c
+      implicit double precision (a-h,o-z)
+      dimension q(1-mbc:maxmx+mbc, meqn)
+      dimension aux(1-mbc:maxmx+mbc, *)
+c
+      pi = 4.d0*datan(1.d0)
+c
+      do 150 i=1,mx
+	      q(i,1) = 1.0
+  150 continue
+c
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/1dadv/setaux.f

@@ -1,0 +1,34 @@
+      subroutine setaux(maxmx,mbc,mx,xlower,dx,maux,aux)
+
+      implicit double precision (a-h,o-z)
+      dimension aux(1-mbc:maxmx+mbc, *)
+
+      double precision xe,xc, pi, dx1, dx2
+      integer i
+
+      common /compi/ pi
+
+
+      do i = 1-mbc,mx+mbc
+c          x = xlower + (i-0.5)*dx
+          xe = xlower + (i-1)*dx
+          xc = xlower + (i-0.5)*dx
+
+          aux(i,1) = cos(2*pi*xe)
+          aux(i,2) = -2*pi*sin(2*pi*xc)
+
+c          aux(i,1) = 0.1d0*sin(2*pi*xe)*sin(16*pi*xe)
+c          dx1 =  2*pi*cos(2*pi*xc)*sin(16*pi*xc)
+c          dx2 = 16*pi*sin(2*pi*xc)*cos(16*pi*xc)
+c          aux(i,2) = 0.1*(dx1 + dx2)
+
+c          if (x .le. 0.5d0) then
+c              aux(i,1) = -0.5d0
+c          else
+c              aux(i,1) = 0.5d0
+c          endif
+      enddo
+
+c
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/1dadv/setprob.f

@@ -1,0 +1,11 @@
+      subroutine setprob
+      implicit double precision (a-h,o-z)
+
+      double precision pi
+
+      common /compi/ pi
+
+      pi = 4.d0*atan(1.d0)
+
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/swirlcons_user.cpp
(clawpack/advection/swirl (cons)) Update to conservative application

@@ -39,6 +39,7 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 void swirlcons_link_solvers(fclaw2d_global_t *glob)
 {
     fclaw2d_vtable_t *vt = fclaw2d_vt();
+    fc2d_clawpack46_options_t *clawopt = fc2d_clawpack46_get_options(glob);
 
     vt->problem_setup = &swirlcons_problem_setup;  /* Version-independent */
 
@@ -48,29 +49,29 @@ void swirlcons_link_solvers(fclaw2d_global_t *glob)
         fc2d_clawpack46_vtable_t *clawpack46_vt = fc2d_clawpack46_vt();
         clawpack46_vt->qinit     = &CLAWPACK46_QINIT;
         clawpack46_vt->setaux    = &CLAWPACK46_SETAUX;
-        clawpack46_vt->b4step2   = &CLAWPACK46_B4STEP2;
-        clawpack46_vt->bc2       = &SWIRLCONS_BC2;
+        //clawpack46_vt->b4step2   = &CLAWPACK46_B4STEP2;
+        //clawpack46_vt->bc2       = &SWIRLCONS_BC2;
         if (user->cons_rp)
         {
-            clawpack46_vt->rpn2      = &RPN2CONS_CC;
-            clawpack46_vt->rpt2      = &RPT2CONS_CC;
-            clawpack46_vt->b4step2   = &SWIRLCONS_B4STEP2;
+            clawopt->use_fwaves = 0;
+            clawpack46_vt->rpn2      = &RPN2CONS_CC;                                
+        }
+        else if (clawopt->use_fwaves)
+        {
+            clawpack46_vt->rpn2      = &RPN2FWAVE;                
         }
         else
         {
-            clawpack46_vt->rpn2      = &CLAWPACK46_RPN2ADV;
-            clawpack46_vt->rpt2      = &CLAWPACK46_RPT2ADV;
-            clawpack46_vt->b4step2   = &CLAWPACK46_B4STEP2;
+            clawopt->use_fwaves = 0;
+            clawpack46_vt->rpn2      = &RPN2CONS;                                
         }
+        clawpack46_vt->rpt2      = &RPT2CONS_CC;
+        clawpack46_vt->b4step2   = &SWIRLCONS_B4STEP2;
     }
     else if (user->claw_version == 5)
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/swirlcons_user.h

@@ -63,6 +63,21 @@ void RPN2CONS_CC(const int* ixy,const int* maxm, const int* meqn, const int* mwa
                  double auxl[], double auxr[], double wave[],
                  double s[], double amdq[], double apdq[]);
 
+
+#define RPN2CONS FCLAW_F77_FUNC(rpn2cons,RPN2CONS)
+void RPN2CONS(const int* ixy,const int* maxm, const int* meqn, const int* mwaves,
+              const int* mbc,const int* mx, double ql[], double qr[],
+              double auxl[], double auxr[], double wave[],
+              double s[], double amdq[], double apdq[]);
+
+#define RPN2FWAVE FCLAW_F77_FUNC(rpn2fwave, RPN2FWAVE)
+void RPN2FWAVE(const int* ixy, const int* maxm, const int* meqn, const int* mwaves,
+               const int* mbc, const int* mx, double ql[], double qr[],
+               double auxl[], double auxr[], double fwave[],
+               double s[], double amdq[], double apdq[]);
+
+
+
 #define RPT2CONS_CC FCLAW_F77_FUNC(rpt2cons_cc, RPT2CONS_CC)
 void RPT2CONS_CC(const int* ixy, const int* maxm, const int* meqn, const int* mwaves,
                  const int* mbc, const int* mx, double ql[], double qr[],

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/qinit.f

@@ -1,28 +1,27 @@
       subroutine clawpack46_qinit(maxmx,maxmy, meqn,mbc,mx,my,
      &      xlower,ylower,dx,dy,q,maux,aux)
-       implicit none
+      implicit none
 
-       integer meqn, mbc, mx, my, maux, maxmx, maxmy
-       double precision xlower, ylower, dx, dy
-       double precision q(1-mbc:maxmx+mbc, 1-mbc:maxmy+mbc, meqn)
-       double precision aux(1-mbc:maxmx+mbc, 1-mbc:maxmy+mbc, maux)
+      integer meqn, mbc, mx, my, maux, maxmx, maxmy
+      double precision xlower, ylower, dx, dy
+      double precision q(1-mbc:maxmx+mbc, 1-mbc:maxmy+mbc, meqn)
+      double precision aux(1-mbc:maxmx+mbc, 1-mbc:maxmy+mbc, maux)
 
-       integer i, j, mq
-       double precision xc,yc
+      integer i, j, mq
+      double precision xc,yc
+      do mq = 1,meqn
+        do i = 1-mbc,mx+mbc
+           xc = xlower + (i-0.5d0)*dx
+           do j = 1-mbc,my+mbc
+              yc = ylower + (j-0.5d0)*dy
+              if (abs(xc-0.5) .lt. 0.15d0) then
+                 q(i,j,mq) = 1.d0
+              else
+                 q(i,j,mq) = 1.0
+              endif
+           enddo
+        enddo
+      enddo
 
-       do mq = 1,meqn
-          do i = 1-mbc,mx+mbc
-             xc = xlower + (i-0.5d0)*dx
-             do j = 1-mbc,my+mbc
-                yc = ylower + (j-0.5d0)*dy
-                if (xc .lt. 0.5d0) then
-                   q(i,j,mq) = 1.d0
-                else
-                   q(i,j,mq) = 0.5
-                endif
-             enddo
-          enddo
-       enddo
-
-       return
-       end
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/swirlcons_b4step2.f

@@ -18,8 +18,8 @@ c        # setaux should be used for all time.
          return
       endif
 
-C       vt = cos(pi2*(time+dt/2.d0)/tperiod)
-      vt = 1.0
+      vt = cos(pi2*(time+dt/2.d0)/tperiod)
+c      vt = 1.0
 
       do i = 1-mbc,mx+mbc
          do j = 1-mbc,my+mbc
@@ -52,26 +52,34 @@ C             aux(i,j,2) = 0.0
       double precision xp,yp,pi
       common /compi/ pi
 
-C       u = -2*((sin(pi*xp))**2 * sin(pi*yp) * cos(pi*yp)) + 2
-C       if(xp .gt. 0) then
-C             ucc = exp(-(xp-0.5)**2/(2*0.1**2))
-C       else
-C             ucc = 0
-C       endif
+c      ucc = 2*((sin(pi*xp))**2 * sin(pi*yp) * cos(pi*yp))
+      ucc = cos(2*pi*xp)
 
-C       if((xp .gt. 0.5 .and. xp .lt. 1) .or. (xp .lt. 0)) then
-C             ucc = 1.0
-C       else
-C             ucc = -1.0
-C       endif
+c      ucc = 0.1*sin(2*pi*xp)*sin(16*pi*xp)
+c      if (xp .le. 0.5d0) then
+c            ucc = -.5d0
+c      else
+c            ucc = .50
+c      endif
+c      if(xp .gt. 0) then
+c            ucc = exp(-(xp-0.5)**2/(2*0.1**2))
+c      else
+c            ucc = 0
+c      endif
 
-      if((xp .gt. 0.5 .and. xp .lt. 0.75)) then
-            ucc = 0.1
-      else if((xp .gt. 0.25 .and. xp .lt. 0.5)) then
-            ucc = -0.1
-      else
-            ucc = 0.0
-      endif
+C      if (xp .gt. 0.5) then
+C            ucc = 0.5
+C      else
+C            ucc = 1.0
+C      endif
+
+c      if((xp .gt. 0.5 .and. xp .lt. 0.75)) then
+c            ucc = 0.1
+c      else if((xp .gt. 0.25 .and. xp .lt. 0.5)) then
+c            ucc = -0.1
+c      else
+c            ucc = 0.0
+c      endif
 
       return
       end
@@ -82,7 +90,7 @@ C       endif
       double precision xp,yp,pi
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/afterframe.m
(clawpack/advection/swirl (cons)) Add plotting routines to see under/over shoots

@@ -1,28 +1,79 @@
-s = 1e-2;
-axis([-s 1+s -s 1+s])
-daspect([1 1 1]);
-axis on;
+if PlotType == 1
 
-yrbcolormap;
-colorbar;
-showpatchborders(1:10);
-caxis([0,5])
-qlo = 0;
-qhi = 1;
-under_label = sprintf('0 - %7.1e',qlo-qmin);
-over_label = sprintf('1 + %7.1e',qmax-qhi);
-fprintf('%6s %12s\n','qmin',under_label);
-fprintf('%6s %12s\n\n','qmax',over_label);
+    s = 0;
+    axis([-s 1+s -s 1+s])
+    daspect([1 1 1]);
+    axis on;
+    
+    hidepatchborders;
+    
+    if (ShowUnderOverShoots == 1)
+        qlo = 0;
+        qhi = 1;
+        under_label = sprintf('%3.1f - %7.1e',qlo,qlo-qmin);
+        over_label = sprintf('%3.1f + %7.1e',qhi,qmax-qhi);
+        fprintf('%-10s %12.4e\n','qmin',qmin);
+        fprintf('%-10s %12.4e\n','qmax',qmax);
+        colorbar_underover(under_label,over_label);
+    else                 
+        yrbcolormap;
+        colorbar;
+        caxis([0 1]);
+    end
+    showpatchborders(1:10);
+    view(2);
+elseif PlotType == 4
+    hold on;
+    dir = './1dadv/';
+    dim = 1;
+    [amrdata1d,t1d] = readamrdata(dim,Frame,dir);
+    [q1d,x1d] = plotframe1ez(amrdata1d,mq,'b.-');    
 
-
-if (ShowUnderOverShoots)
-    qlo = 0;
-    qhi = 1;
-    colorbar_underover(under_label,over_label);
+    % Plot zeros of velocity function
+    vel_case = 1;
+    yl = [-5,25];
+    switch vel_case
+        case 1
+            % Case 1 : u= cos(2*pi*x)
+            u = @(x) cos(2*pi*x);
+            plot([0.25,0.25],yl,'k--');
+            plot([0.75,0.75],yl,'k--');
+        case 2
+            % Case 2 : u = cos(2*pi*x) + 0.5
+            u = @(x) cos(2*pi*x) + 0.5;
+            z = acos(-0.5)/(2*pi);
+            plot([z,z],yl,'k--');
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/map1d.m

@@ -1,0 +1,54 @@
+function [r,q] = map1d(xgrid,ygrid,qgrid)
+
+%  MAP1D is a user-defined function to create line plots from 2d/3d data
+%
+%      MAP1D is a user-defined function called by PLOTCLAW2 or PLOTCLAW3 to
+%      convert 2d or 3d data into 1d data for a line plot.
+%
+%      For 2d data, the MAP1D function has the form
+%
+%           [r,q] = map1d(xgrid,ygrid,qgrid)
+%
+%      where XGRID, YGRID, and QGRID are MxN arrays of grid data, and R and
+%      Q are 1d column vectors of user-defined length.
+%
+%      For 3d data, the MAP1D function has the form
+%
+%           [r,q] = map1d(xgrid,ygrid,zgrid, qgrid)
+%
+%      where XGRID, YGRID, ZGRID and QGRID are MxNx{ arrays of grid data,
+%      R and Q are 1d column vectors of user-defined length.
+%
+%      This function will be called from PLOTCLAW2 or PLOTCLAW3 if
+%      'UserMap1d' = 1.   This plotting parameter can be set in the file
+%      SETPLOT2 or SETPLOT3.
+%
+%      Example :
+%
+%             % To plot q as a function of x only ;
+%             function [r,q] = map1d(xgrid,ygrid,qgrid)
+%
+%             [m,n] = size(xgrid);
+%             r = reshape(xgrid,m*n,1);
+%             q = reshape(qgrid,m*n,1);
+%
+%      Example :
+%
+%           % Plot data interpolated to grid diagonal on [0,1]x[0,1]x[0,1]
+%           % grid
+%           function [r,q] = map1d(xgrid,ygrid,zgrid,qgrid)
+%
+%           r = linspace(0,0.1,1)*sqrt(3);
+%           q = interp3(xgrid,ygrid,zgrid,qgrid,r);
+%
+%
+%      NOTE : This help file can also be found by typing 'map1d_help' at
+%      the Matlab prompt
+%
+%
+%  See also SETPLOTSTYLE, PLOTFRAME1EZ, GETLEGENDINFO, SETPLOT.
+
+r = xgrid(:);
+q = qgrid(:);
+
+end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/setplot2.m

@@ -3,7 +3,7 @@ OutputDir = './';            % Default (reassign them here anyway)
 
 ForestClaw = 1;     % Plot using ForestClaw preferences.
 
-PlotType = 1;                % type of plot to produce:
+PlotType = 4;                % type of plot to produce:
 			     % 1 = pseudo-color (pcolor)
 			     % 2 = contour
 			     % 3 = Schlieren
@@ -31,6 +31,8 @@ ContourValues = [];
 
 %---------------------------------
 
+UserMap1d = 1;
+ScatterStyle = setplotstyle('ko','go','yo','mx','bo','ro');
 
-ShowUnderOverShoots = 0;
-PlotParallelPartitions=0;
+ShowUnderOverShoots = 1;
+PlotParallelPartitions = 0;

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/underover.m

@@ -1,0 +1,70 @@
+function uo = underover()
+%UNDEROVER returns structure for assigning colors to under/overshoots
+%
+%    UO = UNDEROVER.  To have special highlighting for under/over shoot
+%    values, the user should first set the ShowUnderOverShoots flag in
+%    SETPLOT2 or the SETPLOT3 files.
+%    Then, make a local copy of this file and supply values for the the
+%    following fields in the UO structure :
+%
+%             uo.color_under     the rgb triple used to color undershoots
+%             uo.color_over      the rgb triple used to color overshoots
+%             uo.color_nan       the color to use for NAN values
+%             uo.value_lower     lower value of exact solution
+%             uo.value_upper     upper value of exact solution
+%             uo.tol             tolerance for under/over shoots
+%             uo.colormap        colormap for values in [qlow-tol,qhi+tol];
+%
+%    Values are considered in the overshoot/undershoot region if they satisfy
+%
+%                    q < value_lower - tol      (undershoot)
+%    or
+%                    q > value_upper + tol      (overshoot)
+%
+%    Values not in the under/overshoot region are linearly scaled into the
+%    colormap.  Values that lie in [qlow-tol,qlow] are clamped to qlow, and
+%    values in [qhi, qhi + tol] are clamped to qhi.
+%
+%    The under/over shoot colorbar only takes affect if you have set
+%    'ShowUnderOverShoots' in SETPLOT2.m.
+%
+%    Example :  First, create underover structure :
+%    function uo = underover()
+%
+%    colormap('default');
+%    cm = colormap;
+%
+%    uo = struct('color_under',[0 1 1],...  % cyan
+%                'color_over',[1 0 1],...   % magenta
+%                'color_nan',[1 1 1],...    % white
+%                'value_lower',0, ...       % theoretical minimum
+%    	         'value_upper',1,...        % theoretical maximum
+%    	         'tol',1e-4,...             % acceptable numerical over/undershoot
+%    	         'colormap',cm);            % Colors for non under/over shoot values
+%
+%
+%    Example : To make sure that your under/over shoot color map is used :
+%    % In local copy of 'setplot2.m'
+%    ShowUnderOverShoots = 1;
+%
+%    % In local copy of 'afterframe.m'
+%    qlo = 0;
+%    qhi = 1;
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/Riemann_solvers/rpn2fwave.f
(clawpack/advection/swirl (cons)) Move Riemann solvers to user_4.6

@@ -1,85 +1,0 @@
-      subroutine rpn2(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
-     &			auxl,auxr,wave,s,amdq,apdq)
-      implicit none
-
-      integer maxm, mbc,mwaves,meqn,maux, mx
-      integer ixy
-
-      double precision wave(1-mbc:maxm+mbc, meqn, mwaves)
-      double precision    s(1-mbc:maxm+mbc, mwaves)
-      double precision   ql(1-mbc:maxm+mbc, meqn)
-      double precision   qr(1-mbc:maxm+mbc, meqn)
-      double precision amdq(1-mbc:maxm+mbc, meqn)
-      double precision apdq(1-mbc:maxm+mbc, meqn)
-      double precision auxl(1-mbc:maxm+mbc,*)
-      double precision auxr(1-mbc:maxm+mbc,*)
-
-
-      integer i, iface, m, mq, mw
-      double precision ustar, fi,fim
-      double precision nv(3), ul(3), ur(3), vl,vr,v
-      double precision fstar
-
-
-      double precision dtcom, dxcom, dycom, tcom
-      integer icom, jcom
-      integer vflag, get_vflag
-
-      common/comxyt/dtcom,dxcom,dycom,tcom,icom,jcom
-
-      vflag = get_vflag()
-      if (vflag .ne. 3) then
-         write(6,*) 'rpn3fwave : You must supply cell ',
-     &         'centered velocities for this Riemann ',
-     &         'solver;  set vflag = 3 in setprob.data'
-         write(6,100) 'vflag', vflag
-         stop
-  100    format(A10,I3)
-      endif
-
-
-
-      iface = ixy
-      do i = 2-mbc, mx+mbc
-
-         do mq = 1,meqn
-            do mw = 1,mwaves
-               wave(i,mq,mw) = 0
-            enddo
-         enddo
-
-         do mq = 1,meqn
-c           # Get normal (times area) at this edge
-            vl = 0
-            vr = 0
-            do m = 1,3
-               nv(m) = auxl(i,1+(iface-1)*3+m)
-               ul(m) = auxr(i-1,10+(mq-1)*3 + m)
-               ur(m) = auxl(i,10 + (mq-1)*3 + m)
-
-c              # Use the same normal for both left and right
-c              # cell centered velocities
-               vl = vl + nv(m)*ul(m)
-               vr = vr + nv(m)*ur(m)
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpt2cons_cc.f

@@ -1,0 +1,42 @@
+      subroutine rpt2cons_cc(ixy,maxm,meqn,mwaves,mbc,mx,
+     &                       ql,qr,aux1,aux2,aux3,imp,asdq,
+     &                       bmasdq,bpasdq)
+      
+      implicit none
+
+      integer ixy, icoor, maxm, meqn,mwaves,mbc,mx,maux,imp
+
+      double precision     ql(1-mbc:maxm+mbc, meqn)
+      double precision     qr(1-mbc:maxm+mbc, meqn)
+      double precision   asdq(1-mbc:maxm+mbc, meqn)
+      double precision bmasdq(1-mbc:maxm+mbc, meqn)
+      double precision bpasdq(1-mbc:maxm+mbc, meqn)
+      double precision   aux1(1-mbc:maxm+mbc, *)
+      double precision   aux2(1-mbc:maxm+mbc, *)
+      double precision   aux3(1-mbc:maxm+mbc, *)
+
+
+      integer iuv, iface,i,j,i1
+      double precision vl,vr,vhat
+
+      iface = 3-ixy
+      do i = 2-mbc, mx+mbc
+         i1 = i-2+imp    !#  =  i-1 for amdq,  i for apdq
+
+c        # Lower face (at imp faces)
+         vr = aux2(i1,iface)
+         vl = aux1(i1,iface)
+         vhat = (vl+vr)/2.d0
+
+         bmasdq(i,1) = min(vhat,0.d0)*asdq(i,1)
+
+c        # Upper face (at imp faces)
+         vr = aux3(i1,iface)
+         vl = aux2(i1,iface)
+         vhat = (vl + vr)/2.d0
+         bpasdq(i,1) = max(vhat,0.d0)*asdq(i,1)
+      enddo
+
+
+      return
+      end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/notes/fwave.tex
(clawpack/advection/swirl (cons)) Adding some written notes on conservative approach

\documentclass{article}

\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{cite}

\begin{document}
\paragraph{Scalar approach to conservation}
In \cite{le:2002}, LeVeque describes a solution to the conservation law
\begin{align}
q_t + (uq)_x &= 0 \\
\end{align}
where $u = u(x)$ is a spatially varying velocity field, and $q = q(x,t)$ is a conserved 
tracer in  the flow field.  At a mesh cell interface, we have two velocities, 
$u_r$ and $u_\ell$, which will not in general be equal.  As a result, satisfying a jump
condition 
\begin{equation}
u_r q_r - u_\ell q_\ell = s(q_r - q_\ell)
\end{equation}
for some speed $s$ will not in general be possible.   Assuming that $u_r > 0$The correct solution actually has two waves, one moving at speed 0 and a second at speed $u_r$.  To determine the 
intermediate state $q^*$ we impose flux continuity across the speed 0 wave.
\begin{equation}
\end{equation}

\paragraph{2-wave decomposition approach.}
Consider the following conservation law for the conservative form of the scalar advection equation.
\begin{align}
q_t + (uq)_x &= 0 \\
u_t + (au)_x & = 0
\end{align}
for conserved quantity $q(x,t)$, and  $u = u(x,t)$ is a prescribed velocity field.  We will only 
interested in the case $a = 0$ so that $u = u(x)$ depends on space only.

We set $\mathbf q = (q,u)$ and compute the flux Jacobian $f'(\mathbf q)$
\begin{equation}
f'(\mathbf q) = \left(\begin{array}{rr} u & q \\ 0 & a\end{array}\right).
\end{equation}
The eigenvalue and eigenvector pairs for $f'(\mathbf q)$ are given by
\begin{equation}
\lambda_1 = a,  \quad  {\mathbf r}^1 = \left(\begin{array}{c} q \\ a-u\end{array}\right) \qquad \mbox{and} \qquad
\lambda_2 = u,  \quad  {\mathbf r}^2 = \left(\begin{array}{c} 1 \\ 0\end{array}\right)
\end{equation}
where we have assumed that $a < u$.  For $a \ne u$, the eigenvectors are linear independent, and so the system is strictly hyperbolic.  Furthermore, both fields satisfy $\nabla \lambda_p(\mathbf q) \cdot r^p(\mathbf q) = 0$, and so both waves are linearly degenerate.

The first wave satisfies $\lambda_1(\mathbf q_\ell) = \lambda_1(\mathbf q_r)$ and so is only supports contact discontinuities with jumps proportional to $r^1$ and neither wave supports rarefaction
waves.  One consequence of this is that if $u_\ell < u_r$, the solution in the classic sense does not exist.

We write the problem in quasi-linear form as 
\begin{equation}
\mathbf q_t + f'(\mathbf q) \mathbf q_x = 0
\end{equation}
or
\begin{equation}
\mathbf q_t + A(\mathbf q) \mathbf q_x = 0
\end{equation}
where $A(\mathbf q) \equiv f'(\mathbf q)$.

Let $R = R(\mathbf q)$ be a matrix of eigen-vectors, and solve
\begin{equation}
R \alpha = \mathbf \delta,
\end{equation}
where  $\mathbf \delta = \mathbf q_r - \mathbf q_\ell$ is a general jump in
left and right states.  Solving for $\alpha$, we get 
\begin{equation}
\left[\begin{array}{c} \alpha_1 \\ \alpha_2 \end{array}\right] = \frac{1}{u-a}
\left[
\begin{array}{rr}
0 & -1 \\ u-a & q
\end{array}
\right]
\left[
\begin{array}{r}
\delta_1 \\ \delta_2
\end{array}
\right]
\end{equation}
or 
\begin{equation}
\alpha_1(\mathbf q) = \frac{-(u_r - u_\ell)}{u-a}, \qquad 
\alpha_2(\mathbf q) = q_r - q_\ell + \frac{u_r - u_l}{u-a} q
\end{equation}
To determine which values to use for $\mathbf q$, we seek values $\widehat{\mathbf q}$ 
so that 
\begin{align}
f(\mathbf q_r) - f(\mathbf q_\ell) & = A(\widehat{\mathbf q})(\mathbf q_r - \mathbf q_\ell) 
\end{align}
is satisfied.  It is straightforward to show that 
setting $\widehat{\mathbf q} = (\widehat{q},\widehat{u})$ to
\begin{equation}
\widehat{q} = \frac{q_\ell + q_r}{2} 
\qquad \mbox{and} \qquad
\widehat{u} = \frac{u_\ell + u_r}{2}
\end{equation}
we can guarantee the conservation of the method.
Using these values, we can then write
down the solution to the Riemann problem as a two-wave
solution separated by constant states

The Riemann solution to the quasi-linear problem at at interface $x=0$ can be written as 
\begin{equation}
\mathbf q(x,t) = \left\{\begin{array}{ll}
(q_\ell, u_\ell) &   x/t < a \\ 
(q^*, u^*)       &   a   \le x/t < \widehat{u} \\ 
(q_r, u_r)       &   \widehat{u} \le x/t
\end{array}
\right.
\end{equation}
where
\begin{align}
\mathbf q^* 
& = \mathbf q_\ell + \alpha_1(\widehat{\mathbf q}) {\mathbf r}^1 \\
& = \mathbf q_r - \alpha_2(\widehat{\mathbf q}) {\mathbf r}^2
\end{align}
Solving for $q^*$ and $u^*$, we get
\begin{equation}
q^*  = q_\ell - \frac{u_r - u_\ell}{\widehat{u}-a} \widehat{q}, \qquad
u^*  = u_\ell - \frac{u_r - u_\ell}{\widehat{u}-a} (a - \widehat{u}) = u_r
\end{equation}
Waves, speeds and fluctuations ae given by
\begin{align}
\mathcal W_1 & \equiv \alpha_1 \mathbf r^1 = - \frac{u_r - u_\ell}{\widehat{u}-a} 
\left(\begin{array}{c} \widehat{q} \\ a - \widehat{u}\end{array}\right) 
= (u_r - u_\ell)
\left(\begin{array}{c} \widehat{q}/\varepsilon \\ 
1 \end{array}\right) \\
\mathcal W_2 & \equiv \alpha_2 \mathbf r^2 = 
\left(q_r - q_\ell + \frac{u_r - u_\ell}{\widehat{u}-a}\widehat{q} \right)
\left(\begin{array}{c} 1 \\ 0\end{array}\right) 
= \left(\begin{array}{c} q_r - q_\ell - (u_r - u_\ell) \widehat{q}/\varepsilon \\ 
0 \end{array}\right) 
\end{align}
where we have set $\varepsilon \equiv a - \widehat{u}$. 
The fluctuations are  defined as
\begin{align}
\mathcal A^- \Delta Q & \equiv \lambda_1 \alpha_1(\widehat{q}) \mathbf r^1 
= a \mathcal W_1 \\
\mathcal A^+ \Delta Q & \equiv \lambda_2 \alpha_2(\widehat{q}) \mathbf r^2 
= \widehat{u} \mathcal W_2
\end{align}
Our choice of $\widehat{\mathbf q}$ guarantees that the following conservation property is satisfied
\begin{equation}
f(\mathbf q_r) - f(\mathbf q_\ell) = \mathcal A^- \Delta Q + \mathcal A^+ \Delta Q
\end{equation}
For a spatially varying velocity field, we have $a = 0$.  In this case, $\mathcal A^- \Delta Q = 0$, $\varepsilon = -\widehat{u}$, and the waves are not defined if $u_\ell$ and $u_r$ are equal in magnitude, but opposite in sign. However, the fluctuations are still well defined.

\begin{align}
\lim_{\varepsilon \rightarrow 0} \mathcal A^+ \Delta Q
& = \lim_{\varepsilon \rightarrow 0} -\varepsilon \mathcal W_2 \\
& = \lim_{\varepsilon \rightarrow 0} \left[-\varepsilon(q_r - q_\ell) + (u_r - u_\ell) \widehat{q}\right] \\
& = (u_r - u_\ell) \widehat{q}
\end{align}
where, for convenience, we have ignored the second component of $\mathcal W_2$.

Following the above approach, we can also find waves for the case $\widehat{u} < a = 0$.   Combining these two cases, we get the following Clawpack Riemann solver.

\vspace{0.25cm}

\begin{minipage}{\textwidth}
\begin{verbatim}
# Riemann solver for wave decomposition approach
def rpn2(ql,qr,ul,ur):
    uavg = (ul + ur)/2        
    uhat = sign(uavg)*max([abs(uavg),tol])
    qhat = (ql + qr)/2
    wave = qr - ql + (ur-ul)*qhat/uhat
    if uavg == 0:
        amdq = -ul*qhat     # amdq+apdq = (ur-ul)*qhat
        apdq = ur*qhat
        s = uavg
    else:
        apdq = max([uhat,0])*wave
        amdq = min([uhat,0])*wave        
        s = uhat
    return wave,s,amdq,apdq
\end{verbatim}
\end{minipage}

\vspace{0.25cm}

\paragraph{The F-wave approach.} To correct this problem, we use the f-wave approach and
split the jump in the flux directly, avoiding the need to create waves $\mathcal W_p$.  
We decompose the jump in flux as
\begin{equation}
f(\mathbf q_r) - f(\mathbf q_\ell) = \beta_1 \mathbf r^1 + \beta_2\mathbf r^2
\end{equation}
where $\mathbf r^1$ and $\mathbf r^2$ are the eigenvectors from above, evaluated at the $\widehat{\mathbf q}$ state.   Solve this leads to

\begin{equation}
\left[\begin{array}{c} \beta_1 \\ \beta_2 \end{array}\right] = \frac{1}{u-a}
\left[
\begin{array}{rr}
0 & -1 \\ u-a & q
\end{array}
\right]
\left[
\begin{array}{r}
\delta_1 \\ \delta_2
\end{array}
\right]
\end{equation}
or 
\begin{equation}
\beta_1(\mathbf q) = \frac{-a(u_r - u_\ell)}{\widehat{u}-a}, \qquad 
\beta_2(\mathbf q) = u_rq_r - q_rq_\ell + \frac{a(u_r - u_l)}{\widehat{u}-a} \widehat{q}
\end{equation}
F-waves are then defined as 
\begin{align}
\mathcal Z_1 & \equiv \beta_1(\mathbf q) \mathbf r^1 = - \frac{a(u_r - u_\ell)}{\widehat{u}-a} 
\left(\begin{array}{c} \widehat{q} \\ a - \widehat{u}\end{array}\right) 
= a(u_r - u_\ell) \left(\begin{array}{c} \widehat{q}/\varepsilon \\ 1 \end{array}\right) \\
\mathcal Z_2 & \equiv \beta_2(\mathbf q) \mathbf r^2 =
\left(u_r q_r - u_\ell q_\ell + \frac{a(u_r - u_\ell)}{\widehat{u}-a}\widehat{q} \right)
\left(\begin{array}{c} 1 \\ 0\end{array}\right) 
= \left(\begin{array}{c} u_r q_r - q_r q_\ell - a(u_r - u_\ell) \widehat{q}/\varepsilon \\ 
0 \end{array}\right) 
\end{align}

As in the above, we set $a = 0$.  Following the f-wave approach, we set $\mathcal A^- \Delta Q = \mathcal Z_1$ and $\mathcal A^+ \Delta Q = \mathcal Z_2$ to get
\begin{align}
\mathcal A^- \Delta Q & = 0 \\
\mathcal A^+ \Delta Q & = u_r q_r - q_r q_\ell
\end{align}
As before, the fluctuations are well-defined for all $u_\ell$, $u_r$, but now the fwaves needed by the algorithm are defined as well and are given by
\begin{align}
\mathcal Z_1 & = 0 \\
\mathcal Z_2 & = u_r q_r - q_r q_\ell
\end{align}
The f-wave approach leads to the following Clawpack Riemann solver.

\vspace{0.25cm}

\begin{minipage}{\textwidth}
\begin{verbatim}
# Riemann solver for f-wave approach
def rpn2fwave(ql,qr,ul,ur):
    uhat = (ul + ur)/2.0
    fwave = ur*qr - ql*ul
    s = uhat
    if uhat < 0:
        amdq = wave
        apdq = 0
    else:
        amdq = 0
        apdq = wave
\end{verbatim}
\end{minipage}

\vspace{0.25cm}

\noindent
The advantage of this approach is that we don't have to make any artifical choices for the case $u_r \approx -u_\ell$.

\paragraph{numerical results.}  We test the two above approaches

\bibliographystyle{plain}
\bibliography{cons}

\end{document}

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/1dadv/setaux.f
(clawpack/advection/swirl(cons)) Updates to Riemann solvers"

@@ -14,13 +14,13 @@ c          x = xlower + (i-0.5)*dx
           xe = xlower + (i-1)*dx
           xc = xlower + (i-0.5)*dx
 
-          aux(i,1) = cos(2*pi*xe)
-          aux(i,2) = -2*pi*sin(2*pi*xc)
+c          aux(i,1) = cos(2*pi*xe)
+c          aux(i,2) = -2*pi*sin(2*pi*xc)
 
-c          aux(i,1) = 0.1d0*sin(2*pi*xe)*sin(16*pi*xe)
-c          dx1 =  2*pi*cos(2*pi*xc)*sin(16*pi*xc)
-c          dx2 = 16*pi*sin(2*pi*xc)*cos(16*pi*xc)
-c          aux(i,2) = 0.1*(dx1 + dx2)
+          aux(i,1) = 0.1d0*sin(2*pi*xe)*sin(16*pi*xe)
+          dx1 =  2*pi*cos(2*pi*xc)*sin(16*pi*xc)
+          dx2 = 16*pi*sin(2*pi*xc)*cos(16*pi*xc)
+          aux(i,2) = 0.1*(dx1 + dx2)
 
 c          if (x .le. 0.5d0) then
 c              aux(i,1) = -0.5d0

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/afterframe.m

@@ -30,7 +30,7 @@ elseif PlotType == 4
     [q1d,x1d] = plotframe1ez(amrdata1d,mq,'b.-');    
 
     % Plot zeros of velocity function
-    vel_case = 1;
+    vel_case = 5;
     yl = [-5,25];
     switch vel_case
         case 1

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/swirlcons_user.cpp

@@ -49,8 +49,6 @@ void swirlcons_link_solvers(fclaw2d_global_t *glob)
         fc2d_clawpack46_vtable_t *clawpack46_vt = fc2d_clawpack46_vt();
         clawpack46_vt->qinit     = &CLAWPACK46_QINIT;
         clawpack46_vt->setaux    = &CLAWPACK46_SETAUX;
-        //clawpack46_vt->b4step2   = &CLAWPACK46_B4STEP2;
-        //clawpack46_vt->bc2       = &SWIRLCONS_BC2;
         if (user->cons_rp)
         {
             clawopt->use_fwaves = 0;

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/qinit.f

@@ -15,9 +15,9 @@
            do j = 1-mbc,my+mbc
               yc = ylower + (j-0.5d0)*dy
               if (abs(xc-0.5) .lt. 0.15d0) then
-                 q(i,j,mq) = 1.d0
-              else
                  q(i,j,mq) = 1.0
+              else
+                 q(i,j,mq) = 0.0
               endif
            enddo
         enddo

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_new.f

@@ -18,6 +18,7 @@
 
       integer i, iface, m, mw, version
       double precision ur,ul,qrr,qll, uhat,qhat, tau,uavg
+      double precision a,b,c
 
 
       double precision dtcom, dxcom, dycom, tcom
@@ -43,21 +44,23 @@
          qhat = (qrr + qll)/2.d0
 
          uavg = (ur+ul)/2.d0
-         tau = 1e-8
-         uhat = sign(1.d0,uavg)*max(abs(uavg),tau) 
 
-         wave(i,1,1) = qrr-qll + (ur-ul)*qhat/uhat         
          if (uavg .eq. 0) then
-C           # lim(uavg --> 0) uavg*W = (ur-ul)*qhat
-C           #             
+            tau = 1e-12
+            uhat = sign(1.d0,uavg)*max(abs(uavg),tau) 
+         else
+            uhat = uavg
+         endif
+         wave(i,1,1) = qrr-qll + (ur-ul)*qhat/uhat         
+         s(i,1) = uhat
+
+         if (uavg .eq. 0) then
             amdq(i,1) = -ul*qhat
             apdq(i,1) = ur*qhat
-            s(i,1) = uavg
          else
+            amdq(i,1) = min(uhat,0.d0)*wave(i,1,1)
+            apdq(i,1) = max(uhat,0.d0)*wave(i,1,1)
          endif
-         amdq(i,1) = min(uhat,0.d0)*wave(i,1,1)
-         apdq(i,1) = max(uhat,0.d0)*wave(i,1,1)
-         s(i,1) = uhat
       enddo
 
       return

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/user_4.6/swirlcons_b4step2.f

@@ -49,11 +49,15 @@ C             aux(i,j,2) = 0.0
       double precision function ucc(xp,yp)
       implicit none
 
-      double precision xp,yp,pi
+      double precision xp,yp,pi, c,rp2
       common /compi/ pi
 
-c      ucc = 2*((sin(pi*xp))**2 * sin(pi*yp) * cos(pi*yp))
-      ucc = cos(2*pi*xp)
+      rp2 = (xp-0.5d0)**2 + (yp-0.5d0)**2
+      c = exp(-350.d0*rp2)
+
+      ucc = 2*((sin(pi*xp))**2 * sin(pi*yp) * cos(pi*yp))
+      ucc = ucc + c
+c      ucc = cos(2*pi*xp)
 
 c      ucc = 0.1*sin(2*pi*xp)*sin(16*pi*xp)
 c      if (xp .le. 0.5d0) then
@@ -87,11 +91,15 @@ c      endif
       double precision function vcc(xp,yp)
       implicit none
 
-      double precision xp,yp,pi
+      double precision xp,yp,pi, rp2, c
       common /compi/ pi
 
-c      vcc = -2*((sin(pi*yp))**2 * sin(pi*xp) * cos(pi*xp))
-      vcc = 0
+      rp2 = (xp-0.5d0)**2 + (yp-0.5d0)**2
+      c = exp(-350.d0*rp2)
+
+      vcc = -2*((sin(pi*yp))**2 * sin(pi*xp) * cos(pi*xp))
+      vcc = vcc + c
+c      vcc = 0
 
       return
       end

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/swirlcons.cpp
(clawpack/advection/swirl (cons)) Update to latest file names, headers

@@ -31,10 +31,7 @@
 #include <fclaw2d_clawpatch.h>
 
 #include <fc2d_clawpack46_options.h>
-#include <fc2d_clawpack5_options.h>
-
 #include <fc2d_clawpack46.h>
-#include <fc2d_clawpack5.h>
 
 static int s_user_options_package_id = -1;
 
@@ -45,7 +42,7 @@ swirlcons_register (user_options_t *user, sc_options_t * opt)
     sc_options_add_double (opt, 0, "period", &user->period, 4,
                            "Period of the flow field [4]");
 
-    sc_options_add_int (opt, 0, "claw-version", &user->claw_version, 5,
+    sc_options_add_int (opt, 0, "claw-version", &user->claw_version, 4,
                            "Clawpack_version (4 or 5) [5]");
 
     sc_options_add_bool (opt, 0, "cons-rp", &user->cons_rp, 1,
@@ -208,28 +205,17 @@ fclaw2d_domain_t* create_domain(sc_MPI_Comm mpicomm, fclaw_options_t* gparms)
 static
 void run_program(fclaw2d_global_t* glob)
 {
-    const user_options_t           *user_opt;
-
     /* ---------------------------------------------------------------
        Set domain data.
        --------------------------------------------------------------- */
     fclaw2d_domain_data_new(glob->domain);
-
-    user_opt = swirlcons_get_options(glob);
 
     /* Initialize virtual table for ForestClaw */
     fclaw2d_vtable_initialize();
     fclaw2d_diagnostics_vtable_initialize();
 
     /* Initialize virtual tables for solvers */
-    if (user_opt->claw_version == 4)
-    {
-        fc2d_clawpack46_vtable_initialize();
-    }
-    else if (user_opt->claw_version == 5)
-    {
-        fc2d_clawpack5_vtable_initialize();
-    }
+    fc2d_clawpack46_solver_initialize();
 
     swirlcons_link_solvers(glob);
 
@@ -254,7 +240,6 @@ main (int argc, char **argv)
     fclaw_options_t             *fclaw_opt;
     fclaw2d_clawpatch_options_t *clawpatch_opt;
     fc2d_clawpack46_options_t   *claw46_opt;
-    fc2d_clawpack5_options_t    *claw5_opt;
 
     fclaw2d_global_t            *glob;
     fclaw2d_domain_t            *domain;
@@ -269,7 +254,6 @@ main (int argc, char **argv)
[…]

### 2017-07-21 applications/clawpack/advection/2d/swirl_cons/swirlcons_user.cpp

@@ -29,12 +29,10 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 
 /* Two versions of Clawpack */
 #include <fc2d_clawpack46.h>
+#include <fc2d_clawpack46_options.h>
 #include <clawpack46_user_fort.h>  /* Headers for user defined fortran files */
 
-#include <fc2d_clawpack5.h>
-#include <clawpack5_user_fort.h>
-
-#include "../all/clawpack_user.h"
+#include "../all/advection_user_fort.h"
 
 void swirlcons_link_solvers(fclaw2d_global_t *glob)
 {

### 2017-07-21 src/patches/clawpatch/fclaw2d_clawpatch_output_netcdf.c
changes attributes to variables so that we can write them later as independent operators

@@ -33,21 +33,13 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 #include <fclaw2d_options.h>
 
 #include <netcdf.h>
+#include <netcdf_par.h>
 
 #define ERRCODE 2
 #define ERR(e) {printf("Error: %s\n", nc_strerror(e)); exit(ERRCODE);}
 
-typedef struct netcdf_data
-{
-    int ncid;
-    int step_inc;
-    int total_steps;  /* This level takes total_steps of step_inc each */
 
-    double initial_time;
-    double current_time;
-    double dt_step;
-} fclaw2d_level_data_t;
-
+#if 0
 void cb_clawpatch_output_netcdf_defpatch (fclaw2d_domain_t * domain,
                                           fclaw2d_patch_t * this_patch,
                                           int this_block_idx, int this_patch_idx,
@@ -137,6 +129,7 @@ void cb_clawpatch_output_netcdf_defpatch (fclaw2d_domain_t * domain,
     if ((retval = nc_enddef(subgroupid)))
         ERR(retval);
 }
+#endif
 
 void cb_clawpatch_output_netcdf_writeq (fclaw2d_domain_t * domain,
                                         fclaw2d_patch_t * this_patch,
@@ -147,14 +140,21 @@ void cb_clawpatch_output_netcdf_writeq (fclaw2d_domain_t * domain,
     fclaw2d_global_t *glob = (fclaw2d_global_t*) s->glob;
 
     int ncid = *((int *) s->user);
-    int groupid,qid;
+    int groupid,qid,patchlocationid,patchinfoid;
     int retval;
-    int patch_num, level;
+    int patch_num, level, meqn, mx, my, mbc;
+    double xlower, ylower, dx, dy;
+    double* q;
+    double patchlocation[4];
+    int patchinfo[2];
 
     /* Get info not readily available to user */
     fclaw2d_patch_get_info(glob->domain,this_patch,
                            this_block_idx,this_patch_idx,
                            &patch_num,&level);
+    
+    fclaw2d_clawpatch_grid_data(glob,this_patch,&mx,&my,&mbc,
+                                &xlower,&ylower,&dx,&dy);
     
     char patchname[10];
     sprintf(patchname, "patch%04d", patch_num);
[…]

### 2017-07-21 src/patches/clawpatch/fclaw2d_clawpatch_output_netcdf.c
add some comments

@@ -217,7 +217,7 @@ void fclaw2d_clawpatch_header_netcdf(fclaw2d_global_t* glob, int iframe, int nci
 }
 
 static
-void fclaw2d_clawpatch_defpatch_netcdf(fclaw2d_global_t* glob, int ipatch, int ncid)
+void fclaw2d_clawpatch_defgroup_netcdf(fclaw2d_global_t* glob, int ipatch, int ncid)
 {
     const fclaw2d_clawpatch_options_t *clawpatch_opt = fclaw2d_clawpatch_get_options(glob);
     
@@ -294,21 +294,24 @@ void fclaw2d_clawpatch_output_netcdf(fclaw2d_global_t* glob,int iframe)
     char filename[12];
     sprintf(filename,"claw%04d.nc",iframe);
 
+    /* serial */
     // if ((retval = nc_create(filename, NC_NETCDF4, &ncid)))
         // ERR(retval);
+
+    /* parallel */
     if ((retval = nc_create_par(filename, NC_NETCDF4|NC_MPIIO, glob->mpicomm, 
                                 MPI_INFO_NULL, &ncid)));
 
     fclaw2d_clawpatch_header_netcdf(glob, iframe, ncid);
     
+    /* Create groups for patches */
     ngrids = glob->domain->global_num_patches;
     for (int i = 0; i < ngrids; ++i)
     {
-        fclaw2d_clawpatch_defpatch_netcdf(glob,i,ncid);
+        fclaw2d_clawpatch_defgroup_netcdf(glob,i,ncid);
     }
 
-    /* Write out each patch to clawXXXX.nc */
-    // fclaw2d_global_iterate_patches (glob, cb_clawpatch_output_netcdf_defpatch, &ncid);
+    /* Write out patch data*/
     fclaw2d_global_iterate_patches (glob, cb_clawpatch_output_netcdf_writeq, &ncid);
     
     /* Close the file. */

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/1dadv/driver.f
(clawpack/advection/1d reference solution) Update 1d reference solution

@@ -12,7 +12,7 @@ c     # set parameters for maximum array sizes used in declarations
 c     # these must be increased for larger problems.
 c
 c
-      parameter (maxmx =   4096)
+      parameter (maxmx =     8192)
       parameter (mwork =   122864)
 
       parameter (mbc = 2)

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/1dadv/setaux.f

@@ -3,7 +3,7 @@
       implicit double precision (a-h,o-z)
       dimension aux(1-mbc:maxmx+mbc, maux)
 
-      double precision xe,xc, pi, dx1, dx2
+      double precision xe,xc, pi, dx1, dx2, a
       integer i, example
 
       common /compi/ pi
@@ -26,13 +26,21 @@ c             # Velocity should have a delta function at x=0.5
                   aux(i,2) = 0            
               endif
           elseif (example .eq. 2) then
-              aux(i,1) = cos(2*pi*xe) + 1.d0
+              aux(i,1) = cos(2*pi*xe)
               aux(i,2) = -2*pi*sin(2*pi*xc)
           elseif (example .eq. 3) then
               aux(i,1) = 0.1d0*sin(2*pi*xe)*sin(16*pi*xe)
               dx1 =  2*pi*cos(2*pi*xc)*sin(16*pi*xc)
               dx2 = 16*pi*sin(2*pi*xc)*cos(16*pi*xc)
               aux(i,2) = 0.1*(dx1 + dx2)
+          elseif (example .eq. 4) then
+              a = 0.01d0
+              aux(i,1) = tanh((xe-0.5d0)/a)
+              aux(i,2) = (1-tanh((xc-0.5d0)/a)**2)/a
+          elseif (example .eq. 5) then
+              a = 0.1d0
+              aux(i,1) = -tanh((xe-0.5d0)/a)
+              aux(i,2) = -(1-tanh((xc-0.5d0)/a)**2)/a
           endif
       enddo
 

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/notes/fwave.tex
(clawpack/advection/swirl(cons)) Updates to Riemann solvers"
(clawpack/advection/swirl (cons)) Updates to conservative updates

@@
 \begin{document}
-\paragraph{Scalar approach to conservation}
+\section{Introduction}
+\section{Numerical approaches}
+
+\subsection{Scalar approach to conservation}
 In \cite{le:2002}, LeVeque describes a solution to the conservation law
@@
 \end{align}
-where $u = u(x)$ is a spatially varying velocity field, and $q = q(x,t)$ is a conserved 
-tracer in  the flow field.  At a mesh cell interface, we have two velocities, 
-$u_r$ and $u_\ell$, which will not in general be equal.  As a result, satisfying a jump
-condition 
+where $u = u(x)$ is a spatially varying velocity field, and $q = q(x,t)$ is a conserved  tracer in  the flow field.  Using the analogy of boxes moving along a series of conveyor belts operating
+at speeds prescribed by $u(x)$, LeVeque describes a solution involving two waves.
+One moves at local speeds given by each conveyor belts, and stationary waves at the interfaces between conveyor belts.
+To see this, we let suppose we have two conveyor belts.
+The left moves at speed $u_\ell$ and the right moves at speed $u_r$.
+Each conveyor belt has a density of boxes, given by $q_\ell$ and $\q_r$.
+If we have $u_\ell > u_r > 0$, boxes will pile up on the right conveyor belt  as faster moving boxes from the left are transfered to the slower moving conveyor belt on the right.
+If $0 < u_\ell < u_r$, the density of boxes on the right conveyor belt will drop as faster moving boxes on the right move away from boxes arriving at a slower rate from the left.
+In either case, we expect a jump
+in box density both at the wave moving at speed $u_r$ along the right conveyor belt, and at the interface between the left and right conveyor belts.
+
+Mathematically, we can see that single wave is not enough by considering the Riemann problem at an interface between conveyor belts.   To solve the Riemann problem, we might try to connect left and 
+right states $q_\ell$ and $q_r$ by a single wave moving at some speed $s$ satisfying
 \begin{equation}
@@
 \end{equation}
-for some speed $s$ will not in general be possible.   Assuming that $u_r > 0$The correct solution actually has two waves, one moving at speed 0 and a second at speed $u_r$.  To determine the 
-intermediate state $q^*$ we impose flux continuity across the speed 0 wave.
+Since $u_r \ne u_\ell$ in general, a single wave that connects the left and 
+right states $q_\ell$ and $q_r$ does not exist.  Instead, we write the conservation 
+law as
+\begin{align*}
+q_t + u_r q_x & = 0, \qquad \mbox{in the left cell} \\
+q_t + u_\ell q_x & = 0, \qquad \mbox{in the right cell}
+\end{align*}
+and introduce two waves, one moving at speed zero, and the second moving at speed $u_r > 0$.  
+We seek an intermediate state $q^*$ that can be used to connect to left and right states.
+Across the wave moving at speed $u_r > 0$, we impose the condition
 \begin{equation}
+u_r q_r - u^* q^* = u_r(q_r - q^*)
 \end{equation}
+to get $u^* = u_r$.  Across the stationary wave, we impose a flux continuity condition
+\begin{equation}
+u^* q^* - u_\ell q_\ell = 0
+\end{equation}
+to get
+\begin{equation*}
+q^* = \frac{u_\ell q_\ell}{u_r}.
+\end{equation*}
+The waves moving at speeds 0 and speed $u_r > 0$ are given by
+\begin{align*}
+\mathcal{W}_1 & \equiv q^* - q_\ell = \frac{u_\ell q_\ell}{u_r} - q_\ell = 
+-\frac{u_r - u_\ell}{u_r}q_\ell\\
+\mathcal{W}_2 & \equiv q_r - q^* = q_r - \frac{u_\ell q_\ell}{u_r} = 
+\frac{1}{u_r}\left(u_r q_r - u_\ell q_\ell\right)
+\end{align*}
+The fluctuations are then defined in terms of these waves as
+\begin{align*}
+\mathcal{A}^+ \Delta Q & \equiv u_r \mathcal{W}_1 = u_r q_r - u_\ell q_\ell \\
+\mathcal{A}^- \Delta Q & \equiv 0
+\end{align*}
+The sum of the fluctuations satisfy the conservation requirement
+\begin{equation*}
+\mathcal{A}^+ \Delta Q + \mathcal{A}^- \Delta Q = f(q_r) - f(q_\ell)
+\end{equation*}
 
-\paragraph{2-wave decomposition approach.}
+\subsection{2-wave decomposition approach.}
 Consider the following conservation law for the conservative form of the scalar advection equation.
@@
 \begin{align}
-\mathcal W_1 & \equiv \alpha_1 \mathbf r^1 = - \frac{u_r - u_\ell}{\widehat{u}-a} 
+\mathcal W_1  \equiv \alpha_1 \mathbf r^1 & = - \frac{u_r - u_\ell}{\widehat{u}-a} 
 \left(\begin{array}{c} \widehat{q} \\ a - \widehat{u}\end{array}\right) 
@@
 1 \end{array}\right) \\
-\mathcal W_2 & \equiv \alpha_2 \mathbf r^2 = 
+\mathcal W_2  \equiv \alpha_2 \mathbf r^2 & = 
 \left(q_r - q_\ell + \frac{u_r - u_\ell}{\widehat{u}-a}\widehat{q} \right)
@@
 
-\paragraph{The F-wave approach.} To correct this problem, we use the f-wave approach and
+\subsection{The F-wave approach.} To correct this problem, we use the f-wave approach and
 split the jump in the flux directly, avoiding the need to create waves $\mathcal W_p$.  
@@
 
-\paragraph{numerical results.}  We test the two above approaches
+\section{Numerical results.}  We test the two above approaches
 

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/afterframe.m

@@ -26,35 +26,46 @@ elseif PlotType == 4
     dir = './1dadv/';
     dim = 1;
     [amrdata1d,t1d] = readamrdata(dim,Frame,dir);
-    [q1d,x1d] = plotframe1ez(amrdata1d,mq,'b.-');    
+    [q1d,x1d] = plotframe1ez(amrdata1d,mq,'b.-');  
+    
+    if (abs(t1d-t) > 1e-12) 
+        fprintf('t1d = %16.8e\n',t1d);
+        fprintf('t   = %16.8e\n',t);
+        error('Reference solution time incorrect');
+    end
+    
+    % Compute errors
+    pp = csape(x1d,q1d,'periodic');
+    qxc = ppval(pp,xcenter);
+    qerr = abs(qxc(:)-q(:,2));
+    qnorm(1) = sum(qerr)*dx;
+    qnorm(2) = sqrt(sum(qerr.^2*dx));
+    qnorm(3) = max(qerr);
+    fprintf('%5d %12.4e %12.4e %12.4e %12.4e %12.4e\n',mx,qnorm(1),...
+        qnorm(2),qnorm(3),qmin,qmax);
 
     % Plot zeros of velocity function
-    vel_case = 3;
+    vel_case = 5;
     yl = [-5,25];
     switch vel_case
         case 1
-            % Case 1 : u= cos(2*pi*x)
-            u = @(x) cos(2*pi*x);
-            plot([0.25,0.25],yl,'k--');
-            plot([0.75,0.75],yl,'k--');
+            u = @(x) (x < 0.5)*(-0.5) + (x >= 0.5)*(0.5);
         case 2
-            % Case 2 : u = cos(2*pi*x) + 0.5
-            u = @(x) cos(2*pi*x) + 0.5;
-            z = acos(-0.5)/(2*pi);
+            % Case 2 : u = cos(2*pi*x) + a
+            a = 0;
+            u = @(x) cos(2*pi*x) + a;
+            z = acos(-a)/(2*pi);
             plot([z,z],yl,'k--');
             plot(1-[z,z],yl,'k--');
         case 3
-            % Case 3 : u = cos(2*pi*x) + 1
-            u = @(x) cos(2*pi*x) + 1;
-            plot([0.5,0.5],yl,'k--');
-        case 4
-            % Case 4 : u = cos(2*pi*x) + 2
-            u = @(x) cos(2*pi*x) + 2;
-            % no zeros
-        case 5
[…]

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/plot_error.m

@@ -1,0 +1,223 @@
+function plot_error(ex,n)
+
+close all;
+
+if (nargin == 0)
+    ex = 1;
+    n = 1;
+end
+
+% Using wave-decomposition approach (no limiters)
+qerr = cell(5,3);
+
+% u(x) = cos(2*pi*x)
+qerr{1,1} = [32   1.5912e-01   4.1814e-01   1.4841e+00
+   64   9.9424e-02   3.7146e-01   1.7177e+00
+  128   4.6475e-02   2.2194e-01   1.4135e+00
+  256   1.2133e-02   6.9347e-02   6.5336e-01
+  512   3.5412e-03   1.9011e-02   2.0626e-01
+ 1024   1.1881e-03   5.3847e-03   5.4493e-02];
+
+qerr{1,2} = [32   4.6568e-01   1.4373e+00   6.9858e+00
+  64   2.8923e-01   1.2635e+00   7.2547e+00
+ 128   1.8978e-01   1.0750e+00   9.8899e+00
+ 256   1.0903e-01   8.4040e-01   1.1862e+01
+ 512   5.5548e-02   6.0228e-01   1.2282e+01
+1024   2.7773e-02   4.2619e-01   1.2328e+01]; 
+
+qerr{1,3} = [32   2.9178e-01   1.1711e+00   6.0072e+00
+   64   1.3997e-01   7.4514e-01   4.3342e+00
+  128   6.8514e-02   4.0987e-01   3.9477e+00
+  256   2.5097e-02   1.7597e-01   2.4684e+00
+  512   3.5412e-03   1.9011e-02   2.0626e-01
+ 1024   1.8254e-03   1.1264e-02   2.3070e-01];
+
+
+% u(x) = cos(2*pi*x) + 1
+
+qerr{2,1} =[32   1.8009e-01   4.7518e-01   2.0792e+00
+   64   6.3383e-02   2.4100e-01   1.4229e+00
+  128   1.9258e-02   7.8394e-02   4.7784e-01
+  256   5.1323e-03   2.1607e-02   1.5401e-01
+  512   1.5520e-03   6.0791e-03   4.0763e-02
+ 1024   6.8544e-04   2.3465e-03   1.5291e-02];
+
+qerr{2,2} = [32   1.8524e-01   4.7071e-01   1.7192e+00
+   64   9.2308e-02   3.0194e-01   1.4694e+00
[…]

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_new.f

@@ -20,7 +20,6 @@
       double precision ur,ul,qrr,qll, uhat,qhat, tau,uavg
       double precision a,b,c
 
-
       double precision dtcom, dxcom, dycom, tcom
       integer icom, jcom
 
@@ -50,13 +49,7 @@
           endif
 
           s(i,1) = uavg
-
-          if (ur .eq. ul) then
-              wave(i,1,1) = qrr - qll
-          else
-              wave(i,1,1) = (ur*qrr - ul*qll)/uhat   !! Same as above
-          endif
-
+          wave(i,1,1) = (ur*qrr - ul*qll)/uhat   !! Same as above
 
           if (uavg .eq. 0) then
 c             # This definition is needed to guarantee conservation, i.e. 

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_qstar.f

@@ -17,7 +17,7 @@
 
 
       integer i, iface
-      double precision qstar, qrr,qll,ur,ul, uhat
+      double precision qstar, qrr,qll,ur,ul, uhat, uavg
 
 
       double precision dtcom, dxcom, dycom, tcom
@@ -39,25 +39,21 @@ C     # ur > 0 : us*qs - ul*qll = 0 -->  us = ur; qs = ul*qll/ur;
           qrr = ql(i,1)
           qll = qr(i-1,1)
 
-          uhat = (ur+ul)/2.d0
-
 c         # ur and ul have the same sign
           if (ur .gt. 0 .and. ul .gt. 0) then
-              qstar = ul*qll/ur
-              wave(i,1,1) = qrr - qstar
+              wave(i,1,1) = (ur*qrr - ul*qll)/ur
               s(i,1) = ur
               amdq(i,1) = 0.d0
               apdq(i,1) = ur*qrr - ul*qll
           elseif (ur .lt. 0 .and. ul .lt. 0) then
-              qstar = ur*qrr/ul
-              wave(i,1,1) = (ur*qrr - ul*qll)/ur
-              s(i,1) = ur
+              wave(i,1,1) = (ur*qrr - ul*qll)/ul
+              s(i,1) = ul
               amdq(i,1) = ur*qrr - ul*qll
               apdq(i,1) = 0.d0
           else
 C             # ul and ur have different signs
-              qstar = 0
-              uhat = max(abs(uhat),1e-12)
+              uavg = (ur+ul)/2.d0
+              uhat = max(abs(uavg),1d-12)
               s(i,1) = uhat
               wave(i,1,1) = (ur*qrr - ul*qll)/uhat
               amdq(i,1) = -ul*qll

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/user_4.6/swirlcons_b4step2.f

@@ -41,7 +41,7 @@ c           # Cell-centered velocity
       double precision function ucc(xp,yp)
       implicit none
 
-      double precision xp,yp,pi, c,rp2,ul,ur,uavg
+      double precision xp,yp,pi, c,rp2,ul,ur,uavg, a
       integer example
       common /compi/ pi
       common /comex/ example
@@ -55,14 +55,20 @@ c         # 1d : Piecewise constant velocity (-0.5, 0.5)
           endif
       elseif (example .eq. 2) then
 c         # 1d : u = cos(2*pi*xp)       
-          ucc = cos(2*pi*xp) + 1.d0            
+          ucc = cos(2*pi*xp)
       elseif (example .eq. 3) then
 c         # 1d : More complicated 1d example
           ucc = 0.1*sin(2*pi*xp)*sin(16*pi*xp)
-      elseif ((example .eq. 4) .or. (example .eq. 5)) then
+      elseif (example .eq. 4) then 
+          a = 0.01d0                   
+          ucc = tanh((xp-0.5d0)/a)
+      elseif (example .eq. 5) then 
+          a = 0.1d0                   
+          ucc = -tanh((xp-0.5d0)/a)
+      elseif ((example .eq. 6) .or. (example .eq. 7)) then
 c         # 2d : Swirl example (computed from streamfunction)           
           ucc = 2*((sin(pi*xp))**2 * sin(pi*yp) * cos(pi*yp))
-          if (example .eq. 5) then
+          if (example .eq. 7) then
 c             # Add non-zero divergence            
               rp2 = (xp-0.5d0)**2 + (yp-0.5d0)**2
               rp2 = (xp-0.5d0)**2
@@ -87,13 +93,13 @@ c             # Add non-zero divergence
       common /comex/ example
 
 
-      if (example .le. 3) then
+      if (example .le. 4) then
 c         # 1d examples
           vcc = 0        
-      elseif ((example .eq. 4) .or. (example .eq. 5)) then
+      elseif ((example .eq. 5) .or. (example .eq. 6)) then
 c         # 2d examples
           vcc = -2*((sin(pi*yp))**2 * sin(pi*xp) * cos(pi*xp))
-          if (example .eq. 5) then
+          if (example .eq. 6) then
 c             # Add non-zero divergence              
               rp2 = (xp-0.5d0)**2 + (yp-0.5d0)**2
               rp2 = (xp-0.5d0)**2

### 2017-07-24 applications/clawpack/advection/2d/swirl_cons/plot_error.m
(clawpack/advection/swirl (cons)) New results

@@ -8,7 +8,7 @@ if (nargin == 0)
 end
 
 % Using wave-decomposition approach (no limiters)
-qerr = cell(5,3);
+qerr = cell(7,5);
 
 % u(x) = cos(2*pi*x)
 qerr{1,1} = [32   1.5912e-01   4.1814e-01   1.4841e+00
@@ -152,6 +152,44 @@ qerr{6,4} = [   32   6.8161e-02   1.5296e-01   6.0469e-01  -2.7187e-01   1.0000e
   512   2.9574e-03   7.1511e-03   2.7692e-02  -1.3439e-02   1.0000e+00
  1024   1.2571e-03   3.1260e-03   1.2836e-02  -6.1793e-03   1.0000e+00];
 
+% u(x) = -tanh((x-0.5)/0.1)
+qerr{7,1} = [   32   6.8220e-02   2.0568e-01   7.8131e-01   1.0005e+00   6.4950e+00
+   64   2.6516e-02   1.0842e-01   5.7044e-01   1.0005e+00   8.6233e+00
+  128   1.9785e-02   1.0044e-01   6.1693e-01   1.0005e+00   1.0389e+01
+  256   6.1397e-03   3.8040e-02   3.7708e-01   1.0005e+00   1.1476e+01
+  512   1.8022e-03   1.0875e-02   1.3820e-01   1.0005e+00   1.1959e+01
+ 1024   6.3045e-04   3.0375e-03   3.9603e-02   1.0005e+00   1.2121e+01];
+
+qerr{7,2} = [   32   6.7161e-02   2.5823e-01   1.0322e+00   1.0005e+00   6.7459e+00
+   64   2.4742e-02   1.0449e-01   5.5100e-01   1.0005e+00   9.3944e+00
+  128   2.5639e-02   1.3199e-01   8.3336e-01   1.0005e+00   1.1840e+01
+  256   2.3454e-02   1.5924e-01   1.5093e+00   1.0005e+00   1.3362e+01
+  512   1.4004e-02   1.2688e-01   1.8041e+00   1.0005e+00   1.3901e+01
+ 1024   7.1874e-03   8.9704e-02   1.8482e+00   1.0005e+00   1.4009e+01];
+
+qerr{7,3} = [  32   9.9205e-02   4.6539e-01   2.5771e+00   1.0005e+00   8.2908e+00
+   64   9.8526e-02   5.3836e-01   3.4631e+00   1.0005e+00   1.2307e+01
[…]

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/afterframe.m
(clawpack/1d conservative example) Add files

@@ -1,0 +1,65 @@
+hold on;
+dir = './1dadv/';
+dim = 1;
+[amrdata1d,t1d] = readamrdata(dim,Frame,dir);
+[q1d,x1d] = plotframe1ez(amrdata1d,mq,'b.-');
+
+if (abs(t1d-t) > 1e-12)
+    fprintf('t1d = %16.8e\n',t1d);
+    fprintf('t   = %16.8e\n',t);
+    error('Reference solution time incorrect');
+end
+
+% Compute errors
+pp = csape(x1d,q1d,'periodic');
+qxc = ppval(pp,xcenter);
+qerr = abs(qxc(:)-q(:));
+qnorm(1) = sum(qerr)*dx;
+qnorm(2) = sqrt(sum(qerr.^2*dx));
+qnorm(3) = max(qerr);
+fprintf('%5d %12.4e %12.4e %12.4e %12.4e %12.4e\n',mx,qnorm(1),...
+    qnorm(2),qnorm(3),qmin,qmax);
+
+% Plot zeros of velocity function
+vel_case = 3;
+yl = [-5,25];
+switch vel_case
+    case 1
+        % Case 1 : u = cos(2*pi*x)  + 2
+        u = @(x) cos(2*pi*x) + a;
+        z = acos(-a)/(2*pi);
+        plot([z,z],yl,'k--');
+        plot(1-[z,z],yl,'k--');
+
+    case 2
+        % Case 2 : u = cos(2*pi*x) 
+        u = @(x) cos(2*pi*x) + a;
+        z = acos(-a)/(2*pi);
+        plot([z,z],yl,'k--');
+        plot(1-[z,z],yl,'k--');
+        
+    case 3
+        y1 = @(x) 0.5*tanh((x-0.65)/0.025) + 1;
+        y2 = @(x) 0.5*tanh((0.35-x)/0.04) + 1;
+        u = @(x) y1(x) + y2(x) - 1.5;
+    otherwise
+        %u = @(x) x*nan;
+end
+
+xc = linspace(0,1,200);
+plot(xc,u(xc),'r--');   % Velocity
+plot(xcenter,0*xcenter,'k');
+
+fprintf('qmin = %12.4e\n',qmin);
+
+qm = min([max([q1d(:); qmax]),15]);
+ylim([-2 max([2.1,1.1*qm]);])
+
+set(gca,'fontsize',16);
+set(gca,'box','on');
+
+shg
+
+hold off;
+
+clear afterframe

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/driver.f

@@ -1,0 +1,32 @@
+      program driver
+c
+c  Generic driver routine for claw1ez
+c
+c  Author: Randall J. LeVeque
+c  Version of March, 1999 --  CLAWPACK Version 4.0
+c
+c
+      implicit double precision (a-h,o-z)
+
+c     # set parameters for maximum array sizes used in declarations
+c     # these must be increased for larger problems.
+c
+c
+      parameter (maxmx =     8192)
+      parameter (mwork =   122864)
+
+      parameter (mbc = 2)
+      parameter (meqn = 1)
+      parameter (mwaves = 1)
+      parameter (maux = 2)
+c
+      dimension q(1-mbc:maxmx+mbc, meqn)
+      dimension  aux(1-mbc:maxmx+mbc,maux)   !# dummy variable since no aux arrays used
+      dimension work(mwork)
+      dimension mthlim(mwaves)
+
+      call claw1ez(maxmx,meqn,mwaves,mbc,maux,mwork,mthlim,
+     &             q,work,aux)
+
+      stop 
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/rp1cons_ec.f

@@ -1,0 +1,46 @@
+      subroutine rp1(maxmx,meqn,mwaves,mbc,mx,ql,qr,auxl,auxr,
+     &		       wave,s,amdq,apdq)
+
+      implicit none
+
+      integer maxmx, meqn, mwaves, mbc,mx
+
+      double precision   ql(1-mbc:maxmx+mbc, meqn)
+      double precision   qr(1-mbc:maxmx+mbc, meqn)
+      double precision    s(1-mbc:maxmx+mbc, mwaves)
+      double precision wave(1-mbc:maxmx+mbc, meqn, mwaves)
+      double precision amdq(1-mbc:maxmx+mbc, meqn)
+      double precision apdq(1-mbc:maxmx+mbc, meqn)
+      double precision auxl(1-mbc:maxmx+mbc, *)
+      double precision auxr(1-mbc:maxmx+mbc, *)
+
+      integer i
+      double precision qrr,qll,ur,ul,uavg,uhat,qhat,fstar
+
+
+      do i = 2-mbc,mx+mbc
+          ur = auxl(i,1)    !! cell-centered.
+          ul = auxr(i-1,1)
+          uhat = (ur + ul)/2.d0
+
+          qrr = ql(i,1)
+          qll = qr(i-1,1)
+
+          wave(i,1,1) = qrr - qll
+          s(i,1) = uhat
+
+          if (uhat .ge. 0.d0) then
+              fstar = uhat*qll
+          else
+              fstar = uhat*qrr
+          endif
+
+          amdq(i,1) = fstar - ul*qll              !flux
+          apdq(i,1) = ur*qrr - fstar            !-flux
+      enddo
+c
+      return
+      end
+
+
+

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/rp1cons_qstar.f

@@ -1,0 +1,57 @@
+      subroutine rp1(maxmx,meqn,mwaves,mbc,mx,ql,qr,auxl,auxr,
+     &		       wave,s,amdq,apdq)
+
+      implicit none
+
+      integer maxmx, meqn, mwaves, mbc,mx
+
+      double precision   ql(1-mbc:maxmx+mbc, meqn)
+      double precision   qr(1-mbc:maxmx+mbc, meqn)
+      double precision    s(1-mbc:maxmx+mbc, mwaves)
+      double precision wave(1-mbc:maxmx+mbc, meqn, mwaves)
+      double precision amdq(1-mbc:maxmx+mbc, meqn)
+      double precision apdq(1-mbc:maxmx+mbc, meqn)
+      double precision auxl(1-mbc:maxmx+mbc, *)
+      double precision auxr(1-mbc:maxmx+mbc, *)
+
+      integer i
+      double precision qrr,qll,ur,ul,uavg,uhat
+
+
+      do i = 2-mbc,mx+mbc
+         ur = auxl(i,1)
+         ul = auxr(i-1,1)
+
+         qrr = ql(i,1)
+         qll = qr(i-1,1)
+
+c         # ur and ul have the same sign
+          if (ur .gt. 0 .and. ul .gt. 0) then
+              wave(i,1,1) = (ur*qrr - ul*qll)/ur
+              s(i,1) = ur
+              amdq(i,1) = 0.d0
+              apdq(i,1) = ur*qrr - ul*qll
+          elseif (ur .lt. 0 .and. ul .lt. 0) then
+              wave(i,1,1) = (ur*qrr - ul*qll)/ul
+              s(i,1) = ul
+              amdq(i,1) = ur*qrr - ul*qll
+              apdq(i,1) = 0.d0
+          else
+C             # ul and ur have different signs
+              uavg = (ur+ul)/2.d0
+              uhat = max(abs(uavg),1d-12)
+              s(i,1) = uhat
+              wave(i,1,1) = (ur*qrr - ul*qll)/uhat
+              amdq(i,1) = -ul*qll
+              apdq(i,1) = ur*qrr
+          endif
+
+      enddo
+c
+c         
+c
+      return
+      end
+
+
+

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/rp1cons_wd.f

@@ -1,0 +1,58 @@
+      subroutine rp1(maxmx,meqn,mwaves,mbc,mx,ql,qr,auxl,auxr,
+     &		       wave,s,amdq,apdq)
+
+      implicit none
+
+      integer maxmx, meqn, mwaves, mbc,mx
+
+      double precision   ql(1-mbc:maxmx+mbc, meqn)
+      double precision   qr(1-mbc:maxmx+mbc, meqn)
+      double precision    s(1-mbc:maxmx+mbc, mwaves)
+      double precision wave(1-mbc:maxmx+mbc, meqn, mwaves)
+      double precision amdq(1-mbc:maxmx+mbc, meqn)
+      double precision apdq(1-mbc:maxmx+mbc, meqn)
+      double precision auxl(1-mbc:maxmx+mbc, *)
+      double precision auxr(1-mbc:maxmx+mbc, *)
+
+      integer i
+      double precision qrr,qll,ur,ul,uavg,uhat,qhat
+
+
+      do i = 2-mbc,mx+mbc
+          ur = auxl(i,1)
+          ul = auxr(i-1,1)
+
+          qrr = ql(i,1)
+          qll = qr(i-1,1)
+
+          qhat = (qrr + qll)/2.d0
+
+          uavg = (ur+ul)/2.d0
+          if (uavg .eq. 0) then
+              uhat = 1.d-12
+          else
+              uhat = uavg
+          endif
+
+          s(i,1) = uavg
+          wave(i,1,1) = (ur*qrr - ul*qll)/uhat   !! Same as above
+
+          if (uavg .eq. 0) then
+c             # This definition is needed to guarantee conservation, i.e. 
+c             #       f(qr) - f(ql) = apdq + amdq                      
+c             # and a consistent definition of f(qr), f(ql).  
+c             # If we base apdq,amdq on uhat, we will have inconsistent
+c             # definitions of f(q). 
+              amdq(i,1) = -ul*qhat
+              apdq(i,1) = ur*qhat
+          else
+              amdq(i,1) = min(uhat,0.d0)*wave(i,1,1)
+              apdq(i,1) = max(uhat,0.d0)*wave(i,1,1)
+          endif
+      enddo
+c
+      return
+      end
+
+
+

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/rp1fwave.f

@@ -1,0 +1,49 @@
+      subroutine rp1(maxmx,meqn,mwaves,mbc,mx,ql,qr,auxl,auxr,
+     &		       wave,s,amdq,apdq)
+
+      implicit none
+
+      integer maxmx, meqn, mwaves, mbc,mx
+
+      double precision   ql(1-mbc:maxmx+mbc, meqn)
+      double precision   qr(1-mbc:maxmx+mbc, meqn)
+      double precision    s(1-mbc:maxmx+mbc, mwaves)
+      double precision wave(1-mbc:maxmx+mbc, meqn, mwaves)
+      double precision amdq(1-mbc:maxmx+mbc, meqn)
+      double precision apdq(1-mbc:maxmx+mbc, meqn)
+      double precision auxl(1-mbc:maxmx+mbc, *)
+      double precision auxr(1-mbc:maxmx+mbc, *)
+
+      integer i
+      double precision qrr,qll,ur,ul,uavg,uhat,qhat
+
+
+      do i = 2-mbc,mx+mbc
+          ur = auxl(i,1)
+          ul = auxr(i-1,1)
+
+          qrr = ql(i,1)
+          qll = qr(i-1,1)
+
+          uhat = (ul + ur)/2.d0
+
+          wave(i,1,1) = ur*qrr - ul*qll
+          s(i,1) = uhat
+          if (uhat .eq. 0) then
+              amdq(i,1) = -ul*qll
+              apdq(i,1) = ur*qrr            
+          elseif (uhat .lt. 0) then
+              amdq(i,1) = wave(i,1,1)
+              apdq(i,1) = 0.d0
+          else
+              amdq(i,1) = 0.d0
+              apdq(i,1) = wave(i,1,1)
+          endif
+
+      enddo
+c
+      return
+      end
+
+
+

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/setaux.f

@@ -1,0 +1,50 @@
+      subroutine setaux(maxmx,mbc,mx,xlower,dx,maux,aux)
+
+      implicit none
+      integer maxmx, mbc,mx,maux
+      double precision dx, xlower
+      double precision aux(1-mbc:maxmx+mbc, maux)
+
+      double precision xc, pi, a, y1, y2, a1, a2
+      integer i, example
+
+      common /compi/ pi
+      common /comex/ example
+
+c     # aux(i,1) : edge velocity (u_{i-1/2})
+c     # aux(i,2) : cell-centered derivative of velocity (for src2)
+
+      do i = 1-mbc,mx+mbc
+          xc = xlower + (i-0.5)*dx
+
+          if (example .eq. 1) then
+c             # Velocity should have a delta function at x=0.5
+              if (xc .le. 0.5d0) then
+                  aux(i,1) = -0.5d0
+                  aux(i,2) = 0
+              else
+                  aux(i,1) = 0.5d0
+                  aux(i,2) = 0            
+              endif
+          elseif (example .eq. 2) then
+              aux(i,1) = cos(2*pi*xc) + 2.d0
+          elseif (example .eq. 3) then
+              a = 0.1d0
+              aux(i,1) = a*sin(2*pi*xc)*sin(16*pi*xc)
+          elseif (example .eq. 4) then
+              a1 = 0.025d0
+              a2 = 0.04d0
+              y1 = 0.5*tanh((xc-0.65d0)/a1) + 1
+              y2 = 0.5*tanh((0.35d0-xc)/a2) + 1
+              aux(i,1) = y1 + y2 - 1.5
+          elseif (example .eq. 5) then
+              a = 0.01d0
+              aux(i,1) = tanh((xc-0.5d0)/a)
+          elseif (example .eq. 6) then
+              a = 0.1d0
+              aux(i,1) = -tanh((xc-0.5d0)/a)
+          endif
+      enddo
+
+      return
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/setplot1.m

@@ -1,0 +1,12 @@
+OutputFlag = 'ascii';         % default value.
+OutputDir = './';            % Default (reassign them here anyway)
+
+ForestClaw = 0;     % Plot using ForestClaw preferences.
+
+
+mq = 1;                       % which component(s) of q to plot
+UserVariable = 0;             % set to 1 to specify a user-defined variable
+UserVariableFile = ' ';       % name of m-file mapping data to q
+MappedGrid = 0;               % set to 1 if mapc2p.m exists for nonuniform grid
+PlotStyle = setplotstyle('ko');  % used in plot command for line color and type
+MaxFrames = 1000;                % max number of frames to loop over

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/driver.f
(clawpack/1d conservative example) Add more files

@@ -1,0 +1,32 @@
+      program driver
+c
+c  Generic driver routine for claw1ez
+c
+c  Author: Randall J. LeVeque
+c  Version of March, 1999 --  CLAWPACK Version 4.0
+c
+c
+      implicit double precision (a-h,o-z)
+
+c     # set parameters for maximum array sizes used in declarations
+c     # these must be increased for larger problems.
+c
+c
+      parameter (maxmx =     16384)
+      parameter (mwork =   131104)
+
+      parameter (mbc = 2)
+      parameter (meqn = 1)
+      parameter (mwaves = 1)
+      parameter (maux = 2)
+c
+      dimension q(1-mbc:maxmx+mbc, meqn)
+      dimension  aux(1-mbc:maxmx+mbc,maux)   !# dummy variable since no aux arrays used
+      dimension work(mwork)
+      dimension mthlim(mwaves)
+
+      call claw1ez(maxmx,meqn,mwaves,mbc,maux,mwork,mthlim,
+     &             q,work,aux)
+
+      stop 
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/lib/limiter.f

@@ -1,0 +1,60 @@
+c
+c
+c     =====================================================
+      subroutine limiter(maxm,meqn,mwaves,mbc,mx,wave,s,mthlim)
+c     =====================================================
+c
+c     # Apply a limiter to the waves.
+c     # The limiter is computed by comparing the 2-norm of each wave with
+c     # the projection of the wave from the interface to the left or
+c     # right onto the current wave.  For a linear system this would
+c     # correspond to comparing the norms of the two waves.  For a 
+c     # nonlinear problem the eigenvectors are not colinear and so the 
+c     # projection is needed to provide more limiting in the case where the
+c     # neighboring wave has large norm but points in a different direction
+c     # in phase space.
+c
+c     # The specific limiter used in each family is determined by the
+c     # value of the corresponding element of the array mthlim, as used in
+c     # the function philim.
+c     # Note that a different limiter may be used in each wave family.
+c
+c     # dotl and dotr denote the inner product of wave with the wave to
+c     # the left or right.  The norm of the projections onto the wave are then
+c     # given by dotl/wnorm2 and dotr/wnorm2, where wnorm2 is the 2-norm
+c     # of wave.
+c
+      implicit real*8(a-h,o-z)
+      dimension mthlim(mwaves)
+      dimension wave(1-mbc:maxm+mbc, meqn, mwaves)
+      dimension    s(1-mbc:maxm+mbc, mwaves)
+c
+c
+      do 50 mw=1,mwaves
+         if (mthlim(mw) .eq. 0) go to 50
+         dotr = 0.d0
+         do 40 i = 0, mx+1
+            wnorm2 = 0.d0
+            dotl = dotr
+            dotr = 0.d0
+            do 20 m=1,meqn
+               wnorm2 = wnorm2 + wave(i,m,mw)**2
+               dotr = dotr + wave(i,m,mw)*wave(i+1,m,mw)
+   20          continue
+            if (i.eq.0) go to 40
+            if (wnorm2.eq.0.d0) go to 40
+c
+            if (s(i,mw) .gt. 0.d0) then
+                wlimitr = philim(wnorm2, dotl, mthlim(mw))
+              else
+                wlimitr = philim(wnorm2, dotr, mthlim(mw))
+              endif
+c
+            do 30 m=1,meqn
+               wave(i,m,mw) = wlimitr * wave(i,m,mw)
+   30          continue
+   40       continue
+   50    continue
+c
+      return
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/qinit.f

@@ -1,0 +1,20 @@
+c
+c
+c =========================================================
+       subroutine qinit(maxmx,meqn,mbc,mx,xlower,dx,q,maux,aux)
+c =========================================================
+c
+c     # Set initial conditions for q.
+c
+      implicit double precision (a-h,o-z)
+      dimension q(1-mbc:maxmx+mbc, meqn)
+      dimension aux(1-mbc:maxmx+mbc, maux)
+
+      pi = 4.d0*datan(1.d0)
+c
+      do i=1-mbc,mx+mbc
+          q(i,1) = 1.0
+      enddo
+c
+      return
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/rp1adv.f

@@ -1,0 +1,35 @@
+      subroutine rp1(maxmx,meqn,mwaves,mbc,mx,ql,qr,auxl,auxr,
+     &		       wave,s,amdq,apdq)
+
+      implicit double precision (a-h,o-z)
+      dimension   ql(1-mbc:maxmx+mbc, meqn)
+      dimension   qr(1-mbc:maxmx+mbc, meqn)
+      dimension    s(1-mbc:maxmx+mbc, mwaves)
+      dimension wave(1-mbc:maxmx+mbc, meqn, mwaves)
+      dimension amdq(1-mbc:maxmx+mbc, meqn)
+      dimension apdq(1-mbc:maxmx+mbc, meqn)
+      dimension auxl(1-mbc:maxmx+mbc, *)
+      dimension auxr(1-mbc:maxmx+mbc, *)
+
+      do i = 2-mbc,mx+mbc
+         ur = auxl(i,1)
+         ul = auxr(i-1,1)
+
+         qrr = ql(i,1)
+         qll = qr(i-1,1)
+
+         uhat = ur    !! left edge value
+
+         wave(i,1,1) = qrr-qll
+         amdq(i,1) = min(uhat,0.d0)*wave(i,1,1)
+         apdq(i,1) = max(uhat,0.d0)*wave(i,1,1)
+         s(i,1) = uhat
+      enddo
+c
+c         
+c
+      return
+      end
+
+
+

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/setaux.f

@@ -1,0 +1,60 @@
+      subroutine setaux(maxmx,mbc,mx,xlower,dx,maux,aux)
+
+      implicit double precision (a-h,o-z)
+      dimension aux(1-mbc:maxmx+mbc, maux)
+
+      double precision xe,xc, pi, dx1, dx2, a, dy1, dy2, y1, y2
+      double precision a1, a2
+      integer i, example
+
+      common /compi/ pi
+      common /comex/ example
+
+c     # aux(i,1) : edge velocity (u_{i-1/2})
+c     # aux(i,2) : cell-centered derivative of velocity (for src2)
+
+      do i = 1-mbc,mx+mbc
+          xe = xlower + (i-1)*dx
+          xc = xlower + (i-0.5)*dx
+
+          if (example .eq. 1) then
+c             # Velocity should have a delta function at x=0.5
+              if (x .le. 0.5d0) then
+                  aux(i,1) = -0.5d0
+                  aux(i,2) = 0
+              else
+                  aux(i,1) = 0.5d0
+                  aux(i,2) = 0            
+              endif
+          elseif (example .eq. 2) then
+              aux(i,1) = cos(2*pi*xe) + 2.d0
+              aux(i,2) = -2*pi*sin(2*pi*xc)
+          elseif (example .eq. 3) then
+              a = 0.1d0
+              aux(i,1) = a*sin(2*pi*xe)*sin(16*pi*xe)
+              dx1 =  2*pi*cos(2*pi*xc)*sin(16*pi*xc)
+              dx2 = 16*pi*sin(2*pi*xc)*cos(16*pi*xc)
+              aux(i,2) = a*(dx1 + dx2)
+          elseif (example .eq. 4) then
+              a1 = 0.025d0
+              a2 = 0.04d0
+              y1 = 0.5*tanh((xc-0.65d0)/a1) + 1
+              y2 = 0.5*tanh((0.35d0-xc)/a2) + 1
+              aux(i,1) = y1 + y2 - 1.5d0
+              dy1 = 0.5d0*(1-tanh((xc-0.65d0)/a1)**2)/a1
+              dy2 = -0.5d0*(1-tanh((0.35d0-xc)/a2)**2)/a2
+              aux(i,2) = dy1 + dy2
+              aux(i,2) = dy1 + dy2
+          elseif (example .eq. 5) then
+              a = 0.01d0
+              aux(i,1) = tanh((xe-0.5d0)/a)
[…]

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/setprob.f

@@ -1,0 +1,17 @@
+      subroutine setprob
+      implicit none
+
+      double precision pi
+      integer example
+
+      common /compi/ pi
+      common /comex/ example
+
+      open(10,file='setprob.data')
+      read(10,*) example
+      close(10)
+
+      pi = 4.d0*atan(1.d0)
+
+      return
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dcons/1dadv/src1.f

@@ -1,0 +1,31 @@
+
+
+c
+c
+c =========================================================
+      subroutine src1(maxmx,meqn,mbc,mx,xlower,dx,q,maux,aux,t,dt)
+c =========================================================
+      implicit none
+
+      integer maxmx,meqn,mbc,mx,maux
+      double precision t,dt,xlower,dx
+      double precision q(1-mbc:maxmx+mbc, meqn)
+      double precision aux(1-mbc:maxmx+mbc, maux)
+
+      double precision xcell,ux,F1,F2,F3,F4
+      integer i
+c
+c     # 4th-order Runge-Kutta method to integrate q_t = -u_x*q
+c
+      do i = 1,mx
+         xcell = xlower + (i-0.5d0)*dx
+         ux = aux(i,2)
+         F1 = -ux*q(i,1)
+         F2 = -ux*(q(i,1) + 0.5d0*dt*F1)
+         F3 = -ux*(q(i,1) + 0.5d0*dt*F2)
+         F4 = -ux*(q(i,1) + dt*F3)
+         q(i,1) = q(i,1) + dt*(F1 + 4*(F2+F3)/2.d0 + F4)/6.d0
+      enddo
+c
+      return
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/afterframe.m
--amend

@@ -23,7 +23,7 @@ if PlotType == 1
     view(2);
 elseif PlotType == 4
     hold on;
-    dir = './1dadv/';
+    dir = './1dcons/1dadv/';
     dim = 1;
     [amrdata1d,t1d] = readamrdata(dim,Frame,dir);
     [q1d,x1d] = plotframe1ez(amrdata1d,mq,'b.-');  
@@ -45,7 +45,7 @@ elseif PlotType == 4
         qnorm(2),qnorm(3),qmin,qmax);
 
     % Plot zeros of velocity function
-    vel_case = 5;
+    vel_case = 4;
     yl = [-5,25];
     switch vel_case
         case 1
@@ -61,8 +61,11 @@ elseif PlotType == 4
             u = @(x) sin(2*pi*x).*sin(16*pi*x);
         case 4
             a = 0.01;
-            u = @(x) tanh((x-0.5)/a);
+            u = @(x) 1 + 0.5*tanh((x-0.5)/a);
         case 5
+            a = 0.01;
+            u = @(x) tanh((x-0.5)/a);
+        case 6
             a = 0.1;
             u = @(x) -tanh((x-0.5)/a);
         otherwise
@@ -77,8 +80,7 @@ elseif PlotType == 4
     
     qm = min([max([q1d(:); qmax]),15]);
     ylim([-2 max([2.1,1.1*qm]);])
-    
-    
+       
     hold off;
      
 end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/plot_error.m

@@ -1,244 +1,305 @@
-function plot_error(ex,n)
+function plot_error(ex,n,add_limiter)
 
 close all;
 
 if (nargin == 0)
     ex = 1;
     n = 1;
+    add_limiter = 0;
 end
 
 % Using wave-decomposition approach (no limiters)
 qerr = cell(7,5);
 
-% u(x) = cos(2*pi*x)
-qerr{1,1} = [32   1.5912e-01   4.1814e-01   1.4841e+00
-   64   9.9424e-02   3.7146e-01   1.7177e+00
-  128   4.6475e-02   2.2194e-01   1.4135e+00
-  256   1.2133e-02   6.9347e-02   6.5336e-01
-  512   3.5412e-03   1.9011e-02   2.0626e-01
- 1024   1.1881e-03   5.3847e-03   5.4493e-02];
-
-qerr{1,2} = [32   4.6568e-01   1.4373e+00   6.9858e+00
-  64   2.8923e-01   1.2635e+00   7.2547e+00
- 128   1.8978e-01   1.0750e+00   9.8899e+00
- 256   1.0903e-01   8.4040e-01   1.1862e+01
- 512   5.5548e-02   6.0228e-01   1.2282e+01
-1024   2.7773e-02   4.2619e-01   1.2328e+01]; 
-
-qerr{1,3} = [32   2.9178e-01   1.1711e+00   6.0072e+00
-   64   1.3997e-01   7.4514e-01   4.3342e+00
-  128   6.8514e-02   4.0987e-01   3.9477e+00
-  256   2.5097e-02   1.7597e-01   2.4684e+00
-  512   3.5412e-03   1.9011e-02   2.0626e-01
- 1024   1.8254e-03   1.1264e-02   2.3070e-01];
+% 1-2 : ustar (no lim, with lim)
+% 3-4 : wave-decomp (no-lim, with-lim)
+% 5-6 : edge (no-lim, with-lim)
+% 7-8 : fwave (no-lim, with-lim)
 
 
-% u(x) = cos(2*pi*x) + 1
+% -------------------------------------------
+% Ex. 1 : u(x) = cos(2*pi*x) + 2
+% -------------------------------------------
+qerr{1,1} = [  32   4.0960e-02   6.0462e-02   1.7538e-01   6.3310e-01   1.7444e+00
[…]

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/swirlcons.cpp

@@ -42,14 +42,11 @@ swirlcons_register (user_options_t *user, sc_options_t * opt)
     sc_options_add_double (opt, 0, "period", &user->period, 4,
                            "Period of the flow field [4]");
 
-    sc_options_add_int (opt, 0, "claw-version", &user->claw_version, 4,
-                           "Clawpack_version (4 or 5) [5]");
-
     sc_options_add_int (opt, 0, "example", &user->example, 3,
                            "Velocity field (1d, 2d, swirl, ...) (1-5) [3]");
 
-    sc_options_add_bool (opt, 0, "cons-rp", &user->cons_rp, 1,
-                           "Conservative riemann solver (T or F) [T]");
+    sc_options_add_int (opt, 0, "rp-solver", &user->rp_solver, 1,
+                           "Conservative riemann solver (1-4) [T]");
     user->is_registered = 1;
 
     return NULL;

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/swirlcons_user.cpp

@@ -42,33 +42,34 @@ void swirlcons_link_solvers(fclaw2d_global_t *glob)
     vt->problem_setup = &swirlcons_problem_setup;  /* Version-independent */
 
     const user_options_t* user = swirlcons_get_options(glob);
-    if (user->claw_version == 4)
+    fc2d_clawpack46_vtable_t *clawpack46_vt = fc2d_clawpack46_vt();
+    clawpack46_vt->qinit     = &CLAWPACK46_QINIT;
+    clawpack46_vt->setaux    = &CLAWPACK46_SETAUX;
+
+    switch(user->rp_solver)
     {
-        fc2d_clawpack46_vtable_t *clawpack46_vt = fc2d_clawpack46_vt();
-        clawpack46_vt->qinit     = &CLAWPACK46_QINIT;
-        clawpack46_vt->setaux    = &CLAWPACK46_SETAUX;
-        if (user->cons_rp)
-        {
+        case 1:
             clawopt->use_fwaves = 0;
-            clawpack46_vt->rpn2      = &RPN2CONS_CC;                                
-        }
-        else if (clawopt->use_fwaves)
-        {
-            clawpack46_vt->rpn2      = &RPN2FWAVE;                
-        }
-        else
-        {
+            clawpack46_vt->rpn2      = &RPN2CONS_QSTAR; 
+            break; 
+
+        case 2:
             clawopt->use_fwaves = 0;
-            clawpack46_vt->rpn2      = &RPN2CONS;                                
-        }
-        clawpack46_vt->rpt2      = &RPT2CONS_CC;
-        clawpack46_vt->b4step2   = &SWIRLCONS_B4STEP2;
+            clawpack46_vt->rpn2      = &RPN2CONS_WD; 
+            break; 
+
+        case 3:
+            clawopt->use_fwaves = 0;
+            clawpack46_vt->rpn2      = &RPN2CONS_EC; 
+            break;
+
+        case 4:
+            clawopt->use_fwaves = 1;
+            clawpack46_vt->rpn2      = &RPN2FWAVE; 
+            break;
     }
-    else if (user->claw_version == 5)
-    {
-        fclaw_global_essentialf("swirl (cons) : Clawpack 5.x implementation not implemented\n");
-        exit(0);
-    }
+    clawpack46_vt->rpt2      = &RPT2CONS_CC;
+    clawpack46_vt->b4step2   = &SWIRLCONS_B4STEP2;
 }
 
 void swirlcons_problem_setup(fclaw2d_global_t* glob)

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_new.f

@@ -1,5 +1,5 @@
-      subroutine rpn2cons(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
-     &                       auxl,auxr,wave,s,amdq,apdq)
+      subroutine rpn2cons_wd(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
+     &                      auxl,auxr,wave,s,amdq,apdq)
 
       implicit none
 

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_qstar.f

@@ -1,4 +1,4 @@
-      subroutine rpn2cons_cc(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
+      subroutine rpn2cons_qstar(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
      &                       auxl,auxr,wave,s,amdq,apdq)
 
       implicit none

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/user_4.6/swirlcons_b4step2.f

@@ -8,7 +8,7 @@
       double precision aux(1-mbc:maxmx+mbc,1-mbc:maxmy+mbc, maux)
 
       integer i, j, example
-      double precision tperiod, pi2, vt, xc,yc, ucc, vcc
+      double precision tperiod, pi2, vt, xc,yc, ucc, vcc, xe,ye
 
       common /comvt/ tperiod,pi2
       common /comex/ example
@@ -29,9 +29,14 @@ c           # coordinates of lower left corner of grid cell:
             xc = xlower + (i-0.5)*dx
             yc = ylower + (j-0.5)*dy
 
+            xe = xlower + (i-1)*dx
+            ye = ylower + (j-1)*dy
+
 c           # Cell-centered velocity
             aux(i,j,1) = vt*ucc(xc,yc)
             aux(i,j,2) = vt*vcc(xc,yc)
+            aux(i,j,3) = vt*ucc(xe,yc)
+            aux(i,j,4) = vt*vcc(xc,ye)
          enddo
       enddo
 
@@ -55,10 +60,10 @@ c         # 1d : Piecewise constant velocity (-0.5, 0.5)
           endif
       elseif (example .eq. 2) then
 c         # 1d : u = cos(2*pi*xp)       
-          ucc = cos(2*pi*xp)
+          ucc = cos(2*pi*xp) + 1.d0
       elseif (example .eq. 3) then
 c         # 1d : More complicated 1d example
-          ucc = 0.1*sin(2*pi*xp)*sin(16*pi*xp)
+          ucc = 0.2*sin(2*pi*xp)*sin(16*pi*xp)
       elseif (example .eq. 4) then 
           a = 0.01d0                   
           ucc = tanh((xp-0.5d0)/a)
@@ -96,10 +101,10 @@ c             # Add non-zero divergence
       if (example .le. 4) then
 c         # 1d examples
           vcc = 0        
-      elseif ((example .eq. 5) .or. (example .eq. 6)) then
+      elseif ((example .eq. 6) .or. (example .eq. 7)) then
 c         # 2d examples
           vcc = -2*((sin(pi*yp))**2 * sin(pi*xp) * cos(pi*xp))
-          if (example .eq. 6) then
+          if (example .eq. 7) then
 c             # Add non-zero divergence              
               rp2 = (xp-0.5d0)**2 + (yp-0.5d0)**2
               rp2 = (xp-0.5d0)**2

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/notes/fwave.tex
(clawpack/swirl (cons)) UPdate write-up

@@
 
+\newcommand{\Fig}[1]{Figure \ref{fig:#1}}
+
 \begin{document}
@@
 \section{Numerical approaches}
+The variable coefficient conservative form of the transport equation arises in many areas of
+geophysical flow modeling.  For example, the transport of pollutants, volcanic ash or 
+other conserved tracer component must often transported using prescibed reanalysis 
+wind field data.
+
+In \cite{le:2002}, LeVeque describes several approaches to solving the conservative, 
+variable coefficient advection equation, given by
+\begin{align}
+q_t + (uq)_x = 0 \\
+\end{align}
+In his book, LeVeque describes several approaches to solving this equation, although
+there is a question about what to do in the case where the velocity field changes sign. 
+In this case, there is a loss of hyperbolicity that cannot be easily addressed by the 
+wave propagation algorithms implemented in ClawPack.  Here, we propose reasonable approaches
+to handling this case.
+
+In the following, we provide details for four of the approaches described, and pay
+particular attention to ways to treat the change of sign that can lead to loss of 
+hyperbolicity.  The accuracy obtained from such methods is second order for velocity
+fields with no sign changes.  For velocity fields with sign changes, the accuracy can be
+significantly degraded, depending on the use of limiters, and whether the velocity field 
+converges or diverges at the sign change.
 
@@
 \begin{align}
-q_t + (uq)_x &= 0 \\
+q_t + (uq)_x = 0 \\
 \end{align}
@@
 The left moves at speed $u_\ell$ and the right moves at speed $u_r$.
-Each conveyor belt has a density of boxes, given by $q_\ell$ and $\q_r$.
+Each conveyor belt has a density of boxes, given by $q_\ell$ and $q_r$.
 If we have $u_\ell > u_r > 0$, boxes will pile up on the right conveyor belt  as faster moving boxes from the left are transfered to the slower moving conveyor belt on the right.
@@
 
-Mathematically, we can see that single wave is not enough by considering the Riemann problem at an interface between conveyor belts.   To solve the Riemann problem, we might try to connect left and 
+We can see that single wave is not enough by considering the Riemann problem at an interface between conveyor belts.   To solve the Riemann problem, we might try to connect left and 
 right states $q_\ell$ and $q_r$ by a single wave moving at some speed $s$ satisfying
@@
 
+The Riemann solver for this approach is described in \Fig{qstar}.
+
+\begin{figure}
+\begin{center}
+\begin{minipage}{0.5\textwidth}
+\vspace{0.5cm}
+\begin{verbatim}
+# Riemann solver for "conveyor belt" approach
+def rpn2(ql,qr,ul,ur):
+    if ur > 0 and ul > 0:
+        wave = (ur*qr - ul*ql)/ur
+        amdq = 0
+        apdq = ur*qr - ul*ql
+        s = ur
+    else if ur < 0 and ul < 0:
+        wave = (ur*qr - ul*ql)/ul
+        amdq = ur*qr - ul*ql
+        apdq = 0
+        s = ul
+    else:
+        uavg = (ur+ul)/2.0
+        uhat = max(abs(uvg),1e-12)
+        wave = (ur*qr - ul*ql)/uhat
+        amdq = -ul*ql
+        apdq = ur*qr
+        s = uhat
+
+    return wave,s,amdq,apdq
+\end{verbatim}
+\end{minipage}
+\end{center}
+\caption{Algorithm for cell-centered velocities based on scalar equations.}
+\label{fig:qstar}
+\end{figure}
+
 \subsection{2-wave decomposition approach.}
+Problem 13.11 in \cite{le:2002}
 Consider the following conservation law for the conservative form of the scalar advection equation.
@@
 1 \end{array}\right) \\
-\mathcal W_2  \equiv \alpha_2 \mathbf r^2 & = 
-\left(q_r - q_\ell + \frac{u_r - u_\ell}{\widehat{u}-a}\widehat{q} \right)
+\mathcal W_2  \equiv \alpha_2 \mathbf r^2 
+& =  \left(q_r - q_\ell + \frac{u_r - u_\ell}{\widehat{u}-a}\widehat{q} \right)
 \left(\begin{array}{c} 1 \\ 0\end{array}\right) 
@@
 
-Following the above approach, we can also find waves for the case $\widehat{u} < a = 0$.   Combining these two cases, we get the following Clawpack Riemann solver.
+The Riemann solver for this approach is described in \Fig{wave_decomp}.
+
+\begin{figure}
+\begin{center}
+\begin{minipage}{0.5\textwidth}
+\vspace{0.25cm}
+\begin{verbatim}
+# Riemann solver for wave decomposition approach
+def rpn2(ql,qr,ul,ur):
+    qhat = (ql + qr)/2.0
+
+    uavg = (ul + ur)/2.0      
+    if uavg == 0:
+        uhat = 1e-12
+    else
+        uhat = uavg
+
+    s = uavg
+    wave = (ur*qr - ul*ql)/uhat
+
+    if uavg == 0:
+        amdq = -ul*qhat     # amdq+apdq = (ur-ul)*qhat
+        apdq = ur*qhat
+    else:
+        apdq = max([uhat,0])*wave
+        amdq = min([uhat,0])*wave
+
+    return wave,s,amdq,apdq
+\end{verbatim}
+\end{minipage}
+\end{center}
+\caption{2-wave decomposition approach to handling conservative advection equation.}
+\label{fig:wave_decomp}
+\end{figure}
 
@@
 
-\begin{minipage}{\textwidth}
-\begin{verbatim}
-# Riemann solver for wave decomposition approach
-def rpn2(ql,qr,ul,ur):
-    uavg = (ul + ur)/2        
-    uhat = sign(uavg)*max([abs(uavg),tol])
-    qhat = (ql + qr)/2
-    wave = qr - ql + (ur-ul)*qhat/uhat
-    if uavg == 0:
-        amdq = -ul*qhat     # amdq+apdq = (ur-ul)*qhat
-        apdq = ur*qhat
-        s = uavg
-    else:
-        apdq = max([uhat,0])*wave
-        amdq = min([uhat,0])*wave        
-        s = uhat
-    return wave,s,amdq,apdq
-\end{verbatim}
-\end{minipage}
-
-\vspace{0.25cm}
+\subsection{Edge velocities}
+In some sense, the simplest approach is to assume that the velocity field at cell edges is available. 
+In this case, we can easily evaluate a flux at cell interfaces, and formulate the problem 
+in classic flux form, where the flux is defined as
+\begin{equation*}
+F_{i-1/2} = \left\{\begin{array}{cc} 
+u_\ell q & u_\ell, u_r < 0 \\
+u_r q & u_r, u_\ell > 0
+\end{array}
+\right.
+\end{equation*}
+where
 

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/swirlcons_user.h

@@ -39,8 +39,7 @@ extern "C"
 typedef struct user_options
 {
     double period;
-    int claw_version;
-    int cons_rp;
+    int rp_solver;
     int example;
     int is_registered;
 
@@ -58,26 +57,29 @@ const user_options_t* swirlcons_get_options(fclaw2d_global_t* glob);
 /* Mappings */
 fclaw2d_map_context_t* fclaw2d_map_new_nomap();
 
-#define RPN2CONS_CC FCLAW_F77_FUNC(rpn2cons_cc,RPN2CONS_CC)
-void RPN2CONS_CC(const int* ixy,const int* maxm, const int* meqn, const int* mwaves,
-                 const int* mbc,const int* mx, double ql[], double qr[],
-                 double auxl[], double auxr[], double wave[],
-                 double s[], double amdq[], double apdq[]);
+#define RPN2CONS_QSTAR FCLAW_F77_FUNC(rpn2cons_qstar,RPN2CONS_QSTAR)
+void RPN2CONS_QSTAR(const int* ixy,const int* maxm, const int* meqn, const int* mwaves,
+                    const int* mbc,const int* mx, double ql[], double qr[],
+                    double auxl[], double auxr[], double wave[],
+                    double s[], double amdq[], double apdq[]);
 
-
-#define RPN2CONS FCLAW_F77_FUNC(rpn2cons,RPN2CONS)
-void RPN2CONS(const int* ixy,const int* maxm, const int* meqn, const int* mwaves,
+#define RPN2CONS_WD FCLAW_F77_FUNC(rpn2cons_wd,RPN2CONS_WD)
+void RPN2CONS_WD(const int* ixy,const int* maxm, const int* meqn, const int* mwaves,
               const int* mbc,const int* mx, double ql[], double qr[],
               double auxl[], double auxr[], double wave[],
               double s[], double amdq[], double apdq[]);
+
+#define RPN2CONS_EC FCLAW_F77_FUNC(rpn2cons_ec,RPN2CONS_EC)
+void RPN2CONS_EC(const int* ixy,const int* maxm, const int* meqn, const int* mwaves,
+                 const int* mbc,const int* mx, double ql[], double qr[],
+                 double auxl[], double auxr[], double wave[],
+                 double s[], double amdq[], double apdq[]);
 
 #define RPN2FWAVE FCLAW_F77_FUNC(rpn2fwave, RPN2FWAVE)
[…]

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/user_4.6/rpn2cons_ec.f

@@ -1,0 +1,48 @@
+      subroutine rpn2cons_ec(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,
+     &      auxl,auxr,wave,s,amdq,apdq)
+      implicit none
+
+      integer maxm, mbc,mwaves,meqn,mx
+      integer ixy
+
+      double precision wave(1-mbc:maxm+mbc, meqn, mwaves)
+      double precision    s(1-mbc:maxm+mbc, mwaves)
+      double precision   ql(1-mbc:maxm+mbc, meqn)
+      double precision   qr(1-mbc:maxm+mbc, meqn)
+      double precision amdq(1-mbc:maxm+mbc, meqn)
+      double precision apdq(1-mbc:maxm+mbc, meqn)
+      double precision auxl(1-mbc:maxm+mbc,*)
+      double precision auxr(1-mbc:maxm+mbc,*)
+
+
+      integer i, iface
+      double precision fstar, uhat, qll,qrr, ul,ur
+
+      iface = ixy
+      do i = 2-mbc, mx+mbc
+
+         ur = auxl(i,iface)
+         ul = auxL(i-1,iface)         
+
+         qrr = ql(i,1)
+         qll = qr(i-1,1)
+
+c         vn = auxl(i,2 + iface)  !! cell centered ?
+         uhat = (ur + ul)/2.d0
+
+         wave(i,1,1) = qrr - qll
+         s(i,1) = uhat
+
+         if (uhat .ge. 0.d0) then
+            fstar = uhat*qll
+         else
+            fstar = uhat*qrr
+         endif
+
+         amdq(i,1) = fstar - ul*qll              !flux
+         apdq(i,1) = ur*qrr - fstar            !-flux
+      enddo
+
+
+      return
+      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dadv/driver.f
(clawpack/1d conservative example) Move 1d example

@@ -1,32 +1,0 @@
-      program driver
-c
-c  Generic driver routine for claw1ez
-c
-c  Author: Randall J. LeVeque
-c  Version of March, 1999 --  CLAWPACK Version 4.0
-c
-c
-      implicit double precision (a-h,o-z)
-
-c     # set parameters for maximum array sizes used in declarations
-c     # these must be increased for larger problems.
-c
-c
-      parameter (maxmx =     8192)
-      parameter (mwork =   122864)
-
-      parameter (mbc = 2)
-      parameter (meqn = 1)
-      parameter (mwaves = 1)
-      parameter (maux = 2)
-c
-      dimension q(1-mbc:maxmx+mbc, meqn)
-      dimension  aux(1-mbc:maxmx+mbc,maux)   !# dummy variable since no aux arrays used
-      dimension work(mwork)
-      dimension mthlim(mwaves)
-
-      call claw1ez(maxmx,meqn,mwaves,mbc,maux,mwork,mthlim,
-     &             q,work,aux)
-
-      stop 
-      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dadv/lib/inlinelimiter.f

@@ -1,108 +1,0 @@
-c
-c
-c     =====================================================
-      subroutine limiter(maxm,meqn,mwaves,mbc,mx,wave,s,mthlim)
-c     =====================================================
-c
-c     # Apply a limiter to the waves.  
-c
-c     # Version of December, 2002.
-c     # Modified from the original CLAWPACK routine to eliminate calls 
-c     # to philim.  Since philim was called for every wave at each cell
-c     # interface, this was adding substantial overhead in some cases.
-c
-c     # The limiter is computed by comparing the 2-norm of each wave with
-c     # the projection of the wave from the interface to the left or
-c     # right onto the current wave.  For a linear system this would
-c     # correspond to comparing the norms of the two waves.  For a 
-c     # nonlinear problem the eigenvectors are not colinear and so the 
-c     # projection is needed to provide more limiting in the case where the
-c     # neighboring wave has large norm but points in a different direction
-c     # in phase space.
-c
-c     # The specific limiter used in each family is determined by the
-c     # value of the corresponding element of the array mthlim.
-c     # Note that a different limiter may be used in each wave family.
-c
-c     # dotl and dotr denote the inner product of wave with the wave to
-c     # the left or right.  The norm of the projections onto the wave are then
-c     # given by dotl/wnorm2 and dotr/wnorm2, where wnorm2 is the 2-norm
-c     # of wave.
-c
-      implicit real*8(a-h,o-z)
-      dimension mthlim(mwaves)
-      dimension wave(1-mbc:maxm+mbc, meqn, mwaves)
-      dimension    s(1-mbc:maxm+mbc, mwaves)
-c
-c
-      do 200 mw=1,mwaves
-         if (mthlim(mw) .eq. 0) go to 200
-         dotr = 0.d0
-         do 190 i = 0, mx+1
-            wnorm2 = 0.d0
-            dotl = dotr
-            dotr = 0.d0
-            do 5 m=1,meqn
-               wnorm2 = wnorm2 + wave(i,m,mw)**2
-               dotr = dotr + wave(i,m,mw)*wave(i+1,m,mw)
-    5          continue
-            if (i.eq.0) go to 190
-            if (wnorm2.eq.0.d0) go to 190
-c
-            if (s(i,mw) .gt. 0.d0) then
-                r = dotl / wnorm2
-              else
-                r = dotr / wnorm2
-              endif
-c
-            go to (10,20,30,40,50) mthlim(mw)
-c
-   10       continue
[…]

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dadv/lib/limiter.f

@@ -1,60 +1,0 @@
-c
-c
-c     =====================================================
-      subroutine limiter(maxm,meqn,mwaves,mbc,mx,wave,s,mthlim)
-c     =====================================================
-c
-c     # Apply a limiter to the waves.
-c     # The limiter is computed by comparing the 2-norm of each wave with
-c     # the projection of the wave from the interface to the left or
-c     # right onto the current wave.  For a linear system this would
-c     # correspond to comparing the norms of the two waves.  For a 
-c     # nonlinear problem the eigenvectors are not colinear and so the 
-c     # projection is needed to provide more limiting in the case where the
-c     # neighboring wave has large norm but points in a different direction
-c     # in phase space.
-c
-c     # The specific limiter used in each family is determined by the
-c     # value of the corresponding element of the array mthlim, as used in
-c     # the function philim.
-c     # Note that a different limiter may be used in each wave family.
-c
-c     # dotl and dotr denote the inner product of wave with the wave to
-c     # the left or right.  The norm of the projections onto the wave are then
-c     # given by dotl/wnorm2 and dotr/wnorm2, where wnorm2 is the 2-norm
-c     # of wave.
-c
-      implicit real*8(a-h,o-z)
-      dimension mthlim(mwaves)
-      dimension wave(1-mbc:maxm+mbc, meqn, mwaves)
-      dimension    s(1-mbc:maxm+mbc, mwaves)
-c
-c
-      do 50 mw=1,mwaves
-         if (mthlim(mw) .eq. 0) go to 50
-         dotr = 0.d0
-         do 40 i = 0, mx+1
-            wnorm2 = 0.d0
-            dotl = dotr
-            dotr = 0.d0
-            do 20 m=1,meqn
-               wnorm2 = wnorm2 + wave(i,m,mw)**2
-               dotr = dotr + wave(i,m,mw)*wave(i+1,m,mw)
-   20          continue
-            if (i.eq.0) go to 40
-            if (wnorm2.eq.0.d0) go to 40
-c
-            if (s(i,mw) .gt. 0.d0) then
-                wlimitr = philim(wnorm2, dotl, mthlim(mw))
-              else
-                wlimitr = philim(wnorm2, dotr, mthlim(mw))
-              endif
-c
-            do 30 m=1,meqn
-               wave(i,m,mw) = wlimitr * wave(i,m,mw)
-   30          continue
-   40       continue
-   50    continue
-c
-      return
-      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dadv/lib/philim.f

@@ -1,58 +1,0 @@
-c
-c
-c     =====================================================
-      double precision function philim(a,b,meth)
-c     =====================================================
-      implicit real*8(a-h,o-z)
-c
-c     # Compute a limiter based on wave strengths a and b.
-c     # meth determines what limiter is used.
-c     # a is assumed to be nonzero.
-c
-c     # NOTE: This routine is obsolete.  Instead of using limiter.f,
-c     # which calls philim.f for every wave, it is more efficient to 
-c     # use inlinelimiter.f, which eliminates all these function calls
-c     # to philim.  If you wish to change the limiter function and are
-c     # using inlinelimiter.f, the formulas must be changed in that routine.
-c
-      r = b/a
-      go to (10,20,30,40,50) meth
-
-c
-   10 continue
-c     --------
-c     # minmod
-c     --------
-      philim = dmax1(0.d0, dmin1(1.d0, r))
-      return
-c
-   20 continue
-c     ----------
-c     # superbee
-c     ----------
-      philim = dmax1(0.d0, dmin1(1.d0, 2.d0*r), dmin1(2.d0, r))
-      return
-c
-   30 continue
-c     ----------
-c     # van Leer
-c     ----------
-      philim = (r + dabs(r)) / (1.d0 + dabs(r))
-      return
-c
-   40 continue
-c     ------------------------------
-c     # monotinized centered 
-c     ------------------------------
-      c = (1.d0 + r)/2.d0
-      philim = dmax1(0.d0, dmin1(c, 2.d0, 2.d0*r))
-      return
-c
-   50 continue
-c     ------------------------------
-c     # Beam-Warming
-c     ------------------------------
-      philim = r
-
-      return
-      end

### 2017-07-26 applications/clawpack/advection/2d/swirl_cons/1dadv/setaux.f

@@ -1,48 +1,0 @@
-      subroutine setaux(maxmx,mbc,mx,xlower,dx,maux,aux)
-
-      implicit double precision (a-h,o-z)
-      dimension aux(1-mbc:maxmx+mbc, maux)
-
-      double precision xe,xc, pi, dx1, dx2, a
-      integer i, example
-
-      common /compi/ pi
-      common /comex/ example
-
-c     # aux(i,1) : edge velocity (u_{i-1/2})
-c     # aux(i,2) : cell-centered derivative of velocity (for src2)
-
-      do i = 1-mbc,mx+mbc
-          xe = xlower + (i-1)*dx
-          xc = xlower + (i-0.5)*dx
-
-          if (example .eq. 1) then
-c             # Velocity should have a delta function at x=0.5
-              if (x .le. 0.5d0) then
-                  aux(i,1) = -0.5d0
-                  aux(i,2) = 0
-              else
-                  aux(i,1) = 0.5d0
-                  aux(i,2) = 0            
-              endif
-          elseif (example .eq. 2) then
-              aux(i,1) = cos(2*pi*xe)
-              aux(i,2) = -2*pi*sin(2*pi*xc)
-          elseif (example .eq. 3) then
-              aux(i,1) = 0.1d0*sin(2*pi*xe)*sin(16*pi*xe)
-              dx1 =  2*pi*cos(2*pi*xc)*sin(16*pi*xc)
-              dx2 = 16*pi*sin(2*pi*xc)*cos(16*pi*xc)
-              aux(i,2) = 0.1*(dx1 + dx2)
-          elseif (example .eq. 4) then
-              a = 0.01d0
-              aux(i,1) = tanh((xe-0.5d0)/a)
-              aux(i,2) = (1-tanh((xc-0.5d0)/a)**2)/a
-          elseif (example .eq. 5) then
-              a = 0.1d0
-              aux(i,1) = -tanh((xe-0.5d0)/a)
-              aux(i,2) = -(1-tanh((xc-0.5d0)/a)**2)/a
-          endif
-      enddo
-
-      return
-      end
